---
title: "Analyses for Symmetry"
output: html_notebook
---

I wrote really quick code to get the percentage of each interval above and below the optimum - it's shown below using the new dataset and the P metric, though it could easily be used for the S and D metrics. It ended up being really simple, given the new summary table!

Contents
1. Import data and housekeeping
2. Exploratory plots
3. Bias in performance
4. Bias in development (not in presentation)
5. Bias in survival (not in presentation)


```{r}
# load packages
library(tidyverse)
library(readxl)
library(tidyverse)
library(readxl)
library(cowplot)
library(psych) # to get geometric mean
library(lme4)
library(car)
library("dotwhisker")  # to make coefficient plots
library(sjPlot) #prediction plots
library(sjmisc)
library(broom)
```

# 1. Import data and housekeeping

```{r}
#Anna's Laptop

data <- read_csv("~/Desktop/ThermalResponsesII.csv")

head(data)
```

```{r}
# Mariana's desktop

data <- read_csv("/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/ThermalResponsesII.csv")

```


```{r}
# convert character to factor
data <- data %>%
  mutate_if(is.character, factor)
# discard non-pertinent data

data <- data %>% 
  filter(lifestage == "egg"| lifestage == "larva"| lifestage == "pupa", quality!= "inferred", quality != "combination") %>% 
  mutate(absLat = abs(lat))
data$lifestage <- factor(data$lifestage)
data$sp <- factor(data$sp)

```



```{r}
data$belowPopt <- ((abs(data$Plower - data$Popt)) / (data$Pwidth))
data$abovePopt <- ((data$Pupper - data$Popt) / (data$Pwidth))
data$aboveP <- data$Pupper - data$Popt
data$belowP <- data$Popt - data$Plower
data$biasP <- data$aboveP - data$belowP


data$aboveD <- data$Dupper - data$Dopt
data$belowD <- data$Dopt - data$Dlower
data$biasD <- data$aboveD - data$belowD


data$aboveS <- data$Supper - data$maxSopt
data$belowS <- data$maxSopt - data$Slower
data$biasS <- data$aboveS - data$belowS





head(data)
data_nona <- filter(data, !is.na(belowPopt))
```


# 2. Exploratory plots
```{r}
ggplot(data, aes(x = lifestage, y = biasP))+
  geom_boxplot()#+
  #geom_boxplot(data = data, mapping = aes(lifestage, y = biasD), col = "blue")+
  #geom_boxplot(data = data, mapping = aes(lifestage, y = biasS), col = "darkred")

```



```{r}
ggplot(data, aes(x = lifestage, y = biasP))+
  geom_jitter(col = "darkgreen")+
  geom_jitter(data = data, mapping = aes(lifestage, y = biasD), col = "blue")+
  geom_jitter(data = data, mapping = aes(lifestage, y = biasS), col = "darkred")


```



```{r}


ggplot(data, aes(x = Pwidth, y = biasP))+
  geom_point()

```



```{r}

ggplot(data, aes(x = Max_dr, y = biasP))+
  geom_point()

```



```{r}
ggplot(data, aes(x = absLat, y = biasP))+
  geom_point(col = "darkgreen")+
  geom_point(data = data, mapping = aes(absLat, y = biasD), col = "blue")+
  geom_point(data = data, mapping = aes(absLat, y = biasS), col = "darkred")
```

# 3. Bias in performance
```{r}
A.set <- with(data_nona, cbind(biasP, belowPopt, abovePopt, Max_dr, lifestage, absLat, sp))

# Full models, with random intercept, and then random slopes

## random slopes: need to input dataset as dataframe
 fit1_fullrandom <- lmer(biasP ~ lifestage + absLat + lifestage:absLat + (lifestage|sp), data = as.data.frame(A.set)) # This model doesn't work

```



```{r}
# models with species (1|species) as a random effect

# Full model

fullP <- lmer(biasP ~ lifestage + absLat + lifestage:absLat + (1|sp), data = data) #
latlifeP <- lmer(biasP ~ absLat + lifestage + (1|sp), data = data) #
latP <- lmer(biasP ~ absLat + (1|sp), data = data) #
lifeP <- lmer(biasP ~ lifestage + (1|sp), data = data) #

anova(fullP, latlifeP, latP, lifeP, test.statistic = "Chisq")

fullP <- lmer(biasP ~ -1 + absLat + lifestage  + lifestage:absLat+ (1|sp), REML = F, data = data) #



Anova(fullP)


summary(fullP)
```
 
```{r}
# get data for plot
latlifeI_effects <- tidy(fullP)
Esizes_wide <- tibble(o_mean = mean(data$biasP, na.rm = T),
                   lat = filter(latlifeI_effects, term == "absLat")[[2]], 
                   lat10 = filter(latlifeI_effects, term == "absLat")[[2]]*10, 
                   egg = filter(latlifeI_effects, term == "lifestageegg")[[2]] - o_mean, 
                   larva = filter(latlifeI_effects, term == "lifestagelarva")[[2]] - o_mean, 
                   pupa = filter(latlifeI_effects, term == "lifestagepupa")[[2]] - o_mean, 
                   SD_sp = filter(latlifeI_effects, term == "sd_(Intercept).sp")[[2]])

names(Esizes_wide)

Esizes <- pivot_longer(Esizes_wide, c("o_mean", "lat", "lat10", "egg", "larva", "pupa", "SD_sp"), names_to = "effect", values_to = "degrees")
Esizes$effect <- factor(Esizes$effect, levels = c("lat", "lat10", "egg", "larva", "pupa", "SD_sp"))

ggplot(filter(Esizes, effect != "o_mean"), aes(x = effect, y = degrees))+
  geom_bar(stat = "identity")+
  theme_cowplot()#+
  #ylim(-2,6)


```
 
 
```{r}
# Plot data and predictions


data$group <- data$lifestage
pred <- ggpredict(fullP, terms = c("absLat","lifestage"))


ggplot(pred, aes(x = x, y = predicted, colour = group, fill = group)) +
  geom_line()+
   geom_ribbon(aes(ymin = conf.low, ymax = conf.high, col = group), alpha = 0.1)+
  scale_color_viridis_d()+
  scale_fill_viridis_d()+
  ylab("Performance skew")+
  xlab("Absolute latitude")+
 # ylim(5, 20)+
  geom_point(data = data, mapping = aes(x = absLat, y = biasP, col =group ), alpha = 0.6)+
  theme_cowplot()


```
 
 
```{r}
#type = c("est", "re", "eff", "pred", "int", "std", "std2", "slope", "resid", "diag")
plot_model(fullP, type = "pred", terms = c("lifestage", "absLat"))+
  theme_cowplot()+
  scale_color_viridis_d()
```




# 4. Bias in development (not in presentation)


```{r}
# models with species (1|species) as a random effect

# Full model

fullD <- lmer(biasD ~ lifestage + absLat + lifestage:absLat + (1|sp), data = data) #
latlifeD <- lmer(biasD ~ absLat + lifestage + (1|sp), data = data) #
latD <- lmer(biasD ~ absLat + (1|sp), data = data) #
lifeD <- lmer(biasD ~ lifestage + (1|sp), data = data) #

anova(fullD, latlifeD, latD, lifeD, test.statistic = "Chisq")

latlifeD <- lmer(biasD ~ -1 + absLat + lifestage + (1|sp), REML = F, data = data) #



Anova(latlifeD)


summary(latlifeD)
```
 
```{r}
# get data for plot
latlifeI_effects <- tidy(latlifeD)
Esizes_wide <- tibble(o_mean = mean(data$biasD, na.rm = T),
                   lat = filter(latlifeI_effects, term == "absLat")[[2]], 
                   lat10 = filter(latlifeI_effects, term == "absLat")[[2]]*10, 
                   egg = filter(latlifeI_effects, term == "lifestageegg")[[2]] - o_mean, 
                   larva = filter(latlifeI_effects, term == "lifestagelarva")[[2]] - o_mean, 
                   pupa = filter(latlifeI_effects, term == "lifestagepupa")[[2]] - o_mean, 
                   SD_sp = filter(latlifeI_effects, term == "sd_(Intercept).sp")[[2]])

names(Esizes_wide)

Esizes <- pivot_longer(Esizes_wide, c("o_mean", "lat", "lat10", "egg", "larva", "pupa", "SD_sp"), names_to = "effect", values_to = "degrees")
Esizes$effect <- factor(Esizes$effect, levels = c("lat", "lat10", "egg", "larva", "pupa", "SD_sp"))

ggplot(filter(Esizes, effect != "o_mean"), aes(x = effect, y = degrees))+
  geom_bar(stat = "identity")+
  theme_cowplot()#+
  #ylim(-2,6)


```
 
 
```{r}
# Plot data and predictions
library(effects)
library(ggeffects)

data$group <- data$lifestage
pred <- ggpredict(latlifeD, terms = c("absLat","lifestage"))


ggplot(pred, aes(x = x, y = predicted, colour = group, fill = group)) +
  geom_line()+
   geom_ribbon(aes(ymin = conf.low, ymax = conf.high, col = group), alpha = 0.1)+
  scale_color_viridis_d()+
  scale_fill_viridis_d()+
  ylab("BiasD")+
  xlab("Absolute latitude")+
 # ylim(5, 20)+
  geom_point(data = data, mapping = aes(x = absLat, y = biasD, col =group ), alpha = 0.6)+
  theme_cowplot()


```
 
 
```{r}
#type = c("est", "re", "eff", "pred", "int", "std", "std2", "slope", "resid", "diag")
plot_model(latlifeD, type = "pred", terms = c("lifestage", "absLat"))+
  theme_cowplot()+
  scale_color_viridis_d()
```



# 5. Bias in survival (not in presentation)

```{r}
# models with species (1|species) as a random effect
dataS <- filter(data, !is.na(biasS))
dataS$sp <- factor(dataS$sp)
dataS$lifestage <- factor(dataS$lifestage)
# Full model

fullS <- lmer(biasS ~ lifestage + absLat + lifestage:absLat + (1|sp), data = data) #
latlifeS <- lmer(biasS ~ absLat + lifestage + (1|sp), data = data) #
latS <- lmer(biasS ~ absLat + (1|sp), data = data) #
lifeS <- lmer(biasS ~ lifestage + (1|sp), data = data) #
# Models don't work
#anova(fullP, latlifeP, latP, lifeP, test.statistic = "Chisq")

#fullP <- lmer(biasP ~ -1 + absLat + lifestage + (1|sp), REML = F, data = data) #
```
 
 
```{r}
hist(data_nona$biasP)
positive <- filter(data_nona, biasP > 0)
id_pos <-unique(positive$set) # 1035 1122 1128 1136 1140 1197 1206
# You need to run other scripts to get Ana_performance_consolidated table
#posi1 <- inner_join(Ana_performance_consolidated, add_info, by = "set")

# posi <- posi1 %>% filter(set %in% id_pos)
# 
# ggplot(posi, aes(x = temp, y = P2))+
#   geom_point()+
#   theme_cowplot()+
#   facet_wrap(lifestage~sp)
# 
# add_info <- Ana_performance%>% 
#   select(set, sp, family, lifestage, lat, lon, locality, quality) %>% 
#   distinct()
# 
# Popt_table <- inner_join(singles, add_info, by = "set")
```
 
 
 
 