---
title: "Anna's Working Script"
output: html_notebook
---

Here, I will put working code for any side projects I'm working on, which will eventually be added to Report.rmd 

```{r}
library(tidyr)
```


## Working on making the dataset wide instead of long to calculate cumulative survival. 

```{r}
head(combsurvival)
```

```{r}
combsurvival2 <- combsurvival[, c("set", "id", "sp", "lifestage", "temp", "survival")]
head(combsurvival2)
```


Make it wider:
```{r}
combsurvival3 <- combsurvival2 %>% 
  pivot_wider(names_from = lifestage, values_from = survival)

head(combsurvival3)
```

Combine data into one row for each temp:

(This is a very round-about way to do this - if you have a smoother suggestion, please implement it instead of this mess!!)

```{r}
#Separate them into different dataframes

eggsurv <- combsurvival3[, c("id", "sp", "temp", "egg")]
larvasurv <- combsurvival3[, c("id", "sp", "temp", "larva")]
pupasurv <- combsurvival3[, c("id", "sp", "temp", "pupa")]

#Remove rows with NAs

eggsurv <- eggsurv %>% 
  drop_na()

larvasurv <- larvasurv %>%
  drop_na()

pupasurv <- pupasurv %>%
  drop_na()

# Merge into one dataframe

combsurvival4 <- merge(eggsurv, larvasurv, by = c("id", "sp", "temp"))
combsurvival5 <- merge(combsurvival4, pupasurv, by = c("id", "sp", "temp"))
```


Make cumulative survival column:
```{r}
combsurvival5$total <- (((combsurvival5$egg)/100) * ((combsurvival5$larva)/100) * ((combsurvival5$pupa)/100))
combsurv <- combsurvival5
```

Here we have it! 

Now, we plot!

## Plotting Cumulative survival at different temperatures

```{r}
ggplot(combsurv, aes(x = temp, y = total, group = id))+
  geom_point() +
  geom_line() +
  theme(legend.position = "none") +
  labs(x = "Temperature", y = "Cumulative Survival across lifestages") + 
  facet_wrap(id~sp)
```

