---
title: "Anna's Working Script - Summary Graphs"
output: html_notebook
---

## Load in the dataset in the same manner as "Report.Rmd"

```{r}
library(tidyverse)
library(readxl)
library(cowplot)
# Anna's laptop
Data <- read_xlsx("~/Desktop/PhysiologyDatabaseVersion5.xlsx", 
                  sheet = "T3", na = c("NA", "")) 
unique(Data$status)
Ana <- Data[Data$status != "parasitoid",]
rm(Data)

# convert character to factor
Ana <- Ana %>%
  mutate_if(is.character, factor)
```

### Calculate development rate
When all larvae in a treatment die, their development time is "NA", however
their development rate should be "0". The loop below calculates development rate (1/dt) and assigns 0 to all cases where both survival is 0 and development time is "NA".
There is one case in which development time is less than 1, it was rounded to 1.

```{r}
fast <- filter(Ana, dt < 1)
Ana <- Ana %>% 
  mutate(dt = ifelse(dt < 1, ceiling(dt),dt))
  

sets <- unique(Ana$set) 
rates <- tribble(~set, ~temp, ~dr)
rates2 <- tribble(~set, ~temp, ~dr)
output.table <- tibble()

for(i in seq_along(sets)) {
  
  # make a table per set and determine whether it includes dt data
  seti <- sets[i]
  table <- filter(Ana, set == seti)
  suma <- sum(table$dt, na.rm = T)
  
  # if there is dt data
  if (suma > 0){ 
    
    clement <- filter(table, dt > 0) [["temp"]] # temperatures that allow for development
    hotlim <- max(clement, na.rm = T) # max temp that permits development
    coldlim <- min(clement, na.rm = T) # min temp that permits development
    tempes <- unique(table$temp) # each temperature included in set i
    
    # evaluate each temperature of a set
    for(ii in 1:length(unique(tempes))){
      
      tempii <- tempes[ii]
      
      # if it is lower than coldlim, dr should be zero
      if (table$temp[ii] < coldlim){
        dr <- 0
      output.row <- cbind(set = seti, temp = tempii, dr = dr)
      
      # if it is higher than hotlim, dr should be zero  
      } else if (table$temp[ii] > hotlim) {
        
        dr <- 0
      output.row <- cbind(set = seti, temp = tempii, dr = dr)
      
      # in all other cases, it should be 1/dt
      } else{
        dr <- 1/table$dt[ii]
        output.row <- cbind(set = seti, temp = tempii, dr = dr)
      }
      
      output.table <- rbind(output.table, output.row)
    }
    
    # print stuff to make sure loop is working
    #cat(seti, " ")
    
    
  # if there is no development time in that set, dr should be zero
    } else {
    #cat("set: ", seti, "has no dt \n")
  }
  
}

rates <- output.table

Ana_rates <- left_join(Ana, rates, by = c("set","temp"))

rm(fast, output.row, output.table, rates, rates2, table)
rm(clement, coldlim, dr, hotlim, i, ii, seti, sets, suma, tempes, tempii)

```

## Sets with both development rate and survival

```{r}
interval_sets <- Ana_rates %>% 
  group_by(set) %>% 
   
  mutate(dr_sum = sum(dr), 
         n_temps = length(unique(temp)), 
         validcount = sum(!is.na(dt)), 
         validcount_s = sum(!is.na(survival))) %>% 
  filter(n_temps > 3, dr_sum > 0, validcount > 3, validcount_s > 3)

interval_set_list <- unique(interval_sets$set) 

interval <- Ana_rates %>% filter(set %in% interval_set_list)

main_stages <- Ana_rates %>% 
  filter(lifestage == "egg"|lifestage == "larva"|lifestage == "pupa")
main_stages_list <- unique(main_stages$lifestage)
inter <- interval %>% filter(lifestage %in% main_stages_list) 
inter$lifestage <- factor(inter$lifestage)

```
There are `r length(unique(interval$set))` sets with both survival and
development time data, from `r length(unique(interval$sp))` families and
`r length(unique(interval$lifestage))` lifestages. 

For the three main lifestages (egg, larva, pupa), there are `r length(unique(inter$set))` sets with both survival and
development time data, from `r length(unique(inter$sp))` species.

**Using dataset "inter" for future analyses** 


## Exploratory Graphs
These graphs represent some early data exploration, using the subset of the data that contains both survival and development time for the three main lifestages (egg, larva, pupa). 

### Subset the data by lifestage (Anna)

### Graph the survival data vs. temperature within each lifestage, with each set as a line, faceted by family (Anna will attempt)

### Graph the development time data vs. temperature within each lifestage, with each set as a line, faceted by family (Anna will attempt)