---
title: "Cumulative survival"
output: html_notebook
---

## Contents
1. Import data and housekeeping

```{r}
# load packages
library(tidyverse)
library(readxl)
library(cowplot)
library(psych) # to get geometric mean
library(lme4)
library(car)
library("dotwhisker")  # to make coefficient plots
library(sjPlot) #prediction plots
library(sjmisc)
library(broom)
```


Import data 
```{r}
# Mariana's desktop
Data <- read_xlsx("/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/PhysiologyDatabaseVersion5.xlsx", 
                  sheet = "WorkingTable", na = c("NA", ""))
```

```{r}
# Mariana's laptop
Data <- read_xlsx("/Users/mar/Desktop/Projects/ThermalPerformance/PhysiologyDatabaseVersion5.xlsx", 
                  sheet = "WorkingTable", na = c("NA", "")) 

```


```{r}
# Anna's Laptop
Data <- read_xlsx("~/Desktop/PhysiologyDatabaseVersion5.xlsx", 
                  sheet = "WorkingTable", na = c("NA", "")) 

```


```{r}
# Discard parasitoids and change character to factor 
unique(Data$status)
Ana <- Data[Data$status != "parasitoid",]
rm(Data)

# convert character to factor
Ana <- Ana %>%
  mutate_if(is.character, factor)
```


Consolidate redundant sets
Redundant sets occur beacuse researchers varied experimental conditions such as host plant, humidity and daylength, or because data were reported for males and females separately. We considered multiple measurements reported in the same publication that were obtained at the same temperature, of the same ontogenetic stage of individuals of the same population (sp and locality) to be redundant, and we obtained the geometric mean of all treatments (host, sex, daylength, humidity,etc) at the same temperature for analyses. 
Consolidated sets kept the number (unique set identifier) of the first set they include.


```{r}
# make a new column that unifies redundant sets
Ana_survival <- Ana %>% 
  mutate(redundancy_id = as.numeric(factor(paste(sp, locality, lifestage, id))))

#length(unique(Ana_survival$set)) # there were 156 sets
#length(unique(Ana_survival$redundancy_id)) # there were 137 non-redundant sets

# Function to get the geometric mean of development rate (dr_c), and mean (dr2)
mean_redundant <- function(dat){
newset <- dat$set [[1]]
dat %>%
  select(temp, sp, survival) %>% 
  group_by(temp) %>% 
  summarise(survival_c = geometric.mean(survival, na.rm = T),
            survival2 = mean(survival,  na.rm = T), 
            newset = newset,
            n_sets = length(survival)) 
}

# Apply function to Ana_survival

Ana_survival_consolidated <- Ana_survival %>% 
  group_by(redundancy_id) %>% 
  nest() %>% 
  mutate(consolidated = map(data, mean_redundant)) %>% 
  select(redundancy_id, consolidated) %>% 
  unnest(col = consolidated)# Function

# re-name variables. Note that consolidated sets summarize multiple sets and got assigned the number (set id value) of the first set.
names(Ana_survival_consolidated) <- c("redundancy_id", "temp", "survival", "survival2", "set", "n_sets")
```

```{r}
add_info <- Ana_survival %>% 
  select(set, sp, family, lifestage, lat, lon, locality, quality, id) %>% 
  distinct()

Survival_table <- inner_join(Ana_survival_consolidated, add_info, by = "set")
```





```{r}
# Select sets of life stages of interest: egg, larva, pupa, egg to emergence


life_stages <- c("egg", "larva", "pupa", "eggtoemergence")

Ana_survival <- Ana %>% 
  filter(lifestage %in% life_stages) # and only the life stages of interest

Ana_survival$lifestage <- factor(Ana_survival$lifestage, levels = c("egg", "larva", "pupa", "eggtoemergence"))
```



# get ids of sets that report survival in all three life stages

```{r}
interval_sets <- Ana_survival %>% 
  group_by(set) %>% 
  mutate(n_temps = length(unique(temp)), 
         validcount = sum(!is.na(survival)), 
         validcount_s = sum(!is.na(survival))) %>% 
  filter(n_temps > 3, validcount > 3, validcount_s > 3)

interval_set_list <- unique(interval_sets$set) 

interval <- Ana_survival %>% filter(set %in% interval_set_list)

main_stages <- Ana_survival %>% 
  filter(lifestage == "egg"|lifestage == "larva"|lifestage == "pupa")
main_stages_list <- unique(main_stages$lifestage)
inter <- interval %>% filter(lifestage %in% main_stages_list) 
inter$lifestage <- factor(inter$lifestage)


egg <- filter(inter, lifestage == "egg", survival >= 0)
egg$set <- factor(egg$set)
larva <- filter(inter, lifestage == "larva",  survival >= 1)
larva$set <- factor(larva$set)
pupa <- filter(inter, lifestage == "pupa",  survival >= 1)
pupa$set <- factor(pupa$set)

egg_ids <- tibble(id = unique(egg$id))
larva_ids <- tibble(id = unique(larva$id))
pupa_ids <- tibble(id = unique(pupa$id))

fullstudies1 <- intersect(egg_ids,larva_ids)
fullstudies <- intersect(fullstudies1,pupa_ids)
# the new table "ontosurvival" contains only those sets
ontosurvival <- inter %>% filter(id %in% fullstudies$id)
ontosurvival$stagenum <- ifelse(ontosurvival$lifestage == "egg", 1, ifelse(ontosurvival$lifestage == "larva",2, 3))

```



```{r}
# select only relevant columns (no tidyverse, so "set" can be removed as the grouping variable)
combsurvival <- ontosurvival %>% 
  filter(sp != "Chilo auricilius", sp != "Metisa plana", sp !="Pteroma pendula")
combsurvival2 <- combsurvival[, c("id","sp", "lifestage", "temp", "survival","family")]
head(combsurvival2)


```

Make it wider:
```{r}
combsurvival <- combsurvival2 %>% 
  ungroup() %>% # by ungrouping you get rid of the issue of having a bunch of NA and non- matching rows
  pivot_wider(names_from = lifestage, values_from = survival) 


```

Ploting Cumulative survival at different temperatures

```{r}
names(combsurvival)
combsurvival$total <- combsurvival$egg/100 * combsurvival$larva/100 * combsurvival$pupa/100
combsurvival$sp_id <- factor(paste(combsurvival$sp, combsurvival$id, sep = "-"))
combsurvival$family <- factor(combsurvival$family)
ggplot(combsurvival, aes(x = temp, y = total, group = factor(id)))+
  geom_point(col = "black") +
  geom_point(data = combsurvival, mapping = aes(x = temp, y = egg/100, group = factor(sp_id)), col = "grey", shape = 1) +
  geom_line(data = combsurvival, mapping = aes(x = temp, y = egg/100, group = factor(sp_id)), col = "grey") +
  
  geom_point(data = combsurvival, mapping = aes(x = temp, y = larva/100, group = factor(sp_id)), col = "grey70", shape = 2) +
  geom_line(data = combsurvival, mapping = aes(x = temp, y = larva/100, group = factor(sp_id)), col = "grey70", linetype = 2)+

geom_point(data = combsurvival, mapping = aes(x = temp, y = pupa/100, group = factor(sp_id)), col = "grey32", shape = 4) +
  geom_line(data = combsurvival, mapping = aes(x = temp, y = pupa/100, group = factor(sp_id)), col = "grey32", linetype = 3)+
  
  geom_line( col = "black") +
  theme_cowplot()+
  theme(legend.position = "none") +
  labs(x = "Temperature", y = "Cumulative Survival across lifestages") + 
  facet_wrap(.~sp_id)

```


```{r}
# families in comsurvival: Pyralidae      Noctuidae      Tortricidae    Lycaenidae     Crambidae Arctiidae      Oecophoridae   Plutellidae    Elachistidae   Depressariidae Geometridae    Bellelliidae
unique(combsurvival$family)
combsurvival %>% 
  group_by(family) %>% 
  tally(sort = T)

xfamily <- filter(combsurvival, family != "Crambidae", family !="Pyralidae", family!= "Tortricidae",family != "Lycaenidae")

xfamily <- filter(combsurvival, family == "Pyralidae")


ggplot(xfamily, aes(x = temp, y = total, group = factor(id)))+
  geom_point(col = "black") +
  geom_point(data = xfamily, mapping = aes(x = temp, y = egg/100, group = factor(sp_id)), col = "grey", shape = 1) +
  geom_line(data = xfamily, mapping = aes(x = temp, y = egg/100, group = factor(sp_id)), col = "grey") +
  
  geom_point(data = xfamily, mapping = aes(x = temp, y = larva/100, group = factor(sp_id)), col = "grey70", shape = 2) +
  geom_line(data = xfamily, mapping = aes(x = temp, y = larva/100, group = factor(sp_id)), col = "grey70", linetype = 2)+

geom_point(data = xfamily, mapping = aes(x = temp, y = pupa/100, group = factor(sp_id)), col = "grey32", shape = 4) +
  geom_line(data = xfamily, mapping = aes(x = temp, y = pupa/100, group = factor(sp_id)), col = "grey32", linetype = 3)+
  
  geom_line( col = "black") +
  theme_cowplot()+
  #theme(legend.position = "none") +
  labs(x = "Temperature", y = "Cumulative Survival across lifestages") + 
  facet_wrap(.~sp_id)
```


```{r}

xfamily <- filter(combsurvival2, family == "Pyralidae")

ggplot(xfamily, aes(x = temp, y = survival, col = lifestage, shape = lifestage))+
  geom_point() +
  geom_line()+
  theme_cowplot()+
  #theme(legend.position = "none") +
  labs(x = "Temperature", y = "Cumulative Survival across lifestages") + 
  facet_grid(sp~lifestage)


```


Eggs

```{r}
Eggs <- filter(Ana_survival, lifestage == "egg")
Eggs$family <- factor(Eggs$family)
Eggs$sp_id <- factor(paste(Eggs$sp, Eggs$id, sep = "-"))
ggplot(Eggs, aes(x = temp, y = survival/100, col = sp_id))+
  geom_point() +
  geom_line()+
  theme_cowplot()+
  theme(legend.position = "none")+
  facet_wrap(.~family)


```

```{r}
ggplot(Eggs, aes(x = temp, y = survival/100, col = factor(set)))+
#ggplot(filter(Eggs, family == "Pyralidae"), aes(x = temp, y = survival/100, col = factor(set)))+
  #geom_point()+
  geom_line()+
  theme_cowplot()+
  theme(legend.position = "none")#+
  #facet_wrap(.~family)



```

Larvae

```{r}
Larvae <- filter(Ana_survival, lifestage == "larva")
Larvae$family <- factor(Larvae$family)
Larvae$sp_id <- factor(paste(Larvae$sp, Larvae$id, sep = "-"))

ggplot(Larvae, aes(x = temp, y = survival/100, col = factor(set)))+
#ggplot(filter(Larvae, family == "Pyralidae"), aes(x = temp, y = survival/100, col = factor(set)))+
  xlim(-10,40)+
  geom_line()+
  theme_cowplot()+
  theme(legend.position = "none")#+
  #facet_wrap(.~family)


```

Pupae

```{r}
Pupae <- filter(Ana_survival, lifestage == "pupa")
Pupae$family <- factor(Pupae$family)
Pupae$sp_id <- factor(paste(Pupae$sp, Pupae$id, sep = "-"))

ggplot(Pupae, aes(x = temp, y = survival/100, col = factor(set)))+
  #geom_point()+
  geom_line()+
  theme_cowplot()+
  xlim(-10,40)+
  theme(legend.position = "none")#+
  #facet_wrap(.~family)


```

```{r}
# Most common temps
Ana_common <- filter(Ana_survival, temp == 5| temp == 10| temp == 15| temp == 20| temp == 25| temp == 30| temp == 35| temp == 40)
sum(!is.na(Ana_common$survival))
length(unique(Ana_common$sp))
ggplot(Ana_common, aes(x = factor(temp), y = survival))+
  geom_boxplot()+
  theme_cowplot()+
  #geom_jitter()+
  facet_wrap(.~ lifestage)
```

