---
title: "GLM curve fitting to estimate the favorable interval"
output: html_notebook
---
Fit a binomial curve using glm to survival and development time data

```{r}
# load packages
library(tidyverse)
library(readxl)
library(cowplot)
```

Import data
```{r}
# Mariana's desktop
Data <- read_xlsx("/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/PhysiologyDatabaseVersion5.xlsx", 
                  sheet = "WorkingTable", na = c("NA", "")) 
SetQuality <- read_csv("SetCharacterization.csv")
# cols 1 to 7 correspond to dt, 8 to 13 to survival, 14 to development time

```

```{r}
# Mariana's laptop
Data <- read_xlsx("/Users/mar/Desktop/Projects/ThermalPerformance/PhysiologyDatabaseVersion5.xlsx", 
                  sheet = "WorkingTable", na = c("NA", "")) 
SetQuality <- read_csv("SetCharacterization.csv")
# cols 1 to 7 correspond to dt, 8 to 13 to survival, 14 to development time

```



```{r}
# Discard parasitoids and change character to factor 
unique(Data$status)
Ana <- Data[Data$status != "parasitoid",]
rm(Data)

# convert character to factor
Ana <- Ana %>%
  mutate_if(is.character, factor)
```

# Favorable interval estimation. 
We included only sets with both development time and survival data
```{r}
interval_set_list <- unique(SetQuality$set) 

inter <- Ana %>% filter(set %in% interval_set_list)

complete_dr_list <- SetQuality %>% 
 filter(just.rise == FALSE, is.min.at.hottest == FALSE) # only just.rise: 114, both: 109

complete_survival_list <- SetQuality %>% 
 filter(just.rise1 == FALSE) # 148
complete_list <- intersect(complete_dr_list$set, complete_survival_list$set)

complete <- Ana %>% filter(set %in% complete_list)

```
There are `r length(unique(factor(inter$set)))` sets from `r length(unique(factor(inter$sp)))` species. Of these, there are `r length(unique(factor(complete_dr_list$set)))` sets with complete data for development, `r length(unique(factor(complete_survival_list$set)))` for survival and 
 `r length(unique(factor(complete_list)))` for both variables from `r length(unique(factor(complete$sp)))` species of `r length(unique(complete$family))` Lepidoptera families: `r unique(factor(complete$family))`
 
#  Development rate, adjusted development rate, and scaled development time.
Choose one of the above to estimate performance from development time data.
dr and adjusted dr have already been calculated.
Function to extract longest development time and create a new column with scaled development time:
```{r}
scale.dt <- function(table){
  maxdt <- max(table$dt, na.rm = T)
  table$s_dt <- 1 - table$dt/maxdt
  table
}
```

Apply the function to each set
```{r}
inter <- inter %>% 
  group_by(set) %>% 
  nest() %>% 
  mutate(s_dt = map(data, scale.dt)) %>% 
  unnest(cols = data,s_dt)
```


 Functions to fit binomial curves and calculate predicted dr and survival. Adjust if using scaled development time.
```{r}
# 1. Model: development rate ~ temperature 
fit_dt <- function(df){
  glm(dr ~ temp + I(temp^2), data = df, family = quasibinomial)
}
# 2. Model: survival ~ temperature 
fit_survival <- function(df) {
  glm(survival/100 ~ temp + I(temp^2), data = df, family = quasibinomial)
}

# 3. Calculate predicted values over the full temperature range
get_predicted <- function(model){
  pred <- tibble(temp = c(1:40), 
                 predicted = predict.glm(model, newdata = tibble(temp = c(1:40)), type = "response"))
}

# 4. Extract model coefficients
get_coefficients <- function(model){
  tibble(intercept = coefficients(model)[1], 
         temperature = coefficients(model)[2], 
         tempsquare = coefficients(model)[3])
}

# 5. Plot observed points and model line

plo_dr <- function(df, predf){
  ggplot(df, aes(x = temp, y = dr))+
    geom_point()+
    theme_cowplot()+
    #ggtitle(set)+
    geom_line(data = predf, mapping = aes(x = temp, y = predicted), col = "darkgrey")
}

plo_s <- function(df, predf){
  ggplot(df, aes(x = temp, y = survival/100))+
    geom_point()+
    theme_cowplot()+
    #ggtitle(set)+
    geom_line(data = predf, mapping = aes(x = temp, y = predicted), col = "darkgrey")
}



```


Apply functions, obtain predictions and calculate the favorable interval

```{r}
# 3. Calculate Fl: length, Fmin: min and Fmax: max ----------------------------------------------------------
unique(inter$lifestage)
# 3.1. nest data (divide into analysis units)
nested <- complete %>% 
  group_by(set) %>% 
  nest()

# 3.2 Apply the model to each set, save predicted and coefficients

nested <- nested %>%
  mutate(model_dt = map(data, fit_dt), 
         predicted_dt = map(model_dt, get_predicted),
         coefficients_dt = map(model_dt, get_coefficients), 
         model_s = map(data, fit_survival),
         predicted_s = map(model_s, get_predicted),
         coefficients_s = map(model_s, get_coefficients))


v_nested <- nested %>%
  mutate(plot_dr = map2(data,predicted_dt, plo_dr), 
         plot_s = map2(data, predicted_s, plo_s))

extract_predictions1 <-v_nested %>% 
  select(set, predicted_dt) %>% 
  unnest(cols = c("predicted_dt")) %>% 
  transmute(temperature = temp, predicted_dt = predicted)

extract_predictions2 <-v_nested %>% 
  select(set, predicted_s) %>% 
  unnest(cols = c("predicted_s")) %>% 
  transmute(temperature = temp, predicted_s = predicted)

predictions <- merge(extract_predictions1, extract_predictions2)




get_performance <- function(predicted){
  #predicted <- enquo(predicted)
  Max_dr <-  filter(predicted, predicted_dt == max(predicted_dt))[["predicted_dt"]] # maximum development rate (value)
  Topt <-  filter(predicted, predicted_dt == max(predicted_dt))[["temperature"]] # maximum development rate (value)
  Inte <- filter(predicted, predicted_dt >= Max_dr/2 & predicted_s >= 0.75)
  Imin <-  Inte[1, "temperature"][[1]]
  Imax <-  tail(Inte, n = 1) [["temperature"]]
  performance <- tibble(Maxdr = Max_dr,
                        Imin = Imin,
                        Imax = Imax,
                        Topt = Topt)
}



nestpre <-  predictions %>% 
  group_by(set) %>% 
  nest() %>% 
  mutate(lims = map(data, get_performance)) %>% 
  select(set, lims) %>% 
  unnest( cols = c(lims)) %>% 
  mutate(Ilength = Imax - Imin)

# add back species and family

add <- complete %>% 
  select(set, sp, family, lifestage) %>% 
  distinct()


nestpre <- merge(nestpre, add)

```


Preliminary plots (model validation still pending!!!)
```{r}

# Boxplots by lifestage
toadd <- select(inter, set, host, lat, lon, quality)
plotable <- inner_join(nestpre, toadd, by = "set")
names(plotable)
plotable$lifestage <- factor(plotable$lifestage)
ggplot(filter(plotable, set != 748, set != 1102), aes(x = lifestage, y = Imax))+
  geom_boxplot()+
  theme_cowplot()+
  xlab("Life stage")+
  ylab("Interval Max limit (C)")+
  ylim(0,40)

```
Figure 1. Higher limit of the favorable interval by lifestage.
```{r}
# Exclude sets with concave/ Incomplete survival curves
plotable %>% 
  filter(Imax >= 40)
Plotss <- v_nested %>%
  select(set, plot_dr, plot_s)
set748 <- Plotss %>% filter(set == 748)
set748[[2]]# dr plot
set748[[3]] # survival plot
set1102 <- Plotss %>% filter(set == 1102)
set1102[[2]]
set1102[[3]]

```
Discard sets 748 & 1102


```{r}
ggplot(filter(plotable, set != 1102, set != 748), aes(x = lifestage, y = Imin))+
  geom_boxplot()+
  theme_cowplot()+
  xlab("Life stage")+
  ylab("Interval Min limit (C)")+
  ylim(0,40)

```
Figure 2. Lower limit of the favorable interval by lifestage.

```{r}
ggplot(filter(plotable, set != 1102, set != 748), aes(x = lifestage, y = Ilength))+
  geom_boxplot()+
  #geom_jitter(data = plotable, mapping = aes(x = lifestage, y = Imax, col = sp))+
  theme_cowplot()+
  xlab("Life stage")+
  ylab("Favorable interval lenght (C)")+
  ylim(0,30)

```
Figure 4. Lenght of the favorable interval by life stage 

#

```{r}

library(lme4)

unique(plotable$set)
fit <- lmer(Ilength~ lifestage + (1|sp), data = plotable)
hist(plotable$Ilength)
shapiro.test(plotable$Ilength)
summary(fit)

hist(plotable$Ilength)
fit2 <-glm(Ilength~ lifestage, data = plotable, family = Gamma ) 
summary(fit2)




```


```{r}
# Scatterplot by latitude
ggplot(filter(plotable, set != 1102, set != 748), aes(x = abs(lat), y = Topt, col = lifestage))+
  geom_point()+
 # geom_smooth()+
 
  facet_grid(.~lifestage)

spmean <- plotable %>% 
  group_by(sp, family, lifestage) %>% 
  summarise_if(is.numeric, mean, na.rm = T)

```


```{r}
# Scatterplot by latitude
ggplot(filter(plotable, set != 1102, set != 748), aes(x = abs(lat), y = Imin, col = lifestage))+
  geom_point()+
 # geom_smooth()+
 
  facet_grid(.~lifestage)

```



```{r}
# Scatterplot by latitude
ggplot(spmean, aes(x = abs(lat), y = Ilength))+
  geom_point()+
 # geom_smooth()+
 
  facet_grid(.~lifestage)

```


```{r}
# Scatterplot by latitude
ggplot(filter(plotable, set != 1102, set != 748), aes(x = abs(lat), y = Imax, col = sp))+
  geom_point()+
 # geom_smooth()+
 theme(legend.position = "none")+
  facet_grid(.~lifestage)

```

```{r}
# Scatterplot by latitude
ggplot(filter(plotable, set != 1102, set != 748), aes(x = abs(lat), y = Ilength, col = lifestage))+
  geom_point()+
 # geom_smooth()+
 
  facet_grid(.~lifestage)

```

```{r}
# Scatterplot by latitude
ggplot(filter(plotable, set != 1102, set != 748), aes(x = abs(lat), y = Maxdr, col = lifestage))+
  geom_point()+
 # geom_smooth()+
 
  facet_grid(.~lifestage)

```

