---
title: "Development rate"
output: html_notebook
---
## Sections
1. Import data
2. Select sets with at least 4 data points and a complete curve (rise, peak, fall) for egg, larva, pupa, and egg to adult.
3. Consolidate redundant sets
4. Obtain predicted dr in the range from 0 to 40 C using linear interpolation
5. Calculate maximum development rate and Dopt (temperature that allows for it)
6. Calculate D-interval (min, max, length)
7. Visualization
8. Analyses

We used r (tydiverse) for data wrangling, (ggplot2) for data visualization, (psych) for obtaining geometric mean
```{r}
# load packages
library(tidyverse)
library(readxl)
library(cowplot)
library(psych) # to get geometric mean
```

Import data 
```{r}
# Mariana's desktop
Data <- read_xlsx("/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/PhysiologyDatabaseVersion5.xlsx", 
                  sheet = "WorkingTable", na = c("NA", ""))
```

```{r}
# Mariana's laptop
Data <- read_xlsx("/Users/mar/Desktop/Projects/ThermalPerformance/PhysiologyDatabaseVersion5.xlsx", 
                  sheet = "WorkingTable", na = c("NA", "")) 

```

```{r}
# Anna's Laptop
Data <- read_xlsx("~/Desktop/PhysiologyDatabaseVersion5.xlsx", 
                  sheet = "WorkingTable", na = c("NA", "")) 

```


```{r}
# Discard parasitoids and change character to factor 
unique(Data$status)
Ana <- Data[Data$status != "parasitoid",]
rm(Data)

# convert character to factor
Ana <- Ana %>%
  mutate_if(is.character, factor)
```

# 2. Select sets 
To be included in development rate analyses, sets should have at least 4 temperature treatments reporting development time and form a complete curve (rise, peak, fall). We restricted analyses to four life stages: egg, larva, pupa and egg to adult.

```{r}
# Function to determine whether a set is:
# "Complete" (has rise, peak, fall)
# "Incomplete" 
# "No data" (no development rate data available)
Ana$set2 <- Ana$set
is_complete <- function(dat){
  dat <- dat[!is.na(dat$dr) & !is.na(dat$temp),] # Remove treatments with no dr data
 if(nrow(dat) > 3 ){
  
   y_max <- max(dat$dr, na.rm = T) # get maximum value for dr
  x_max <- dat[dat$dr == y_max, "temp"][[1]][1]# get temperature that maximizes dr 
  
  species <- unique(dat$sp)
  lifestage <- unique(dat$lifestage)
  seti <- unique(dat$set2)
  print(paste("set:",seti))
  #print(species)
  #print(lifestage)
  #print(x_max)
  #print(y_max)
  cold <- filter(dat, temp < x_max)  
  hot <- filter(dat, temp > x_max)  
  # if(sum(cold$dr >=0) && sum(hot$dr >=0)){
  #     print("Complete")
  #   } else {(print("Incomplete"))}
  result = nrow(cold[!is.na(cold$dr),]) && nrow(hot[!is.na(hot$dr),])
  if(result){
    print("Complete")
  } else {
    print("Incomplete")
  }
 } else{
   print("Less than 4 data points")
 }
  
  
}


```


```{r}
# Apply is_complete() to all sets
SetAssessment <- Ana %>% 
  group_by(set) %>% 
  nest() %>% 
  mutate(curve = map(data,is_complete)) %>% 
  select(set, curve) %>% 
  unnest(cols = curve) 

```

From a total of `r length(unique(Ana$set))`, only  `r nrow(filter(SetAssessment, curve == "Complete"))` had a complete curve for development rate.
```{r}
# Select sets of life stages of interest: egg, larva, pupa, egg to emergence
complete_sets <- SetAssessment %>% 
  filter(curve == "Complete") 
list_complete <- (unique(complete_sets$set))
life_stages <- c("egg", "larva", "pupa", "eggtoemergence")
Ana_dr <- Ana %>% 
  filter(set %in% list_complete) %>% 
  filter(lifestage %in% life_stages)
egg_dr <- Ana_dr %>%
  filter(lifestage == "egg")
larva_dr <- Ana_dr %>%
  filter(lifestage == "larva")
pupa_dr <- Ana_dr %>%
  filter(lifestage == "pupa")
eggtoemergence_dr <- Ana_dr %>%
  filter(lifestage == "eggtoemergence")

Ana_dr$lifestage <- factor(Ana_dr$lifestage, levels = c("egg", "larva", "pupa", "eggtoemergence"))
```
There are `r length(unique(Ana_dr$set))` sets of the relevant lifestages (`r length(unique(Ana_dr$sp))` species). Of those, `r length(unique(egg_dr$set))` correspond to egg (N = `r length(unique(egg_dr$sp))` species), `r length(unique(larva_dr$set))` to larva (N = `r length(unique(larva_dr$sp))` species), `r length(unique(pupa_dr$set))`to pupa (N = `r length(unique(pupa_dr$sp))` species) and `r length(unique(eggtoemergence_dr$set))` to eggtoadult (N = `r length(unique(eggtoemergence_dr$sp))`)

#3. Consolidate redundant sets
Redundant sets occur beacuse researchers varied experimental conditions such as host plant, humidity and daylength, or because data were reported for males and females separately. We considered multiple measurements reported in the same publication that were obtained at the same temperature, of the same ontogenetic stage of individuals of the same population (sp and locality) to be redundant, and we obtained the geometric mean of all treatments (host, sex, daylength, humidity,etc) at the same temperature for analyses. 
Consolidated sets kept the number (unique set identifier) of the first set they include.


```{r}
# make a new column that unifies redundant sets
Ana_dr <- Ana_dr %>% 
  mutate(redundancy_id = as.numeric(factor(paste(sp, locality, lifestage, id))))

#length(unique(Ana_dr$set)) # there were 287 sets
#length(unique(Ana_dr$redundancy_id)) # there were 216 non-redundant sets

# Function to get the geometric mean of development rate (dr_c), and mean (dr2)
mean_redundant <- function(dat){
newset <- dat$set [[1]]
dat %>%
  select(temp, sp, dr) %>% 
  group_by(temp) %>% 
  summarise(dr_c = geometric.mean(dr),
            dr2 = mean(dr), 
            newset = newset,
            n_sets = length(dr)) 
}

# Apply function to Ana_dr

Ana_dr_consolidated <- Ana_dr %>% 
  group_by(redundancy_id) %>% 
  nest() %>% 
  mutate(consolidated = map(data, mean_redundant)) %>% 
  select(redundancy_id, consolidated) %>% 
  unnest(col = consolidated)# Function

# re-name variables. Note that consolidated sets summarize multiple sets and got assigned the number (set id value) of the first set.
names(Ana_dr_consolidated) <- c("redundancy_id", "temp", "dr", "dr2", "set", "n_sets")
```



# 4. Obtain predicted dr in the range from 0 to 40C using linear interpolation. 

```{r}
# 1- Make interpolation function:
interpolate <- function(dat) {
  out <- seq(from = 0, to = 40, by = 0.1)
  as.data.frame(approx(x = dat$temp, y = dat$dr, xout = out, method = "linear", rule = 1))
  
}
# 2. Apply interpolation function to all sets
Predicted_Ana_dr <-  Ana_dr_consolidated %>% 
  group_by(set) %>% 
  nest() %>% 
  mutate(predicted = map(data,interpolate)) %>% 
  select(set, predicted) %>% 
  unnest(cols = predicted)
names(Predicted_Ana_dr) <- c("set", "temp", "dr")
```



# 5. Obtain Ropt (rate at Topt) and minTopt, maxTopt
```{r}
# function to get optimum

get_Dopt <- function(dat){
  dat <- dat[!is.na(dat$dr) & !is.na(dat$temp),] # Remove treatments with no dr data
   y_max <- max(dat$dr, na.rm = T) # get maximum value for dr
  x_max_list <- dat[dat$dr == y_max, "temp"][[1]]# get temperature that maximizes dr   
  if(length(x_max_list) > 1){
    Dopt <- tibble(DTopt = x_max_list,
                   Dropt = y_max, 
                   Opt = "optima")
  } else{
    Dopt <- tibble(DTopt = x_max_list,
                   Dropt = y_max, 
                   Opt = "optimum")
  }
} 


# apply function to all sets
Dopts <- Predicted_Ana_dr %>% 
  group_by(set) %>% 
  nest() %>% 
  mutate(Dopt = map(data,get_Dopt)) %>% 
  select(set, Dopt)

# get a table with only one optimum temperature
singles <- Dopts %>% 
  unnest(cols = c(Dopt)) %>% 
  filter(Opt == "optimum")


# get a table with sets with multiple optima
multiple <- Dopts %>% 
  unnest(cols = c(Dopt)) %>% 
  filter(Opt == "optima")

# obtain minimum and maximum Dopt
```

There were `r length(unique(singles$set))` sets with a single optimum temperature and `r length(unique(multiple$set))` sets for which several temperatures allowed for development at optimum rates. I extracted both the minimum and maximum temperatures that allow for optimum development rate. Patterns are very simmilar.For analyses we kept the maximum temperature that allows for optimum development rate.

```{r}
# Function to create a table with : set, Dropt, maxTopt, minTopt from sets with multiple optima
get_2_optima <- function(dat){
  dr <- max(dat$Dropt)
  maxTDopt <- max(dat$DTopt, na.rm = T)
  minTDopt <- min(dat$DTopt, na.rm = T)
  outpu <- tibble(dr = dr, 
                  maxTDopt = maxTDopt,
                  minTDopt = minTDopt)
}


```


```{r}
#Apply get_2_optima() to sets with multiple optimum temperatures
two_optima_table <- multiple %>% 
  group_by(set) %>% 
  nest() %>% 
  mutate(Opt = map(data,get_2_optima)) %>% 
  select(set, Opt) %>% 
  unnest( cols = c(Opt))
```

```{r}
# minTopt (Topt for singles, minTopt for multiple)
# 1- Get values from "singles" table:
min_Topt_table1 <- singles %>% 
  mutate(minTDopt = DTopt) %>% # minTDopt = DTopt in sets with a single optimum
  select(set, Dropt, minTDopt) 

min_Topt_table2 <- two_optima_table %>% 
   mutate(Dropt = dr) %>%  # so columns have the same name in both tables
  select(set, Dropt, minTDopt)  

min_Topt_table3 <- rbind(min_Topt_table1, min_Topt_table2)
# Add columns for lifestage, lat, lon, population, sp and family for analyses

add_info <- Ana_dr %>% 
  select(set, sp, family, lifestage, lat, lon, locality, quality) %>% 
  distinct()

min_Topt_table <- inner_join(min_Topt_table3, add_info, by = "set")
rm(min_Topt_table1, min_Topt_table2, min_Topt_table3)
```

```{r}
# maxTopt (Topt for singles, maxTopt for multiple)
# 1- Get values from "singles" table:
max_Topt_table1 <- singles %>% 
  mutate(maxTDopt = DTopt) %>% # minTDopt = DTopt in sets with a single optimum
  select(set, Dropt, maxTDopt) 

max_Topt_table2 <- two_optima_table %>% 
   mutate(Dropt = dr) %>%  # so columns have the same name in both tables
  select(set, Dropt, maxTDopt)  

max_Topt_table3 <- rbind(max_Topt_table1, max_Topt_table2)
Topt_table <- inner_join(max_Topt_table3, add_info, by = "set")
rm(max_Topt_table1, max_Topt_table2, max_Topt_table3)

```

#6. Calculate interval (min 50% dr, max 50% dr, length)

```{r}


get_interval <- function(dat){
  dat <- dat[!is.na(dat$dr) & !is.na(dat$temp),] # Remove treatments with no dr data
  y_max <- max(dat$dr) # get maximum value for dr
  x_max <- dat[dat$dr == y_max, "temp"][[1]][[1]]# get temperature that maximizes dr
  cold <- filter(dat, temp < x_max)  
  hot <- filter(dat, temp > x_max)  
  limit <- y_max/2
  if (min(cold$dr) > limit){
    print("cold incomplete")
    y_mincold <- NA
    x_min_cold <- NA
    grade_cold <- "incomplete"
  } else{
    y_mincold <- min(filter(cold, dr >= y_max/2)[["dr"]])
    x_min_cold <- min(filter(cold, dr >= y_max/2)[["temp"]])
    grade_cold <- "complete"
  }
  
  if(min(hot$dr) > limit){
    print("heat incomplete")
    y_minhot <- NA
    x_min_hot <- NA 
    grade_hot <- "incomplete"
  } else {
    y_minhot <- min(filter(hot, dr >= y_max/2)[["dr"]])
    x_min_hot <- min(filter(hot, dr >= y_max/2)[["temp"]])
    grade_hot <- "complete"
  }
  
  
  output <- tibble(Imin = x_min_cold, 
                   Imax = x_min_hot, 
                   Ilength = Imax - Imin, 
                   gradecold = grade_cold, 
                   gradehot = grade_hot)
  
  
}




```


```{r}
# Apply get interval to everything
# apply function to all sets
Interval <- Predicted_Ana_dr %>% 
  group_by(set) %>% 
  nest() %>% 
  mutate(interval = map(data,get_interval)) %>% 
  select(set, interval) %>% 
  unnest(cols = interval)
Interval_table1 <- inner_join(Interval, add_info, by = "set")
Interval_table1$lifestage <- factor(Interval_table1$lifestage, levels = c("egg", "larva", "pupa", "eggtoemergence"))
Interval_table_cold <- filter(Interval_table1, gradecold == "complete")
Interval_table_hot <- filter(Interval_table1, gradehot == "complete")
Interval_table_both <- filter(Interval_table1, gradecold == "complete", gradehot == "complete")
rm(Interval_table1)
```


```{r}
# clean-up so only relevant tables for analysis remain in workspace
#rm(add_info) # columns with sp, lifestage, and lat lon
#rm(Dopts) #list of set and otimum temperatures
#rm(Interval) # table missing columns relevant for analysis
#rm(multiple)

#rm(singles)
#rm(two_optima_table)
#rm(complete_sets)
```

Description of tables relevant for analysis
Ana: original table
Ana_dr: table with complete sets of relevant lifestages
Interval_table_cold: Imin
Interval_table_hot: Imax 
Interval_table_both: Ilength
Topt_table: table of maximum temperatures that allow for development rate at maximum efficiency
min_Topt_table: table of minimum temperatures that allow for development rate at maximum efficiency. There are two tables because in some cases there are several optima instead of one. plots for these tables are almost the same, so we should choose 1 (Topt_table) to focus but analyze both and report if results were the same or different

# 7. Visualization
## Dopt
Temperature that allows for development rate at maximum efficiency

```{r}
ggplot(Topt_table, aes(x = lifestage, y = maxTDopt))+
  geom_boxplot()+
  theme_cowplot()+
  ylab("Dopt (C)")+
  ylab("TDopt (C)")+
  xlab("Life stage")+
  #geom_jitter(mapping = aes(x = lifestage, y = maxTDopt, col = family))+
  #geom_text(mapping = aes(x = lifestage, y = maxTDopt, label = sp))+
  ylim(10,40)
  
```


Optimum development rates of all life stages occurred between `r min(Topt_table$maxTDopt, na.rm = T)` and `r max(Topt_table$maxTDopt, na.rm = T)`C. The mean tempeature allowing for optimum development was `r mean(Topt_table$maxTDopt, na.rm = T)` sd = `r sd(Topt_table$maxTDopt, na.rm = T)`. N = `r length(unique(Topt_table$set))` consolidated sets of `r length(unique(Topt_table$sp))` species.


```{r}
# I'm exluding points where locality was not reported ("inferred" value in field quiality)

ggplot(filter(Topt_table, quality != "inferred"), aes(x = abs(lat), y = maxTDopt, col = family))+
  geom_point()+
  theme_cowplot()+
  ylab("TDopt")+
  xlab("Absolute latitude")+
  #geom_text(data = filter(max_Topt_table, sp == "Pteroma pendula"), mapping = aes(x = abs(lat), y = maxTDopt, label = sp))+
  facet_wrap(.~lifestage)

#filter(max_Topt_table, abs(lat) < 10, sp == "Pteroma pendula")
```
Point with TDopt = 35 at 3.2 latitude corresponds to Pteroma pendula (Psychidae), a bag worm, pest of oil palm.
Same patters if instead of using the max value for TDopt we use the minimum for those 26 sets that had multiple temperatures allowing for optimum development rate.
```{r}
# ggplot(min_Topt_table, aes(x = lifestage, y = minTDopt))+
# geom_boxplot()+
# #geom_jitter(mapping = aes(x = lifestage, y = minTDopt, col = family))
#   theme_cowplot()+
#   ylab("TDopt (min)")+
#   #geom_jitter(mapping = aes(x = lifestage, y = maxTDopt, col = family))+
#   ylim(10,40)
```

```{r}
# ggplot(min_Topt_table, aes(x = abs(lat), y = minTDopt), col = "quality")+
#   geom_point()+
#   theme_cowplot()+
#   ylab("TDopt (min)")+
#   #geom_jitter(mapping = aes(x = lifestage, y = maxTDopt, col = family))+
#    facet_wrap(.~lifestage)
```


```{r}
# ggplot(Topt_table, aes(x = lifestage, y = maxTDopt, col = family, label = set))+
#   #geom_jitter()+
#   geom_text()

```
This is to check outliers (sets 1005, 175).1005 is incomplete, 175 is fine.
```{r}
ggplot(filter(Ana_dr, set == 1005), aes(x = temp, y = dr))+
  geom_point()

```

```{r}
ggplot(Interval_table_both, aes(x = lifestage, y = Ilength))+
  geom_boxplot()+
  theme_cowplot()+
  ylab("D-interval length")+
  #geom_jitter(mapping = aes(x = lifestage, y = maxTDopt, col = family))+
  ylim(0,40)

```

Lenght of the temperature interval that allows for development rates at >=50% efficiency.  


```{r}
# Summary statistics
summary <-Interval_table_both %>% 
  group_by(lifestage) %>% 
  summarise(median = median(Ilength, na.rm = T), q0.25 =quantile(Ilength,probs =0.25, na.rm = T), q0.75 =quantile(Ilength,probs =0.75, na.rm = T), N = n())#, c(0.25, 0.5, 0.75), type = 1,na.rm = T))

summary
```
 
 
 Median length of the favorable interval (allows for development to occur at 50% efficiency) was `r summary[summary$lifestage == "egg", "median"][[1]] `C for eggs (N = `r summary[summary$lifestage == "egg", "N"][[1]] `), `r summary[summary$lifestage == "larva", "median"][[1]] `C for larvae (N = `r summary[summary$lifestage == "egg", "median"][[1]] `), `r summary[summary$lifestage == "pupa", "N"][[1]] `C for pupae (N = `r summary[summary$lifestage == "egg", "N"][[1]] `), and `r summary[summary$lifestage == "eggtoemergence", "median"][[1]] ` for the whole lifecycle (egg to emergence, N = `r summary[summary$lifestage == "egg", "N"][[1]] `)
 
```{r}

ggplot(filter(Interval_table_both, quality != "inferred"), aes(x = abs(lat), y = Ilength, col = family))+
  geom_point()+
  theme_cowplot()+
  ylab("D-interval length")+
  #geom_jitter(mapping = aes(x = lifestage, y = maxTDopt, col = family))+
  #ylim(10,40)+
  facet_wrap(.~lifestage)
```

```{r}
ggplot(Interval_table_cold, aes(x = lifestage, y = Imin))+
  geom_boxplot()+
  theme_cowplot()+
  ylab("D-interval minimum (C)")+
  #geom_jitter(mapping = aes(x = lifestage, y = maxTDopt, col = family))+
  ylim(0,20)
```
Minimum temperature that allows for development at >= 50% efficiency

```{r}
ggplot(filter(Interval_table_cold, quality != "inferred"), aes(x = abs(lat), y = Imin), col = "quality")+
  geom_point()+
  theme_cowplot()+
  ylab("D-interval minimum")+
  #geom_jitter(mapping = aes(x = lifestage, y = maxTDopt, col = family))+
  #ylim(10,40)+
  facet_wrap(.~lifestage)
```





```{r}
ggplot(Interval_table_hot, aes(x = lifestage, y = Imax))+
 geom_boxplot()+
  theme_cowplot()+
  ylab("D-interval max")+
  #geom_jitter(mapping = aes(x = lifestage, y = maxTDopt, col = family))+
  ylim(10,40)
```
Maximum temperature that allows for development at >= 50% efficiency.

```{r}
ggplot(filter(Interval_table_hot, quality != "inferred"), aes(x = abs(lat), y = Imax), col = "quality")+
  geom_point()+
  theme_cowplot()+
  ylab("D-interval maximum")+
  #geom_jitter(mapping = aes(x = lifestage, y = maxTDopt, col = family))+
  #ylim(10,40)+
  facet_wrap(.~lifestage)
```



7. Perform analyses

Dopt. For analysis I discarded "eggtoemergence" because it integrates patterns of eggs, larvae and pupae. I also discarded sets in which the locality was "inferred" instead of reported by authors.

```{r}
library(lme4)
library(car)
hist(Topt_table$maxTDopt)
shapiro.test((Topt_table$maxTDopt))
ToptAnalysis <- filter(Topt_table, lifestage != "eggtoemergence", quality!= "inferred")

ToptAnalysis$absLat <- abs(ToptAnalysis$lat)

fit1 <- lmer(maxTDopt ~ lifestage + absLat + (1|sp), data = ToptAnalysis) #
summary(fit1)
Anova(fit1) # 
plot(fit1) #

```

Optimum development rates of all life stages occurred between `r min(ToptAnalysis$maxTDopt, na.rm = T)` and `r max(ToptAnalysis$maxTDopt, na.rm = T)`C. The mean tempeature allowing for optimum development was `r mean(ToptAnalysis$maxTDopt, na.rm = T)` sd = `r sd(ToptAnalysis$maxTDopt, na.rm = T)`. N = `r length(unique(ToptAnalysis$set))` consolidated sets of `r length(unique(ToptAnalysis$sp))` species.

Absoulte latitude, but not lifestage explained variation in TDopt.


```{r}

library(effects)
library(ggeffects)
ToptAnalysis$group <- ToptAnalysis$lifestage
pred <- ggpredict(fit1, terms = c("absLat","lifestage"))
pred2 <- Effect(c("absLat","lifestage"),fit1)
#plot(pred2)
ggplot(pred, aes(x = x, y = predicted, colour = group, fill = group)) +
  geom_line()+
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, col = group), alpha = 0.1)+
  geom_point(data = ToptAnalysis, mapping = aes(x = abs(lat), y = maxTDopt, col =group, alpha = 0.1))+
  scale_color_viridis_d()+
  scale_fill_viridis_d()+
  ylab("Dopt")+
  xlab("Absolute latitude")+
  #ylim(15, 40)+
  theme_cowplot()#+
  #facet_grid(.~group)




```
Observed (points) and predicted (line, shaded interval) variation of Dopt (temperature that allows for optimum development rate) according to latitude and lifestage using model:  lmer(maxTDopt ~ lifestage + absLat + (1|sp), data = ToptAnalysis)



```{r}
#Imin
IntervalAnalysis_cold <- filter(Interval_table_cold, lifestage != "eggtoemergence", quality!= "inferred")
IntervalAnalysis_cold$absLat <- abs(IntervalAnalysis_cold$lat)


fit2 <- lmer(Imin ~ lifestage + absLat + (1|sp) + (1|family), data = IntervalAnalysis_cold) # 
summary(fit2)
Anova(fit2) # 
plot(fit2) #
```

Absolute latitude, but not life stage explained variation in Imin



```{r}

library(effects)
library(ggeffects)
pred <- ggpredict(fit2, terms = c("absLat","lifestage"))
IntervalAnalysis_cold$group <- IntervalAnalysis_cold$lifestage

ggplot(pred, aes(x = x, y = predicted, colour = group, fill = group)) +
  geom_line()+
   geom_ribbon(aes(ymin = conf.low, ymax = conf.high, col = group), alpha = 0.1)+
  scale_color_viridis_d()+
  scale_fill_viridis_d()+
  ylab("D-interval lower limit")+
  xlab("Absolute latitude")+
 # ylim(5, 20)+
  geom_point(data = IntervalAnalysis_cold, mapping = aes(x = abs(lat), y = Imin, col =group ), alpha = 0.4)+
  theme_cowplot()




```

Observed (points) and predicted (line, shaded interval) variation of the minimum temperature that allows for development at >= 50% efficiency according to latitude and lifestage using model:  lmer(Imin ~ lifestage + absLat + (1|sp) + (1|family), data = IntervalAnalysis).
The lower limit of D- interval (minimum temperature that allowed for development at 50% efficiency or higher) ranged from `r min(IntervalAnalysis_cold$Imin)`ºC to `r max(IntervalAnalysis_cold$Imin)`ºC, and the mean was `r mean(IntervalAnalysis_cold$Imin)` ± `r sd(IntervalAnalysis_cold$Imin)` (mean ± SD, N = `r length(unique(IntervalAnalysis_cold$set))` sets from `r length(unique(IntervalAnalysis_cold$sp))` species).


```{r}
# Imax
IntervalAnalysis_hot <- filter(Interval_table_hot, lifestage != "eggtoemergence", quality!= "inferred")
IntervalAnalysis_hot$absLat <- abs(IntervalAnalysis_hot$lat)
hist(Interval_table_hot$Imax)
shapiro.test((Interval_table_hot$Imax))
fit3 <- lmer(Imax ~ lifestage + absLat + (1|sp) + (1|family), data = IntervalAnalysis_hot) # 
summary(fit3)
Anova(fit3) # 
plot(fit3) #
```

The upper limit of D- interval (maximum temperature that allowed for development at 50% efficiency or higher) ranged from `r min(IntervalAnalysis_hot$Imax, na.rm = T)`ºC to `r max(IntervalAnalysis_hot$Imax, na.rm = T)`ºC, and the mean was `r mean(IntervalAnalysis_hot$Imax, na.rm = T)` ± `r sd(IntervalAnalysis_hot$Imax, na.rm = T)` (mean ± SD, N = `r length(unique(IntervalAnalysis_hot$set))` sets from `r length(unique(IntervalAnalysis_hot$sp))` species).The maximum temperature did not vary across latitudes, or lifestages.

```{r}


ggplot(IntervalAnalysis_hot, aes(x = absLat, y = Imax, col = lifestage))+
 geom_point()+
  theme_cowplot()+
  ylab("D-interval upper limit")+
  scale_color_viridis_d(alpha = 0.5)+
  xlab("Absolute latitude")+
  #geom_jitter(mapping = aes(x = lifestage, y = maxTDopt, col = family))+
  ylim(10,40)
```

Observed (points) variation of the maximum temperature that allows for development at >= 50% efficiency. Neither absolute latitude nor lifestage explained the observed variation.
This temperature is where D-interval starts

```{r}
#Ilength
IntervalAnalysis_both <- filter(Interval_table_both, lifestage != "eggtoemergence", quality!= "inferred")
IntervalAnalysis_both$absLat <- abs(IntervalAnalysis_both$lat)
hist(IntervalAnalysis$Ilength)
fit4 <- lmer(Ilength ~ lifestage + absLat + (1|sp) + (1|family), data = IntervalAnalysis_both) # 
summary(fit4)
Anova(fit4) # 
plot(fit4) #
```

```{r}

ggplot(IntervalAnalysis_both, aes(x = absLat, y = Ilength, col = lifestage))+
 geom_point()+
  theme_cowplot()+
  ylab("D-interval length")+
  scale_color_viridis_d(alpha = 0.5)+
  xlab("Absolute latitude")+
  #geom_jitter(mapping = aes(x = lifestage, y = maxTDopt, col = family))+
  ylim(0,20)

```

Length of the interval of temperatures that allowed for development at >=50% efficiency. interval length varied from  `r min(IntervalAnalysis_both$Ilength, na.rm = T)`  to max `r max(IntervalAnalysis_both$Ilength, na.rm = T)`, mean = `r mean(IntervalAnalysis_both$Ilength, na.rm = T)`, SD = `r sd(IntervalAnalysis_both$Ilength, na.rm = T)`, N = `r length(unique(IntervalAnalysis_both$set))` sets from `r length(unique(IntervalAnalysis_both$sp))` species).


```{r}
Intervar <- select(IntervalAnalysis_both, set, Ilength, Imin, Imax)

allvalues <- full_join(Intervar,ToptAnalysis, by = "set")
ggplot(allvalues, aes(x = absLat, y = Dropt, col = lifestage))+
 geom_point()+
  theme_cowplot()+
  ylab("Maximum development rate")+
  scale_color_viridis_d(alpha = 0.5)+
  xlab("Absolute latitude")
  #geom_jitter(mapping = aes(x = lifestage, y = maxTDopt, col = family))+
  

```


