---
title: "Calculate development rate variables"
output: html_notebook
---
## Contents
1. Import data
2. Set selection
3. Redundant set consolidation
4. Linear interpolation
5. Obtain Dopt, minTopt, maxTopt, and development rate at Dopt

We used r (tydiverse) for data wrangling, (ggplot2) for data visualization, (psych) for obtaining geometric mean, (readxl) to import excel file, (cowplot) to format ggplots
```{r}
# load packages
library(tidyverse)
library(readxl)
library(cowplot)
library(psych) # to get geometric mean
```

Import data 
```{r}
# Mariana's desktop
Data <- read_xlsx("/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/PhysiologyDatabaseVersion5.xlsx", 
                  sheet = "WorkingTable", na = c("NA", ""))
```

```{r}
# Mariana's laptop
Data <- read_xlsx("/Users/mar/Desktop/Projects/ThermalPerformance/PhysiologyDatabaseVersion5.xlsx", 
                  sheet = "WorkingTable", na = c("NA", "")) 

```

```{r}
# Anna's Laptop
Data <- read_xlsx("~/Desktop/PhysiologyDatabaseVersion5.xlsx", 
                  sheet = "WorkingTable", na = c("NA", "")) 

```


```{r}
# Discard parasitoids and change character to factor 
unique(Data$status)
Ana <- Data[Data$status != "parasitoid",]
rm(Data)

# convert character to factor
Ana <- Ana %>%
  mutate_if(is.character, factor)
```

# 2. Select sets 
To be included in development rate analyses, sets should have at least 4 temperature treatments reporting development time, and form a complete curve (rise, peak, fall). We restricted analyses to four life stages: egg, larva, pupa, and egg to adult.

```{r}
# Function to determine whether a set is:
# "Complete" (has rise, peak, fall)
# "Incomplete" (it's a partial curve)
# "No data" (no development rate data available)
Ana$set2 <- Ana$set
is_complete <- function(dat){
  dat <- dat[!is.na(dat$dr) & !is.na(dat$temp),] # Remove treatments with no dr data
 if(nrow(dat) > 3 ){
  
   y_max <- max(dat$dr, na.rm = T) # get maximum value for s_dr
  x_max <- dat[dat$dr == y_max, "temp"][[1]][1]# get temperature that maximizes dr 
  
  species <- unique(dat$sp)
  lifestage <- unique(dat$lifestage)
  seti <- unique(dat$set2)
  print(paste("set:",seti))
  #print(species)
  #print(lifestage)
  #print(x_max)
  #print(y_max)
  cold <- filter(dat, temp < x_max)  
  hot <- filter(dat, temp > x_max)  
  # if(sum(cold$dr >=0) && sum(hot$dr >=0)){
  #     print("Complete")
  #   } else {(print("Incomplete"))}
  result = nrow(cold[!is.na(cold$dr),]) && nrow(hot[!is.na(hot$dr),])
  if(result){
    print("Complete")
  } else {
    print("Incomplete")
  }
 } else{
   print("Less than 4 data points")
 }
  
  
}


```


```{r}
# Apply is_complete() to all sets
SetAssessment <- Ana %>% 
  group_by(set) %>% 
  nest() %>% 
  mutate(curve = map(data,is_complete)) %>% 
  select(set, curve) %>% 
  unnest(cols = curve) 
unique(SetAssessment$curve)
```

From a total of `r length(unique(Ana$set))`, only  `r nrow(filter(SetAssessment, curve == "Complete"))` had a complete curve for development rate. There were `r nrow(filter(SetAssessment, curve == "Incomplete"))` incomplete sets and `r nrow(filter(SetAssessment, curve == "Less than 4 data points"))` with less than four data points.

```{r}
# Select sets of life stages of interest: egg, larva, pupa, egg to emergence
complete_sets <- SetAssessment %>% 
  filter(curve == "Complete") 
list_complete <- (unique(complete_sets$set))
life_stages <- c("egg", "larva", "pupa", "eggtoemergence")
Ana_dr <- Ana %>% 
  filter(set %in% list_complete) %>% 
  filter(lifestage %in% life_stages)
egg_dr <- Ana_dr %>%
  filter(lifestage == "egg")
larva_dr <- Ana_dr %>%
  filter(lifestage == "larva")
pupa_dr <- Ana_dr %>%
  filter(lifestage == "pupa")
eggtoemergence_dr <- Ana_dr %>%
  filter(lifestage == "eggtoemergence")

Ana_dr$lifestage <- factor(Ana_dr$lifestage, levels = c("egg", "larva", "pupa", "eggtoemergence"))
```
There are `r length(unique(Ana_dr$set))` sets of the relevant lifestages (`r length(unique(Ana_dr$sp))` species). Of those, `r length(unique(egg_dr$set))` correspond to egg (N = `r length(unique(egg_dr$sp))` species), `r length(unique(larva_dr$set))` to larva (N = `r length(unique(larva_dr$sp))` species), `r length(unique(pupa_dr$set))`to pupa (N = `r length(unique(pupa_dr$sp))` species) and `r length(unique(eggtoemergence_dr$set))` to eggtoadult (N = `r length(unique(eggtoemergence_dr$sp))`)

# 3. Consolidate redundant sets
Redundant sets occur beacuse researchers varied experimental conditions such as host plant, humidity and daylength, or because data were reported for males and females separately. We considered multiple measurements reported in the same publication that were obtained at the same temperature, of the same ontogenetic stage of individuals of the same population (sp and locality) to be redundant, and we obtained the geometric mean of all treatments (host, sex, daylength, humidity,etc) at the same temperature for analyses. 
Consolidated sets kept the number (unique set identifier) of the first set they include.


```{r}
# make a new column that unifies redundant sets
Ana_dr <- Ana_dr %>% 
  mutate(redundancy_id = as.numeric(factor(paste(sp, locality, lifestage, id))))

#length(unique(Ana_dr$set)) # there were 287 sets
#length(unique(Ana_dr$redundancy_id)) # there were 216 non-redundant sets

# Function to get the geometric mean of development rate (dr_c), and mean (dr2)
mean_redundant <- function(dat){
newset <- dat$set [[1]]
dat %>%
  select(temp, sp, dr) %>% 
  group_by(temp) %>% 
  summarise(dr_c = geometric.mean(dr),
            dr2 = mean(dr), 
            newset = newset,
            n_sets = length(dr)) 
}

# Apply function to Ana_dr

Ana_dr_consolidated <- Ana_dr %>% 
  group_by(redundancy_id) %>% 
  nest() %>% 
  mutate(consolidated = map(data, mean_redundant)) %>% 
  select(redundancy_id, consolidated) %>% 
  unnest(col = consolidated)# Function

# re-name variables. Note that consolidated sets summarize multiple sets and got assigned the number (set id value) of the first set.
names(Ana_dr_consolidated) <- c("redundancy_id", "temp", "dr", "dr2", "set", "n_sets")
```

Function to scale development rate
```{r}
scale.dr <- function(table){
  table <- table[!is.na(table$dr) & !is.na(table$temp),] # Remove treatments with no performance data
  if(nrow(table) > 1 ){
  maxdr <- max(table$dr, na.rm = T)
  table$s_dr <- table$dr/maxdr
  table
  }
}


```


Apply the function to each set to obtain s_dr
```{r}
Ana_dr_consolidated <- Ana_dr_consolidated %>% 
  group_by(set) %>% 
  nest() %>% 
  mutate(s_dr = map(data, scale.dr)) %>% 
  unnest(cols = s_dr) %>% 
  select(-data)

# ggplot(Ana_dr_consolidated, aes(x = dr, y = s_dr))+
#   geom_point()
```

# 4. Obtain predicted dr in the range from 0 to 40C using linear interpolation. 

```{r}
# 1- Make interpolation function:
interpolate <- function(dat) {
  out <- seq(from = 0, to = 40, by = 0.1)
  as.data.frame(approx(x = dat$temp, y = dat$s_dr, xout = out, method = "linear", rule = 1))
  
}
# 2. Apply interpolation function to all sets
Predicted_Ana_dr <-  Ana_dr_consolidated %>% 
  group_by(set) %>% 
  nest() %>% 
  mutate(predicted = map(data,interpolate)) %>% 
  select(set, predicted) %>% 
  unnest(cols = predicted)
names(Predicted_Ana_dr) <- c("set", "temp", "s_dr")
```



# 5. Obtain Ropt (rate at Dopt) and D-width, maxTopt
```{r}
# function to get optimum

get_Dopt <- function(dat){
  dat <- dat[!is.na(dat$s_dr) & !is.na(dat$temp),] # Remove treatments with no dr data
   y_max <- max(dat$s_dr, na.rm = T) # get maximum value for dr
  x_max_list <- dat[dat$s_dr == y_max, "temp"][[1]]# get temperature that maximizes dr   
  if(length(x_max_list) > 1){
    Dopt <- tibble(Dopt = x_max_list,
                   Dropt = y_max, 
                   Opt = "optima")
  } else{
    Dopt <- tibble(Dopt = x_max_list,
                   Dropt = y_max, 
                   Opt = "optimum")
  }
} 


# apply function to all sets
Dopts <- Predicted_Ana_dr %>% 
  group_by(set) %>% 
  nest() %>% 
  mutate(Dopt = map(data,get_Dopt)) %>% 
  select(set, Dopt)

# get a table with only one optimum temperature
singles <- Dopts %>% 
  unnest(cols = c(Dopt)) %>% 
  filter(Opt == "optimum")


# get a table with sets with multiple optima
multiple <- Dopts %>% 
  unnest(cols = c(Dopt)) %>% 
  filter(Opt == "optima")

# obtain minimum and maximum Dopt
```

There were `r length(unique(singles$set))` sets with a single optimum temperature and `r length(unique(multiple$set))` sets for which several temperatures allowed for development at optimum rates. I extracted both the minimum and maximum temperatures that allow for optimum development rate. Patterns are very simmilar.For analyses we kept the maximum temperature that allows for optimum development rate.

```{r}
# Function to create a table with : set, Dropt, maxTopt, minTopt from sets with multiple optima
get_2_optima <- function(dat){
  dr <- max(dat$Dropt)
  maxTDopt <- max(dat$Dopt, na.rm = T)
  minTDopt <- min(dat$Dopt, na.rm = T)
  outpu <- tibble(dr = dr, 
                  Dopt = maxTDopt,
                  minDopt = minTDopt)
}


```


```{r}
#Apply get_2_optima() to sets with multiple optimum temperatures
two_optima_table <- multiple %>% 
  group_by(set) %>% 
  nest() %>% 
  mutate(Opt = map(data,get_2_optima)) %>% 
  select(set, Opt) %>% 
  unnest( cols = c(Opt))
```

```{r}
# minTopt (Topt for singles, minTopt for multiple)
# 1- Get values from "singles" table:
min_Topt_table1 <- singles %>% 
  mutate(minTDopt = Dopt) %>% # minTDopt = DTopt in sets with a single optimum
  select(set, Dropt, minTDopt) 

min_Topt_table2 <- two_optima_table %>% 
   mutate(Dropt = dr) %>%  # so columns have the same name in both tables
  select(set, Dropt, minDopt)  

min_Topt_table3 <- rbind(min_Topt_table1, min_Topt_table2)
# Add columns for lifestage, lat, lon, population, sp and family for analyses

add_info <- Ana_dr %>% 
  select(set, sp, family, lifestage, lat, lon, locality, quality) %>% 
  distinct()

min_Topt_table <- inner_join(min_Topt_table3, add_info, by = "set")
rm(min_Topt_table1, min_Topt_table2, min_Topt_table3)
```

```{r}
# Dopt was (Dopt for singles, Dopt for multiple)
# 1- Get values from "singles" table:
Dopt_table1 <- singles %>% 
  select(set, Dropt, Dopt) 

Dopt_table2 <- two_optima_table %>% 
   mutate(Dropt = dr) %>%  # so columns have the same name in both tables
  select(set, Dropt, Dopt)  

Dopt_table3 <- rbind(Dopt_table1, Dopt_table2)
Dopt_table <- inner_join(Dopt_table3, add_info, by = "set")
rm(Dopt_table1, Dopt_table2, Dopt_table3)

```

#6. Calculate D-width (min 50% dr, max 50% dr, length)

```{r}


get_interval <- function(dat){
  dat <- dat[!is.na(dat$s_dr) & !is.na(dat$temp),] # Remove treatments with no s_dr data
  y_max <- max(dat$s_dr) # get maximum value for s_dr
  x_max <- dat[dat$s_dr == y_max, "temp"][[1]][[1]]# get temperature that maximizes s_dr
  cold <- filter(dat, temp < x_max)  
  hot <- filter(dat, temp > x_max)  
  limit <- y_max/2
  if (min(cold$s_dr) > limit){
    print("cold incomplete")
    y_mincold <- NA
    x_min_cold <- NA
    grade_cold <- "incomplete"
  } else{
    y_mincold <- min(filter(cold, s_dr >= y_max/2)[["s_dr"]])
    x_min_cold <- min(filter(cold, s_dr >= y_max/2)[["temp"]])
    grade_cold <- "complete"
  }
  
  if(min(hot$s_dr) > limit){
    print("heat incomplete")
    y_minhot <- NA
    x_min_hot <- NA 
    grade_hot <- "incomplete"
  } else {
    y_minhot <- min(filter(hot, s_dr >= y_max/2)[["s_dr"]])
    x_min_hot <- max(filter(hot, s_dr >= y_max/2)[["temp"]])
    grade_hot <- "complete"
  }
  
  
  output <- tibble(Imin = x_min_cold, 
                   Imax = x_min_hot, 
                   Ilength = Imax - Imin, 
                   gradecold = grade_cold, 
                   gradehot = grade_hot)
  
  
}




```


```{r}
# Apply get interval to everything
# apply function to all sets
Interval <- Predicted_Ana_dr %>% 
  group_by(set) %>% 
  nest() %>% 
  mutate(interval = map(data,get_interval)) %>% 
  select(set, interval) %>% 
  unnest(cols = interval)
Interval_table1 <- inner_join(Interval, add_info, by = "set")
Interval_table1$lifestage <- factor(Interval_table1$lifestage, levels = c("egg", "larva", "pupa", "eggtoemergence"))
Interval_table_cold <- filter(Interval_table1, gradecold == "complete")
Interval_table_hot <- filter(Interval_table1, gradehot == "complete")
Interval_table_both <- filter(Interval_table1, gradecold == "complete", gradehot == "complete")
rm(Interval_table1)
```

# 7. get a table with maximum development rate (not standardized)
 
```{r}
MaxDrTable <- Ana_dr_consolidated %>%
  select(set, dr) %>% 
  group_by(set) %>%
   nest() %>%
   mutate(Max_dr = map(data, max)) %>%
   select(set, Max_dr) %>%
   unnest(cols = c(Max_dr)) %>% 
  filter(!is.na(Max_dr))
```
 
 
 



Print tables for analysis
Dopt: Optimum development rate
Interval_table_cold: lower limmit of D-witdh
Interval_table_hot: upper limmit of D-witdh 
Interval_table_both: D-witdh 

```{r}
# 
# write_csv(Dopt_table,"/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/Dopt.csv")
# write_csv(Dopt_table,"/Users/mar/Desktop/Projects/ThermalPerformance/Dopt.csv")
# write_csv(Interval_table_cold,"/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/llD-width.csv")
# write_csv(Interval_table_cold,"/Users/mar/Desktop/Projects/ThermalPerformance/llD-width.csv")
# write_csv(Interval_table_hot,"/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/ulD-width.csv")
# write_csv(Interval_table_hot,"/Users/mar/Desktop/Projects/ThermalPerformance/ulD-width.csv")
# write_csv(Interval_table_both,"/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/D-width.csv")
# write_csv(Interval_table_both,"/Users/mar/Desktop/Projects/ThermalPerformance/D-width.csv")

```

Merge tables

```{r}
names(Dopt_table)
```



```{r}
Dopt_to_merge <- Dopt_table %>% mutate(s_dr = Dropt) %>% select(-Dropt)
Merge1 <- select(Interval_table_cold, set, Dlower = Imin)
head(Merge1)
Merge2 <- select(Interval_table_hot, set, Dupper = Imax)
head(Merge2)
Merge3 <- select(Interval_table_both, set, Dwidth =Ilength)
head(Merge3)
DevelopmentRate1 <- full_join(Dopt_to_merge, Merge1,  by = "set")
DevelopmentRate2 <- full_join(DevelopmentRate1, Merge2,  by = "set")
DevelopmentRate3 <- full_join(DevelopmentRate2, Merge3,  by = "set")
DevelopmentRate <- full_join(DevelopmentRate3, MaxDrTable,  by = "set")
head(DevelopmentRate)
names(DevelopmentRate)

```


```{r}
# write_csv(DevelopmentRate,"/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/Development.csv")
```

