---
title: "Analyze development rate variables"
output: html_notebook
---
## Contents
1. Import data and housekeeping

2. Dlow (Dlower)
2.1  Model selection
2.2  Implement phylogenetic correction

3. Dwidth
3.1  Model selection
3.2  Phylogenetic correction

4. Dhigh (Dupper)
4.1  Model selection
4.2  Implement phylogenetic correction

5. Dopt
5.1  Model selection
5.2  Phylogenetic correction


# 1. Import data and housekeeping
```{r}
 
#library(INLA) # install.packages("INLA", repos=c(getOption("repos"), #INLA="https://inla.r-inla-download.org/R/stable"), dep=TRUE)
library(tidyverse)
library(readxl)
library(cowplot)
library(psych) # to get geometric mean
library(lme4)
library(car)
library("dotwhisker")  # to make coefficient plots
library(sjPlot) #prediction plots
library(sjmisc)
library(broom)
library(effects)
library(ggeffects)
library(Hmisc)
library(phyr)
library(ape)
library(nlme)
library(stringr)
library(phylotools)
library(phytools)
library(rr2)
# sessionInfo()
```


```{r}
# Mariana's desktop

responses <- read_csv("/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/ThermalResponsesTable.csv")
tree <- read.tree("GTRG_Unpartitioned_constrain.raxml.bestTree.tre")

```

```{r}
# Mariana's laptop
#responses <- read_csv("/Users/mar/Desktop/Projects/ThermalPerformance/ThermalResponsesIII.csv")
```

```{r}
# Anna's Laptop
#responses <- read_csv("~/Desktop/ThermalResponsesIII.csv")
```

We discarded "eggtoemergence" category in the subsequent analyses because it is a combination of the other three. We also discarded data from sets of uncertain locality ("inferred") or that were a combination of localities "combination".

```{r}
# convert character to factor
responses <- responses %>%
  mutate_if(is.character, factor)
# discard non-pertinent data

responses_table <- responses %>% 
  filter(lifestage == "egg"| lifestage == "larva"| lifestage == "pupa", 
         #quality!= "inferred", 
         quality != "combination", !is.na(Dopt)) %>% 
  mutate(absLat = abs(lat)) 
responses_table$lifestage <- factor(responses_table$lifestage)
responses_table$sp <- factor(responses_table$sp)

# Select variables to include in analyses
D.set <- select(responses_table, set, Dopt, Dupper, Dwidth, Dlower, lifestage, absLat, sp, family, quality)
D.set$genus <- word(D.set$sp, 1)
D.set$species <- word(D.set$sp, 2)
filter(responses, set == 1005)
# save a list of sets included in analyses
D.sets <- D.set %>% 
  select(set)
# write_csv(D.sets,"/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/sets_development.csv")

```
There were `length(unique(D.set$sp))`species
There were `length(unique(D.set$set))`sets
There were `nrow(filter(D.set, quality == "inferred")["quality"])/length(unique(D.set$set))`species
41/173
# 2. Dlow (Dlower)
Dlow ranged from `r min(D.set$Dlower, na.rm = T)` to `r max(D.set$Dlower, na.rm = T)`, mean =  `r mean(D.set$Dlower, na.rm = T)`, SD = `r sd(D.set$Dlower, na.rm = T)`
Latitude ranged from `r min(D.set$absLat, na.rm = T)` to `r max(D.set$absLat, na.rm = T)`

## 2.1 Model selection

```{r}

Dlower_i <- lmer(Dlower ~ -1 + lifestage + absLat + lifestage:absLat + (1|sp), data = as.data.frame(D.set)) #
Dlower_a <- lmer(Dlower ~ -1 + absLat + lifestage + (1|sp), data = as.data.frame(D.set)) #


anova(Dlower_i, Dlower_a, test.statistic = "Chisq") # keep additive model, plot interaction
```

```{r}
Dlower_a_ML <- lmer(Dlower ~ -1 + absLat + lifestage +  (1|sp), REML = F, data = as.data.frame(D.set)) #
summary(Dlower_a_ML)
```
According to this mode, Dlower was 35 for pupae, 35.2 for larvae and 35.4 for eggs. Dlower decreased 0.3 degrees C for every 10 degrees in latitude. Variation among species: 2.6

```{r}
# Predictions plot
D.set$group <- factor(D.set$lifestage)
predlow <- ggpredict(Dlower_i, terms = c("absLat","lifestage"))
Dlow_plot <- ggplot(predlow, aes(x = x, y = predicted, colour = group, fill = group)) +
  geom_line()+
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, col = group), alpha = 0.1)+
  scale_fill_viridis_d(option = "plasma", begin = 0, end = 0.8)+
  scale_color_viridis_d(option = "plasma", begin = 0, end = 0.8)+
  ylab(expression(D[low]))+
  xlab("Absolute latitude")+
  ylim(0, 40)+
  annotate("text", x = 0, y = 40, label = "G", size = 5)+
  geom_point(data = D.set, mapping = aes(x = absLat, y = Dlower, col = group), alpha = 0.6, shape = 16)+
  theme_cowplot()+
  theme(legend.title = element_blank(), legend.position = "none")

Dlow_plot
```
Effect sizes according to ML model 

## 2.2 Phylogenetic correction
### 2.2.1 Adjust data table 
Change species' names of the following species to reflect substitutions due to lack of data and to correct misspellings
Substitutions:
 In tree                  In dataset    
Ephestia columbiella for     Ephestia calidella #### Not in analyses because source population was "inferred"
Episimus tyrius for          Episimus utilis 
Euzopherodes allocrossa for Euzopherodes vapidella ### No development rate data available
Marmara arbutiella for      Marmara gulosa
Protodeltote albidula for   Naranga aenescens

Mispellings: 
Hyphantria cunea instead of   Hypantria cunea
Ameyelois transitella instead of Amyelois transitella,
```{r}
D.set$original_species <- D.set$sp
D.set$sp <- ifelse(D.set$original_species == "Episimus utilis", "Episimus tyrius", ifelse(D.set$original_species == "Marmara gulosa", "Marmara arbutiella", ifelse(D.set$original_species == "Naranga aenescens", "Protodeltote albidula", ifelse(D.set$original_species == "Hypantria cunea", "Hyphantria cunea",  ifelse(D.set$original_species == "Ameyelois transitella", "Amyelois transitella", ifelse(D.set$original_species == "Utethesia ornatrix", "Utetheisa ornatrix", ifelse(D.set$original_species == "Ephestia calidella","Ephestia columbiella", as.character(D.set$original_species))))))))

Dspecies <- tibble(sp = unique(D.set$sp))
V_species <- unique(D.set$sp)
tree_species <-  tibble(sp = as.character(paste(word(tree$tip.label, sep = "_", 4), word(tree$tip.label, sep = "_", 5))))

#setdiff(Dspecies, tree_species)
# all species in D.set are in the tree. 

#setdiff(tree_species, Dspecies) # the tree has 45 species more than D.set
```

Trim the tree so it only includes the species in D.set. First the tips have to be relabeled so they include only the species names
```{r}
rename <- as.data.frame(tibble(oldname = tree$tip.label, 
                 newname = tree_species$sp))
tree_relabel <- sub.taxa.label(tree, rename)
D.tree <- keep.tip(tree_relabel, V_species)
plotTree(D.tree)
```

```{r}
Dlow_phylo <- pglmm(Dlower ~ -1 + lifestage + absLat + (1|sp__), tree = tree, bayes= F, data = D.set, cov_ranef = list(sp = D.tree), REML = F)

Dlow_nophylo <- pglmm(Dlower ~ -1 + lifestage + absLat + (1|sp), tree = tree, bayes= F, data = D.set, cov_ranef = list(sp = D.tree), REML = F)

```

Asses the effect of phylogeny on Dlow
```{r}
R2(Dlow_phylo, Dlow_nophylo) # if this does not work unload and reload rr2
```

#3. Dhigh (Dupper)
Dhigh ranged from `r min(D.set$Dupper, na.rm = T)` to `r max(D.set$Dupper, na.rm = T)`, mean =  `r mean(D.set$Dupper, na.rm = T)`, SD = `r sd(D.set$Dupper, na.rm = T)`
Latitude ranged from `r min(D.set$absLat, na.rm = T)` to `r max(D.set$absLat, na.rm = T)`

## 3.1 Model selection

```{r}
Dupper_i <- lmer(Dupper ~ -1 + lifestage + absLat + lifestage:absLat + (1|sp), data = as.data.frame(D.set)) #
Dupper_a <- lmer(Dupper ~ -1 + absLat + lifestage + (1|sp), data = as.data.frame(D.set)) #

anova(Dupper_i, Dupper_a, test.statistic = "Chisq") # keep aditive model, plot interaction
```

```{r}
Dupper_a_ML <- lmer(Dupper ~ -1 + absLat + lifestage + (1|sp), REML = F, data = as.data.frame(D.set)) #
summary(Dupper_a_ML)
```
According to this model, Dupper was 35 for pupae, 35.2 for larvae and 35.4 for eggs.  Dupper decreased 0.3 degrees C for every 10 degrees in latitude. Variation among species: 2.6

```{r}
# Predictions plot
D.set$group <- factor(D.set$lifestage)
predu <- ggpredict(Dupper_i, terms = c("absLat","lifestage"))
Dupper_plot <- ggplot(predu, aes(x = x, y = predicted, colour = group, fill = group)) +
  geom_line()+
   geom_ribbon(aes(ymin = conf.low, ymax = conf.high, col = group), alpha = 0.1)+
  scale_fill_viridis_d(option = "plasma", begin = 0, end = 0.8)+
  scale_color_viridis_d(option = "plasma", begin = 0, end = 0.8)+
  ylab(expression(D[high]))+
  xlab("Absolute latitude")+
  ylim(10, 40)+
  annotate("text", x = 0, y = 40, label = "X", size = 5)+
  geom_point(data = D.set, mapping = aes(x = absLat, y = Dupper, col = group), alpha = 0.6, shape = 1)+
  theme_cowplot()+
  theme(legend.title = element_blank(), legend.position = "none")

Dupper_plot
```

## 3.2 Phylogenetic correction

```{r}
Dhigh_phylo <- pglmm(Dupper ~ -1 + lifestage + absLat + (1|sp__), tree = tree, bayes= F, data = D.set, cov_ranef = list(sp = D.tree), REML = F)

Dhigh_nophylo <- pglmm(Dupper ~ -1 + lifestage + absLat + (1|sp), tree = tree, bayes= F, data = D.set, cov_ranef = list(sp = D.tree), REML = F)

```

Asses the effect of phylogeny
```{r}
R2(Dhigh_phylo, Dhigh_nophylo) # if this does not work unload and reload rr2
```

Extract effect sizes
```{r}
DhighFixedEffects <- tibble(Response = "Dhigh", Term = row.names(Dhigh_phylo$B), Effect = Dhigh_phylo$B[1:4], 
SE = Dhigh_phylo$B.se[1:4])

DhighRandomEffects <- tibble(Response = "Dhigh", Term = c("sp", "sp__", "residual"), Variance = c(Dhigh_phylo$s2r,Dhigh_phylo$s2resid) , 
SD = sqrt(Variance))

```


# 4. Dwidth
Dwidth ranged from `r min(D.set$Dwidth, na.rm = T)` to `r max(D.set$Dwidth, na.rm = T)`, mean =  `r mean(D.set$Dwidth, na.rm = T)`, SD = `r sd(D.set$Dwidth, na.rm = T)`
Latitude ranged from `r min(D.set$absLat, na.rm = T)` to `r max(D.set$absLat, na.rm = T)`

## 3.1 Model selection

```{r}
Dwidth_i <- lmer(Dwidth ~ -1 + lifestage + absLat + lifestage:absLat + (1|sp), data = as.data.frame(D.set)) #
Dwidth_a <- lmer(Dwidth ~ -1 + absLat + lifestage + (1|sp), data = as.data.frame(D.set)) #

anova(Dwidth_i, Dwidth_a, test.statistic = "Chisq") # keep additive model, interaction for figures
```

```{r}
Dwidth_a_ML <- lmer(Dwidth ~ -1 + absLat + lifestage + (1|sp), REML = F, data = as.data.frame(D.set)) #
summary(Dwidth_a_ML)
```
According to this model, Dwidth was 12.1 for egg and larvae and 11.1 for pupae (one degree smaller, `r ((12.1 - 11.1)/12.1) *100`% smaller)  and Dwidth increased 0.4 degrees C for every 10 degrees in latitude. Variation among species: 1.8

```{r}
# Predictions plot
D.set$group <- factor(D.set$lifestage)
predw <- ggpredict(Dwidth_i, terms = c("absLat","lifestage"))


Dwidth_plot <- ggplot(predw, aes(x = x, y = predicted, colour = group, fill = group)) +
  geom_line()+
   geom_ribbon(aes(ymin = conf.low, ymax = conf.high, col = group), alpha = 0.1)+
  scale_fill_viridis_d(option = "plasma", begin = 0, end = 0.8)+
  scale_color_viridis_d(option = "plasma", begin = 0, end = 0.8)+
  scale_shape_manual(values = c(1,16,8))+
  ylab(expression(D[width]))+
  xlab("Absolute latitude")+
  ylim(0, 40)+
  annotate("text", x = 0, y = 40, label = "X", size = 5)+
  geom_point(data = D.set, mapping = aes(x = absLat, y = Dwidth, col = group), alpha = 0.6, shape = 0)+
  theme_cowplot()+
  theme(legend.title = element_blank(), legend.position = "none")

Dwidth_plot
```

## 4.2 Phylogenetic correction

### 4.2.2 Fit model
```{r}

Dwidth_phylo <- pglmm(Dwidth ~ -1 + lifestage + absLat + (1|sp__), tree = tree, bayes= F, data = D.set, cov_ranef = list(sp = D.tree), REML = F)

Dwidth_nophylo <- pglmm(Dwidth ~ -1 + lifestage + absLat + (1|sp), tree = tree, bayes= F, data = D.set, cov_ranef = list(sp = D.tree), REML = F)

```

Asses the effect of phylogeny
```{r}
R2(Dwidth_phylo,Dwidth_nophylo) # if this does not work unload and reload rr2
```
Extract effect sizes
```{r}
DwidthFixedEffects <- tibble(Response = "Dwidth", Term = row.names(Dwidth_phylo$B), Effect = Dwidth_phylo$B[1:4], 
SE = Dwidth_phylo$B.se[1:4])

DwidthRandomEffects <- tibble(Response = "Dwidth", Term = c("sp", "sp__", "residual"), Variance = c(Dwidth_phylo$s2r,Dwidth_phylo$s2resid) , 
SD = sqrt(Variance))

```

# 
Extract effect sizes
```{r}
DhighFixedEffects <- tibble(Response = "Dhigh", Term = row.names(Dhigh_phylo$B), Effect = Dhigh_phylo$B[1:4], 
SE = Dhigh_phylo$B.se[1:4])

DhighRandomEffects <- tibble(Response = "Dhigh", Term = c("sp", "sp__", "residual"), Variance = c(Dhigh_phylo$s2r,Dhigh_phylo$s2resid) , 
SD = sqrt(Variance))

```


Consolidate effects and save table
```{r}
DrFixedEffects <- rbind(DoptFixedEffects, DwidthFixedEffects, DhighFixedEffects)
DrRandomEffects <- rbind(DoptRandomEffects, DwidthRandomEffects, DhighRandomEffects)
 write_csv(DrFixedEffects,"/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/DrateFixedEffects.csv")
 write_csv(DrRandomEffects,"/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/DrateRandomEffects.csv")
```


Run script for T0 and remove comments from T0 line and T0 points to get full figure

```{r}
#Import output from T0_Analyses to get T0 line and points
predto <- read_csv("/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/Output/predto.csv")
T.set <- read_csv("/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/Output/Tset.csv")

PanelA <- ggplot(predlow, aes(x = x, y = predicted, colour = group, fill = group)) +
  geom_line()+
  # T0 line
  geom_line(data = predto, mapping = aes(x = x, y = predicted, colour = group, fill = group))+
  geom_line(data = predu, mapping = aes(x = x, y = predicted, colour = group, fill = group))+
  scale_fill_viridis_d()+
  scale_color_viridis_d(option = "plasma", begin = 0, end = 0.8)+ # "magma", "inferno", "plasma"
  annotate("text", x = 59, y = 10, label = expression(`T`[0]))+
  #annotate("text", x = 59, y = 10, label = expression(D[width]))+
  annotate("text", x = 59, y = 20, label = expression(D[low]))+
  annotate("text", x = 59, y = 35, label = expression(D[high]))+
  ylab("Temperature")+
  #xlab("Absolute latitude")+ # 
  xlab("")+
  ylim(0, 40)+
  annotate("text", x = 0, y = 40, label = "A", size = 5)+
  geom_point(data = D.set, mapping = aes(x = absLat, y = Dlower, col = group), alpha = 0.6, shape =16)+
  geom_point(data = D.set, mapping = aes(x = absLat, y = Dupper, col = group), alpha = 0.6, shape = 1)+
  # T0 points
  geom_point(data = T.set, mapping = aes(x = absLat, y = T0, col = group), alpha = 0.6, shape = 6)+
  theme_cowplot()+
  theme(legend.title = element_blank(), legend.position = "none")
PanelA








```





```{r}
PanelB <- ggplot(predw, aes(x = x, y = predicted, colour = group, fill = group)) +
  geom_line()+
  #geom_line(data = predto, mapping = aes(x = x, y = predicted, colour = group, fill = group))+
  #geom_line(data = predu, mapping = aes(x = x, y = predicted, colour = group, fill = group))+
  #geom_line(data = predw, mapping = aes(x = x, y = predicted, colour = group, fill = group))+
   #geom_ribbon(aes(ymin = conf.low, ymax = conf.high, col = group), alpha = 0.1)+
  #scale_color_viridis_d()+
  scale_fill_viridis_d()+
  scale_color_viridis_d(option = "plasma", begin = 0, end = 0.8)+ # "magma", "inferno", "plasma"
  #scale_shape_manual(values = c(1,16,8))+
  #ylab(expression(`T`[0]-D[low]-D[high]))+
  #annotate("text", x = 59, y = 10, label = expression(`T`[0]))+
  annotate("text", x = 50, y = 5, label = expression(D[width]))+
  #annotate("text", x = 59, y = 20, label = expression(D[low]))+
  #annotate("text", x = 59, y = 35, label = expression(D[high]))+
  ylab("Temperature")+
  xlab("Absolute latitude")+
  ylim(0, 30)+
  annotate("text", x = 0, y = 30, label = "B", size = 5)+
  #geom_point(data = D.set, mapping = aes(x = absLat, y = Dlower, col = group), alpha = 0.6, shape =16)+
  geom_point(data = D.set, mapping = aes(x = absLat, y = Dwidth, col = group), alpha = 0.6, shape = 0)+
  #geom_point(data = D.set, mapping = aes(x = absLat, y = Dupper, col = group), alpha = 0.6, shape = 1)+
  #geom_point(data = T.set, mapping = aes(x = absLat, y = T0, col = group), alpha = 0.6, shape = 6)+
  theme_cowplot()+
  theme(legend.title = element_blank(), legend.position = "none")
PanelB
```


Save data to make figures:
predlow
predu
D.set
```{r}
#write_csv(predlow,"/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/Output/predlow.csv")
#write_csv(predu,"/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/Output/predu.csv")
#write_csv(D.set,"/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/Output/Dset.csv")
#write_csv(predw,"/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/Output/predw.csv")
```







### Run this for grid plot 

(Run scripts for development, survival, performance and TO to get all objects)
```{r}
plot_grid(Dopt_plot, Sopt_plot, Popt_plot, Pwidth_plot, Phigh_plot, T0_plot, Dlow_plot, Slow_plot, Plow_plot, ncol = 3)
```
```{r}
plot_grid(PanelA, PanelC, PanelE, PanelB, PanelD, PanelF, ncol = 3)
```




# * Old code *****************************************************************************************  
 
Correlation coefficients of development variables 
 
```{r}
names(D.set)
cor(D.set[1:4])
rcorr(as.matrix(D.set[1:4])) # correlation coefficients and then P values
```


# PCA
PCA of development responses 
```{r}
## Omit NA rows for PCA. need dataframe for input
D.set <- with(responses_table, cbind(Dopt, Dupper, Dwidth,Dlower, lifestage, absLat, sp))
D.set.nona <- as.data.frame(na.omit(D.set))
Dpc <- princomp(~Dopt + Dupper + Dwidth, data = D.set.nona)
summary(Dpc)
```
The first principal component explains 74% of the total variation.
The second component explains 19% of the total variation.

```{r}
loadings(Dpc)
```
Dopt and Dupper are highly correlated (have high positive loadings in component 1)
Dopt and Dwith are negatively correlated. 
 
```{r}

## Bind PC scores to dataset
D.set.pcs <- cbind(D.set.nona, Dpc$scores)
head(D.set.pcs)

## Now model Comp.1 (PC scores on 1st PC)
fit1_red1 <- lmer(Comp.1 ~ lifestage + absLat + (lifestage|sp), data = D.set.pcs)
summary(fit1_red1)

## test simpler models
fit1_red2 <- lmer(Comp.1 ~ lifestage  + (lifestage|sp), data = D.set.pcs)
fit1_red3 <- lmer(Comp.1 ~ absLat + (lifestage|sp), data = D.set.pcs)
fit1_red4 <- lmer(Comp.1 ~ (lifestage|sp), data = D.set.pcs)
anova(fit1_red1,fit1_red2, fit1_red3,fit1_red4, test.statistic = "Chisq")

## estimate parameters using ML rather than REML (default)
fit1_red1.ML <- lmer(Comp.1 ~ lifestage + absLat  + (lifestage|sp), data = D.set.pcs, REML = F)
summary(fit1_red1.ML)

dw_plot(fit1_red1.ML) +
  #geom_vline(xintercept= mean(D.set$Dopt),lty=2)+
  geom_vline(xintercept= 0,lty=3)+
  theme_cowplot()+
  scale_color_viridis_d()+
  theme(legend.position = "none")


```
 

 
PCA of ALL development responses (same result as first PCA, not included in report)
```{r}
## Omit NA rows for PCA. need dataframe for input
D.set.nona2 <- as.data.frame(na.omit(D.set))
Dpc2 <- princomp(~Dopt + Dupper + Dwidth + Dlower, data = D.set.nona2)
summary(Dpc2)
```
The first principal component explains 67% of the total variation.
The second component explains 26% of the total variation.

```{r}
loadings(Dpc2)
```
Dopt and Dupper are highly correlated (have high positive loadings in component 1)
Dopt and Dwith are negatively correlated. 
 
```{r}

## Bind PC scores to dataset
D.set.pcs2 <- cbind(D.set.nona2, Dpc2$scores)
head(D.set.pcs2)

## Now model Comp.1 (PC scores on 1st PC)
fit5_red1 <- lmer(Comp.1 ~ lifestage + absLat + (lifestage|sp), data = D.set.pcs2)
summary(fit5_red1)

## test simpler models
fit5_red2 <- lmer(Comp.1 ~ lifestage  + (lifestage|sp), data = D.set.pcs2)
fit5_red3 <- lmer(Comp.1 ~ absLat + (lifestage|sp), data = D.set.pcs2)
fit5_red4 <- lmer(Comp.1 ~ (lifestage|sp), data = D.set.pcs2)
anova(fit5_red1,fit5_red2, fit5_red3,fit5_red4, test.statistic = "Chisq")

## estimate parameters using ML rather than REML (default)
fit5_red3.ML <- lmer(Comp.1 ~ lifestage + absLat  + (lifestage|sp), data = D.set.pcs2, REML = F)
summary(fit5_red3.ML)

dw_plot(fit5_red3.ML) +
  #geom_vline(xintercept= mean(D.set$Dopt),lty=2)+
  geom_vline(xintercept= 0,lty=3)+
  theme_cowplot()+
  scale_color_viridis_d()+
  theme(legend.position = "none")


```
 
 # Optimum
 # 2. Dopt: temperature that maximizes development rate
Dopt ranged from `r min(D.set$Dopt, na.rm = T)` to `r max(D.set$Dopt, na.rm = T)`, mean =  `r mean(D.set$Dopt, na.rm = T)`, SD = `r sd(D.set$Dopt, na.rm = T)`, N =`r length(!is.na(D.set$Dopt))` 
Latitude ranged from `r min(D.set$absLat, na.rm = T)` to `r max(D.set$absLat, na.rm = T)`.

## 2.1 Model selection
To asses the effect of latitude and lifestage on variation in Dopt, we fitted two linear mixed effects models one  including with both terms and their interaction as fixed effects and the other only both terms. In both cases we included species as a random effect, we used AIC and BIC as criteria for model selection.  We used the lmer function from package lme4 version `r packageVersion("lme4")`. We fitted the models without an intercept to facilitate effect comparisons among life stages. We kept the additive model

```{r}

Dopt_i <- lmer(Dopt ~ -1 + lifestage + absLat + lifestage:absLat + (1|sp), data = as.data.frame(D.set)) #
Dopt_a <- lmer(Dopt ~ -1 + absLat + lifestage + (1|sp), data = as.data.frame(D.set)) #

anova(Dopt_i, Dopt_a, test.statistic = "Chisq") # keep interaction model
```


```{r}
Dopt_i_ML <- lmer(Dopt ~ -1 + absLat + lifestage + lifestage:absLat + (1|sp), REML = F, data = as.data.frame(D.set)) #
# Anova(Dopt_a_ML)
summary(Dopt_i_ML)
```
According to this model, Dopt was close to 32. for all lifestages and Dopt increased 0.6 degrees C for every 10 degrees in latitude. Variation among species: 2.3

```{r}
# Predictions plot
D.set$group <- factor(D.set$lifestage)
predo <- ggpredict(Dopt_i_ML, terms = c("absLat","lifestage"))


Dopt_plot <- ggplot(predo, aes(x = x, y = predicted, colour = group, fill = group, shape = group)) +
  geom_line()+
   geom_ribbon(aes(ymin = conf.low, ymax = conf.high, col = group), alpha = 0.1)+
  scale_color_viridis_d(option = "plasma", begin = 0, end = 0.8)+ # "magma", "inferno", "plasma"
  scale_fill_viridis_d()+
  scale_shape_manual(values = c(1,16,8))+
  ylab(expression(D[opt]))+
  xlab("Absolute latitude")+
  ylim(10, 40)+
  annotate("text", x = 0, y = 40, label = "A", size = 5)+
  geom_point(data = D.set, mapping = aes(x = absLat, y = Dopt, col = group), alpha = 0.6)+
  theme_cowplot()+
  theme(legend.title = element_blank(), legend.position = c(0.8,0.3))

Dopt_plot
```


## 2.2 Phylogenetic correction
### 2.2.1 Adjust data table 
Change species' names of the following species to reflect substitutions due to lack of data and to correct misspellings

Substitutions:
 In tree                  In dataset    
Ephestia columbiella for     Ephestia calidella #### Not in analyses because source population was "inferred"
Episimus tyrius for          Episimus utilis 
Euzopherodes allocrossa for Euzopherodes vapidella ### No development rate data available
Marmara arbutiella for      Marmara gulosa
Protodeltote albidula for   Naranga aenescens

Mispellings: 
Hyphantria cunea instead of   Hypantria cunea
Ameyelois transitella instead of Amyelois transitella,
```{r}
D.set$original_species <- D.set$sp
D.set$sp <- ifelse(D.set$original_species == "Episimus utilis", "Episimus tyrius", ifelse(D.set$original_species == "Marmara gulosa", "Marmara arbutiella", ifelse(D.set$original_species == "Naranga aenescens", "Protodeltote albidula", ifelse(D.set$original_species == "Hypantria cunea", "Hyphantria cunea",  ifelse(D.set$original_species == "Ameyelois transitella", "Amyelois transitella", ifelse(D.set$original_species == "Utethesia ornatrix", "Utetheisa ornatrix", ifelse(D.set$original_species == "Ephestia calidella","Ephestia columbiella", as.character(D.set$original_species))))))))

Dspecies <- tibble(sp = unique(D.set$sp))
V_species <- unique(D.set$sp)
tree_species <-  tibble(sp = as.character(paste(word(tree$tip.label, sep = "_", 4), word(tree$tip.label, sep = "_", 5))))

#setdiff(Dspecies, tree_species)
# all species in D.set are in the tree. 

#setdiff(tree_species, Dspecies) # the tree has 45 species more than D.set
```

Trim the tree so it only includes the species in D.set. First the tips have to be relabeled so they include only the species names
```{r}
rename <- as.data.frame(tibble(oldname = tree$tip.label, 
                 newname = tree_species$sp))
tree_relabel <- sub.taxa.label(tree, rename)
D.tree <- keep.tip(tree_relabel, V_species)
plotTree(D.tree)
```

### 2.2.2 Fit models

```{r}
# 1. with phylogenetic correction
Dopt_phylo <- pglmm(Dopt ~ -1 + lifestage + absLat +(1|sp__),  bayes= F, data = D.set, cov_ranef = list(sp = D.tree), REML = F)
# 2. With sp as a random effect 
Dopt_nophylo <- pglmm(Dopt ~ -1 + lifestage + absLat + (1|sp),  bayes= F, data = D.set, cov_ranef = list(sp = D.tree), REML = F)
summary(Dopt_phylo)
```

```{r}
summary(Dopt_nophylo)
```


```{r}

# Run this to asses the effect of phylogeny on Dopt

R2(Dopt_phylo,Dopt_nophylo) # if this does not work unload and reload rr2

```
Small values or R2_lik indicate almost no effect of phylogeny on response variable

Extract effect sizes for plot
```{r}
DoptFixedEffects <- tibble(Response = "Dopt", Term = row.names(Dopt_phylo$B), Effect = Dopt_phylo$B[1:4], 
SE = Dopt_phylo$B.se[1:4])

DoptRandomEffects <- tibble(Response = "Dopt", Term = c("sp", "sp__", "residual"), Variance = c(Dopt_phylo$s2r,Dopt_phylo$s2resid) , 
SD = sqrt(Variance))

```


 
 