---
title: "Relationships among thermal variables"
output: html_notebook
---

Manuscript Figure 3

# 1. Import data and housekeeping
```{r}
 
library(tidyverse)
library(readxl)
library(cowplot)
library(psych) # to get geometric mean
#library(lme4)
#library(car)
#library("dotwhisker")  # to make coefficient plots
#library(sjPlot) #prediction plots
#library(sjmisc)
#library(broom)
#library(effects)
#library(ggeffects)
#library(Hmisc)
#library(phyr)
#library(ape)
#library(nlme)
#library(stringr)
#library(phylotools)
#library(phytools)
#library(rr2)
```




```{r}
responses <- read_csv("/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/ThermalResponsesTable.csv") 
res <- responses %>% 
  filter(lifestage == "egg"| lifestage == "larva"| lifestage == "pupa", quality != "combination") %>% 
  mutate(lifestage = factor(lifestage))
names(res)


res$skew <- (res$Popt - res$Plower) /res$Pwidth
res_plot5 <- res %>% 
  filter(!is.na(Dlower),  !is.na(Slower))

  res_plot3 <- res %>% 
  filter(!is.na(Swidth),  !is.na(Dwidth))

  res_plot6 <- res %>% 
  filter(!is.na(T0),  !is.na(Dlower))
  res_plot6 <- res %>% 
  filter(!is.na(T0),  !is.na(Dlower))
max(res_plot6$Dlower)
res_plot7 <- res %>% 
  filter(!is.na(Plower),  !is.na(Slower), !is.na(Dlower))

  res_plot2 <- res %>% 
  filter( !is.na(Supper),  !is.na(Dupper))
res_plot <- res %>% 
  filter( !is.na(Dopt),  !is.na(maxSopt))

```


Dopt for eggs, larvae and pupae ranged from `r min(res$Dopt, na.rm = T)` to `r max(res$Dopt, na.rm = T)`, with a mean (+_SD) of `r min(res$Dopt, na.rm = T)` +- `r SD(res$Dopt, na.rm = T)`

Predictions from figure 2. 
# Figure 3A, Dlow ~ Slow

```{r}
fitdlows <- lm(Dlower ~ Slower, data = res)
shapiro.test(fitdlows$residuals)
summary(fitdlows)

res$Ddif <- res$Dlower - res$Slower
res$Highdif <- res$Dupper - res$Supper
min(res$Ddif, na.rm = T)
mean(res$Ddif, na.rm = T)
```

The lower limit for development Dlow ranged from `r min(res$Dlower, na.rm = T)` to `r max(res$Dlower, na.rm = T)` and on average it was `r mean(res$Ddif, na.rm = T)` degrees higher than the corresponding Slow; which in turn ranged from `r min(res$Slower, na.rm = T)` to `r max(res$Slower, na.rm = T)` 
The higher limit for development Dhigh ranged from `r min(res$Dupper, na.rm = T)` to `r max(res$Dupper, na.rm = T)` and on average it was `r mean(res$Highdif, na.rm = T)` degrees higher than the corresponding Slow Shigh which in turn ranged from `r min(res$Supper, na.rm = T)` to `r max(res$Supper, na.rm = T)`

T0 from `r min(res$T0, na.rm = T)` to `r max(res$T0, na.rm = T)`  
#Dupper ~ Supper
```{r}
fitb <- lm(Dupper ~ Supper, data = res)
shapiro.test(fitb$residuals)
summary(fitb)
```


Dlow and Slow were correlated,R2= 0.4, P < 0.0001, N = `r nrow(model.frame(fitdlows))`).  
Dlow for eggs, larvae and pupae ranged from `r min(res$low, na.rm = T)` to `r max(res$Dlow, na.rm = T)`, with a mean (+_SD) of `r mean(res$Dopt, na.rm = T)` +- `r SD(res$Dopt, na.rm = T)`

```{r}
res_plot5$AbsLat <- abs(res_plot5$lat)
res_plot2$AbsLat <- abs(res_plot2$lat)
res_plot3$AbsLat <- abs(res_plot3$lat)
res_plot6$AbsLat <- abs(res_plot6$lat)

res_plot7$AbsLat <- abs(res_plot7$lat)
lows <- ggplot(res_plot5, aes(x = Slower, y = Dlower, col = AbsLat))+
  geom_segment(x = 0, xend = 40, y = 0, yend = 40, col = "darkgrey", linetype = 2)+
  geom_point()+
  geom_point(data = res_plot2, aes(x = Supper, y = Dupper, col = AbsLat), shape = 1)+
  #geom_point(data = res_plot3, aes(x = Swidth, y = Dwidth, col = AbsLat), shape = 3)+
  ylim(0,40)+
  xlim(0,40)+
  geom_smooth(method = lm, se = F, col = "black")+
  geom_smooth(data = res_plot2, aes(x = Supper, y = Dupper), method = lm, se = F, col = "black", linetype = 1)+
  #geom_smooth(data = res_plot3, aes(x = Swidth, y = Dwidth), method = lm, se = F, col = "black", linetype = 1)+
  scale_shape_manual(values = c(1,16,8))+
  scale_color_viridis_c()+
  #scale_color_viridis_d()+
  
  theme_cowplot()+
  #ylab(expression("D"[low]))+
  ylab("Development temperature")+
  annotate("text", x = 1, y = 40, label = "A", size = 5)+
  #xlab(expression("S"[low]))+
  xlab("Survival temperature")+
  theme(legend.position = "none")+
  annotate("text", x = 30, y = 10, label = "low")+
  annotate("point", x = 33, y = 10, shape = 16)+
  annotate("text", x = 30, y = 15, label = "high")+
  annotate("point", x = 33, y = 15, shape = 1)
  #annotate("text", x = 30, y = 5, label = "width")+
  #annotate("point", x = 33, y = 5, shape = 3)
lows
```
`r mean(res$Dwidth, na.rm = T)` +- `r SD(res$Dwidth, na.rm = T)`
`r mean(res$Swidth, na.rm = T)` +- `r SD(res$Swidth, na.rm = T)`
`r mean(res$Swidth- res$Dwidth, na.rm = T)`
`r min(res$Swidth, na.rm = T)`
`r max(res$Swidth, na.rm = T)`
`r min(res$Dwidth, na.rm = T)`
`r max(res$Dwidth, na.rm = T)`


```{r}
Figure3b <- ggplot(res_plot5, aes(x = Slower, y = Dlower, col = AbsLat))+
  geom_segment(x = 0, xend = 40, y = 0, yend = 40, col = "darkgrey", linetype = 2)+
  #geom_point()+
  #geom_point(data = res_plot2, aes(x = Supper, y = Dupper, col = AbsLat), shape = 1)+
  geom_point(data = res_plot3, aes(x = Swidth, y = Dwidth, col = AbsLat), shape = 3)+
  ylim(0,40)+
  xlim(0,40)+
  #geom_smooth(method = lm, se = F, col = "black")+
  #geom_smooth(data = res_plot2, aes(x = Supper, y = Dupper), method = lm, se = F, col = "black", linetype = 1)+
  geom_smooth(data = res_plot3, aes(x = Swidth, y = Dwidth), method = lm, se = F, col = "black", linetype = 1)+
  scale_shape_manual(values = c(1,16,8))+
  scale_color_viridis_c()+
  #scale_color_viridis_d()+
  
  theme_cowplot()+
  #ylab(expression("D"[low]))+
  ylab("Development temperature")+
  annotate("text", x = 0, y = 40, label = "B", size = 5)+
  #xlab(expression("S"[low]))+
  xlab("Survival temperature")+
  theme(legend.position = "none")
  #annotate("text", x = 30, y = 10, label = "low")+
  #annotate("point", x = 33, y = 10, shape = 16)+
  #annotate("text", x = 30, y = 15, label = "high")+
  #annotate("point", x = 33, y = 15, shape = 1)
  #annotate("text", x = 20, y = 5, label = "width")+
  #annotate("point", x = 22, y = 5, shape = 3)
Figure3b

```

```{r}
fitT0 <- lm(T0 ~ Dlower, data = res_plot6)

shapiro.test(fitT0$residuals)
hist(fitT0$residuals)
summary(fitT0)
library(quantreg)

fitT0q <- rq(T0 ~ Dlower, data =res_plot6,
             tau = 0.5)

summary(fitT0q)

nullmod <- rq(T0 ~ 1, data = res_plot6, tau = 0.5)
summary(fitT0q)
anova(fitT0q, nullmod)
library(rcompanion)
nagelkerke(fitT0q)


```



```{r}

Ts <- ggplot(res_plot6, aes(x = T0, y = Dlower, col = AbsLat))+
  geom_segment(x = 0, xend = 40, y = 0, yend = 40, col = "darkgrey", linetype = 2)+
  geom_point()+
  ylim(0,40)+
  xlim(0,40)+
  geom_smooth(method = lm, se = F, col = "black")+
  scale_shape_manual(values = c(1,16,8))+
  scale_color_viridis_c()+
  theme_cowplot()+
  ylab(expression("D"[l]))+
  #annotate("text", x = 0, y = 29, label = "E", size = 5)+
  annotate("text", x = 0, y = 40, label = "C", size = 5)+
  xlab(expression("T"[0]))+
  theme(legend.position = c(0.7, 0.4))
Ts
```




#Figure 3B P~S; P~D [Low]
```{r}



fitPlowsD <- lm(Plower ~ Dlower, data = res_plot7)
fitPlowsS <- lm(Plower ~ Slower, data = res_plot7)

shapiro.test(fitPlowsD$residuals)
shapiro.test(fitPlowsS$residuals)

summary(fitPlowsD)
```


Dhigh and Shigh were significantly correlated adjusted R2 = 0.78, F = 119, degrees of freedom: 1,33, P < 0.0001, N = `r nrow(model.frame(fitb))`). 



```{r}

res_plot9 <- res %>% 
  filter(!is.na(Pupper),  !is.na(Supper))
res_plot9$AbsLat <- abs(res_plot9$lat)
res$AbsLat <- abs(res$lat)

Plow <- ggplot(res, aes(x = Slower, y = Plower))+
  # survival low
  geom_segment(x = 0, xend = 40, y = 0, yend = 40, col = "darkgrey", linetype = 2)+
  scale_color_viridis_c()+
  geom_point()+
  # development low
  geom_point(data = res_plot7, aes(x = Dlower, y = Plower), col = "grey48")+
  # survival high
  #geom_point(data = res_plot9, aes(x = Supper, y = Pupper), shape = 1, col = "black")+
  # development high
  #geom_point(data = res_plot9, aes(x = Dupper, y = Pupper), shape = 1, col = "grey")+
  
  ylim(0,40)+
  xlim(0,40)+
  geom_smooth(method = lm, se = F, col = "black", linetype = 1)+
  geom_smooth(data = res_plot7, aes(x = Dlower, y = Plower), col = "grey48", method = lm, se = F, linetype = 1)+
  #geom_smooth(data = res_plot9, aes(x = Supper, y = Pupper), col = "black", method = lm, se = F, linetype = 1)+
  #geom_smooth(data = res_plot9, aes(x = Dupper, y = Pupper), col = "grey", method = lm, se = F, linetype = 1)+
  scale_shape_manual(values = c(1,16,8))+
  
  #geom_segment(x = 5, xend = 30, y = 15, yend = 15, col = "black", linetype = 2)+
  theme_cowplot()+
  ylab(expression("P"[l]))+
  #ylab("Performance temperature")+ 
  #annotate("text", x = 5, y = 30, label = "F", size = 5)+
   annotate("text", x = 1, y = 40, label = "D", size = 5)+
   annotate("text", x = 30, y = 10, label = expression("D"[l]), size = 5, col = "grey48")+
   annotate("text", x = 30, y = 5, label = expression("S"[l]), size = 5, col = "black")+
   #xlab("Survival/ Development \n temperature")+
  #xlab(expression("S(black)/D(grey)"[low]))+
  xlab(expression("S"[l]|"D"[l]))+ 
  xlab("Temperature")+ 
  theme(legend.position = "none")
Plow
```





```{r}
Phigh <- ggplot(res, aes(x = Slower, y = Plower))+
  
  geom_segment(x = 0, xend = 40, y = 0, yend = 40, col = "darkgrey", linetype = 2)+# survival low
  #geom_point(col = "black")+
  # development low
  #geom_point(data = res_plot7, aes(x = Dlower, y = Plower), col = "grey")+
  # survival high
  geom_point(data = res_plot9, aes(x = Supper, y = Pupper), shape = 1, col = "black")+
  # development high
  geom_point(data = res_plot9, aes(x = Dupper, y = Pupper), shape = 1, col = "grey48")+
  
  ylim(0,40)+
  xlim(0,40)+
  #geom_smooth(method = lm, se = F, col = "black", linetype = 1)+
  #geom_smooth(data = res_plot7, aes(x = Dlower, y = Plower), col = "grey", method = lm, se = F, linetype = 1)+
  geom_smooth(data = res_plot9, aes(x = Supper, y = Pupper), col = "black", method = lm, se = F, linetype = 1)+
  geom_smooth(data = res_plot9, aes(x = Dupper, y = Pupper), col = "grey48", method = lm, se = F, linetype = 1)+
  scale_shape_manual(values = c(1,16,8))+
  #scale_color_viridis_d()+
  #geom_segment(x = 5, xend = 30, y = 15, yend = 15, col = "black", linetype = 2)+
  theme_cowplot()+
  ylab(expression("P"[h]))+
  #ylab("Performance temperature")+ 
  #annotate("text", x = 5, y = 30, label = "F", size = 5)+
   annotate("text", x = 1, y = 40, label = "E", size = 5)+
   annotate("text", x = 30, y = 10, label = expression("D"[h]), size = 5, col = "grey48")+
   annotate("text", x = 30, y = 5, label = expression("S"[h]), size = 5, col = "black")+
   #xlab("Survival/ Development \n temperature")+
  #xlab(expression("S(black)/D(grey)"[low]))+
  xlab(expression("S"[l]|"D"[l]))+ 
  xlab("Temperature")+ 
  theme(legend.position = "none")
Phigh
```



```{r}
 plot_grid(lows, Figure3b, Ts,Plow,  Phigh, ncol = 3 )
```






```{r}
res_plot9 <- res %>% 
  filter(!is.na(Pupper),  !is.na(Supper))
res_plot9$AbsLat <- abs(res_plot9$lat)
res$AbsLat <- abs(res$lat)

widthplot <- ggplot(res, aes(x = Pwidth, y = Plower))+
  # survival low
  geom_segment(x = 0, xend = 40, y = 0, yend = 40, col = "darkgrey", linetype = 2)+
  scale_color_viridis_c()+
  geom_point()+
  # development low
  geom_point(data = res_plot7, aes(x = Pwidth, y = Pupper), col = "grey48")+
  # survival high
  #geom_point(data = res_plot9, aes(x = Supper, y = Pupper), shape = 1, col = "black")+
  # development high
  #geom_point(data = res_plot9, aes(x = Dupper, y = Pupper), shape = 1, col = "grey")+
  
  ylim(0,40)+
  xlim(0,40)+
  geom_smooth(method = lm, se = F, col = "black", linetype = 1)+
  geom_smooth(data = res, aes(x = Pwidth, y = Pupper), col = "grey48", method = lm, se = F, linetype = 1)+
  #geom_smooth(data = res_plot9, aes(x = Supper, y = Pupper), col = "black", method = lm, se = F, linetype = 1)+
  #geom_smooth(data = res_plot9, aes(x = Dupper, y = Pupper), col = "grey", method = lm, se = F, linetype = 1)+
  scale_shape_manual(values = c(1,16,8))+
  
  #geom_segment(x = 5, xend = 30, y = 15, yend = 15, col = "black", linetype = 2)+
  theme_cowplot()+
  ylab(expression("P"[breadth]))+
  #ylab("Performance temperature")+ 
  #annotate("text", x = 5, y = 30, label = "F", size = 5)+
   annotate("text", x = 1, y = 40, label = "C", size = 5)+
   annotate("text", x = 30, y = 10, label = expression("D"[low]), size = 5, col = "grey48")+
   annotate("text", x = 30, y = 5, label = expression("S"[low]), size = 5, col = "black")+
   #xlab("Survival/ Development \n temperature")+
  #xlab(expression("S(black)/D(grey)"[low]))+
  xlab(expression("S"[low]|"D"[low]))+ 
  xlab("Temperature")+ 
  theme(legend.position = "none")
widthplot
```



```{r}

high <- ggplot(res_plot2, aes(x = Supper, y = Dupper))+
  geom_segment(x = 0, xend = 40, y = 0, yend = 40, col = "darkgrey", linetype = 2)+
  geom_point(aes(shape = lifestage))+
  ylim(0,40)+
  xlim(0,40)+
  #scale_color_viridis_d()+
  geom_smooth(method = lm, se = F, col = "black")+
  scale_shape_manual(values = c(1,16,8))+
  
  theme_cowplot()+
  #annotate("text", x = 15, y = 40, label = "C", size = 5)+
  annotate("text", x = 1, y = 40, label = "B", size = 5)+
  
  ylab(expression("D"[high]))+
  xlab(expression("S"[high]))+
  theme(legend.position = "none")

high
```



```{r}

summary(fitPlowsS)
```

```{r}


Plow <- ggplot(res, aes(x = Slower, y = Plower))+
  geom_point(aes(shape = lifestage), col = "black")+
  geom_point(data = res_plot7, aes(x = Dlower, y = Plower, shape = lifestage), col = "grey")+
  ylim(0,40)+
  xlim(0,40)+
  geom_smooth(method = lm, se = F, col = "black")+
  geom_smooth(data = res_plot7, aes(x = Dlower, y = Plower), col = "grey", method = lm, se = F)+
  scale_shape_manual(values = c(1,16,8))+
  #scale_color_viridis_d()+
  #geom_segment(x = 5, xend = 30, y = 15, yend = 15, col = "black", linetype = 2)+
  theme_cowplot()+
  ylab(expression("P"[low]))+
   #annotate("text", x = 5, y = 30, label = "F", size = 5)+
   annotate("text", x = 1, y = 40, label = "E", size = 5)+
   annotate("text", x = 5, y = 5, label = expression("D"[low]), size = 5, col = "grey")+
   annotate("text", x = 5, y = 1, label = expression("S"[low]), size = 5, col = "black")+
  #xlab(expression("S(black)/D(grey)"[low]))+
  xlab(expression("S"[low]/"D"[low]))+ 
  theme(legend.position = "none")
Plow
```








```{r}

fitPs <- lm(Pupper ~ Plower, data = res_plot2)
fitPs <- lm(Pupper ~ Plower, data = res_plot2)


shapiro.test(fitPs$residuals)


summary(fitPs)
```




```{r}
some <- ggplot(res_plot2, aes(x = Plower, y = Pupper))+
  geom_point()+
  ylim(20,40)+
  xlim(0,40)+
  #scale_color_viridis_d()+
  geom_smooth(method = lm, se = F, col = "black")+
  #scale_shape_manual(values = c(1,16,8))+
  geom_segment(x = 20, xend = 20, y = 0, yend = 40, col = "black", linetype = 2)+
  geom_segment(x = 0, xend = 40, y = 20, yend = 20, col = "black", linetype = 2)+
  theme_cowplot()+
  #annotate("text", x = 15, y = 40, label = "C", size = 5)+
  annotate("text", x = 1, y = 40, label = "B", size = 5)+
  
  xlab(expression("P"[low]))+
  ylab(expression("P"[high]))+
  theme(legend.position = "none")

some
```



```{r}
ggplot(res, aes(x = Pupper, y = Dupper-Supper))+
  geom_point(col = "red")+
  geom_point(data = res, mapping = aes(x = Plower, y = Dlower-Slower), col = "blue")+
  ylim(-15,15)+
  #xlim(0,40)+
  #scale_color_viridis_d()+
  geom_smooth(method = lm, se = F, col = "black")+
  #scale_shape_manual(values = c(1,16,8))+
  #geom_segment(x = 20, xend = 20, y = 0, yend = 40, col = "black", linetype = 2)+
  geom_segment(x = 0, xend = 40, y = 0, yend = 0, col = "black", linetype = 2)+
  theme_cowplot()+
  #annotate("text", x = 15, y = 40, label = "C", size = 5)+
  annotate("text", x = 1, y = 20, label = "B", size = 5)+
  ylab("Difference (D-S)")+
  xlab("Performance limit, lower blue, upper red")+
  #ylab(expression("P"[high]))+
  theme(legend.position = "none")
```















```{r}
fita <- lm(Dopt ~ maxSopt, data = res)
shapiro.test(fita$residuals)
summary(fita)

```
Dopt and Sopt were significantly correlated adjuster R2 = 0.14, F = 12.7, degrees of freedom: 1,70, P < 0.0001, N = `r nrow(model.frame(fita))` ). Dopt was always higher than Sopt 




```{r}



optima <- ggplot(res_plot, aes(x = maxSopt, y = Dopt))+
  geom_smooth(method = lm, se = FALSE, col = "black")+
  geom_point(aes(shape = lifestage))+
  ylim(15,40)+
  xlim(15,40)+
  ylab(expression("D"[opt]))+
  xlab(expression("S"[opt]))+
  #scale_color_grey()+
  scale_shape_manual(values = c(1,16,8))+
  geom_segment(x = 10, xend = 40, y = 10, yend = 40, col = "black", linetype = 2)+
   annotate("text", x = 15, y = 40, label = "B", size = 5)+
  theme_cowplot()+
  theme(legend.position = "none")

optima

```
Figure 3A N = `r nrow(res_plot)`


Figure 3B N = `r nrow(res_plot2)`

#Figure 2C, Dwidth ~ Swidth

```{r}
fitc <- lm(Dwidth ~ Swidth, data = res)
shapiro.test(fitc$residuals)
summary(fitc)
```
Dwidth and Swidth were significantly correlated adjusted R2 = 0.14, F = 5.124, degrees of freedom: 1,24, P < 0.03, N = `r nrow(model.frame(fitc))`). Dopt was always higher than Sopt 


```{r}
res_plot3 <- res %>% 
  filter(!is.na(Swidth),  !is.na(Dwidth))

width <- ggplot(res_plot3, aes(x = Swidth, y = Dwidth))+
  geom_point(aes(shape = lifestage))+
  ylim(0,40)+
  xlim(0,40)+
  geom_smooth(method = lm, se = F, col = "black")+
  scale_shape_manual(values = c(1,16,8))+
  #scale_color_viridis_d()+
  geom_segment(x = 0, xend = 40, y = 15, yend = 15, col = "black", linetype = 2)+
  theme_cowplot()+
  ylab(expression("D"[width]))+
   #annotate("text", x = 5, y = 30, label = "D", size = 5)+
   annotate("text", x = 1, y = 40, label = "C", size = 5)+
  xlab(expression("S"[width]))+
  theme(legend.position = "none")
width
```
Figure 3B N = `r nrow(res_plot3)`



to describe the unbalance in thermal tolerance at supra vs sub optimal temperatures, we defined skew as Skew = (Phigh – Popt) /(Popt – Plow). A skew value of 1, occurs when Pwidth occurs solely at optimal and suboptimal temperatures; 0 occurs when Pwidth is equally distributed between sub and supra-optimal temperatures (Figure 1F). Negative skew values would indicate the portion of Pwidth above Popt is larger than that below it.

# Old Figure 2D, skew ~ Pupper, discarded

```{r}
fitd <- lm(skew ~ Pupper, data = res)
shapiro.test(fitc$residuals)
summary(fitd)
```
Skew and Phigh varied independently, F = 0.02, degrees of freedom: 1, 58, P < 0.88, N = `r nrow(model.frame(fitd))`).  


```{r}
res_plot4 <- res %>% 
  filter(!is.na(skew),  !is.na(Supper))

skew <- ggplot(res_plot4, aes(x = Supper, y = skew))+
  geom_point(aes(shape = lifestage))+
  ylim(0,1)+
  #xlim(5,30)+
  
  scale_shape_manual(values = c(1,16,8))+
  #scale_color_viridis_d()+
  geom_segment(x = 25, xend = 37, y = 1, yend = 0, col = "black", linetype = 2)+
  theme_cowplot()+
  ylab("skew")+
  annotate("text", x = 24.5, y = 1, label = "D", size = 5)+
  xlab(expression("P"[high]))+
  theme(legend.position = "bottom")
skew
```
Figure 3B N = `r nrow(res_plot3)`

`r min(res$skew, na.rm = T)`

`r mean(res$skew, na.rm = T)`
`r sd(res$skew, na.rm = T)`
`r max(res$skew, na.rm = T)`
`r sum(!is.na(res$skew))`









#Figure T0 T0 ~ Dlow

```{r}
res_plot6 <- res %>% 
  filter(!is.na(T0),  !is.na(Dlower))
fitT0 <- lm(T0 ~ Dlower, data = res_plot6)

shapiro.test(fitT0$residuals)
hist(fitT0$residuals)
summary(fitT0)
```



```{r}
res_plot6 <- res %>% 
  filter(!is.na(T0),  !is.na(Dlower))
max(res_plot6$Dlower)
Ts <- ggplot(res_plot6, aes(x = T0, y = Dlower))+
  geom_point(aes(shape = lifestage))+
  ylim(0,40)+
  xlim(0,40)+
  geom_smooth(method = lm, se = F, col = "black")+
  scale_shape_manual(values = c(1,16,8))+
  #scale_color_viridis_d()+
  geom_segment(x = 0, xend = 40, y = 0, yend = 40, col = "black", linetype = 2)+
  theme_cowplot()+
  ylab(expression("D"[low]))+
  #annotate("text", x = 0, y = 29, label = "E", size = 5)+
  annotate("text", x = 0, y = 40, label = "D", size = 5)+
  xlab(expression("T"[0]))+
  theme(legend.position = "none")
Ts
```
Figure 3D N = `r nrow(res_plot5)`


```{r}

legendplot <- ggplot(res_plot6, aes(x = T0, y = Dlower))+
  geom_point(aes(shape = lifestage))+
  ylim(0,30)+
  xlim(0,30)+
  
  scale_shape_manual(values = c(1,16,8))+
  #scale_color_viridis_d()+
  geom_segment(x = 0, xend = 25, y = 0, yend = 25, col = "black", linetype = 2)+
  theme_cowplot()+
  ylab(expression("D"[low]))+
  annotate("text", x = 0, y = 29, label = "A", size = 5)+
  xlab(expression("T"[0]))+
  theme(legend.position = "right")
legend <- get_legend(legendplot)

```



#Figure 4 F P~S; P~D [Low]
```{r}
res_plot7 <- res %>% 
  filter(!is.na(Plower),  !is.na(Slower), !is.na(Dlower))


fitPlowsD <- lm(Plower ~ Dlower, data = res_plot7)
fitPlowsS <- lm(Plower ~ Slower, data = res_plot7)

shapiro.test(fitPlowsD$residuals)
shapiro.test(fitPlowsS$residuals)

summary(fitPlowsD)
```



```{r}

summary(fitPlowsS)
```

```{r}


Plow <- ggplot(res, aes(x = Slower, y = Plower))+
  geom_point(aes(shape = lifestage), col = "black")+
  geom_point(data = res_plot7, aes(x = Dlower, y = Plower, shape = lifestage), col = "grey")+
  ylim(0,40)+
  xlim(0,40)+
  geom_smooth(method = lm, se = F, col = "black")+
  geom_smooth(data = res_plot7, aes(x = Dlower, y = Plower), col = "grey", method = lm, se = F)+
  scale_shape_manual(values = c(1,16,8))+
  #scale_color_viridis_d()+
  #geom_segment(x = 5, xend = 30, y = 15, yend = 15, col = "black", linetype = 2)+
  theme_cowplot()+
  ylab(expression("P"[low]))+
   #annotate("text", x = 5, y = 30, label = "F", size = 5)+
   annotate("text", x = 1, y = 40, label = "E", size = 5)+
   annotate("text", x = 5, y = 5, label = expression("D"[low]), size = 5, col = "grey")+
   annotate("text", x = 5, y = 1, label = expression("S"[low]), size = 5, col = "black")+
  #xlab(expression("S(black)/D(grey)"[low]))+
  xlab(expression("S"[low]/"D"[low]))+ 
  theme(legend.position = "none")
Plow
```


#Figure 4 G P~S; P~D [Low]



```{r}
res_plot8 <- res %>% 
  filter(!is.na(Popt),  !is.na(maxSopt))

Popt <- ggplot(res_plot8, aes(x = maxSopt, y = Dopt))+
  geom_point(aes(shape = lifestage), col = "black")+
  geom_point(data = res_plot5, aes(x = maxSopt, y = Dopt, shape = lifestage), col = "grey")+
  ylim(0,40)+
  xlim(0,40)+
  geom_smooth(method = lm, se = F, col = "black")+
  geom_smooth(data = res_plot5, aes(x = maxSopt, y = Dopt), col = "grey", method = lm, se = F)+
  scale_shape_manual(values = c(1,16,8))+
  #scale_color_viridis_d()+
  #geom_segment(x = 5, xend = 30, y = 15, yend = 15, col = "black", linetype = 2)+
  theme_cowplot()+
  ylab(expression("P"[opt]))+
   annotate("text", x = 1, y = 40, label = "G", size = 5)+
  xlab(expression("S(black)/D(grey)"[low]))+
  theme(legend.position = "none")
Popt
```


#Figure 4 F P~S; P~D [high]


```{r}
res_plot9 <- res %>% 
  filter(!is.na(Pupper),  !is.na(Supper), !is.na(Dupper))

fitPhD <- lm(Pupper ~ Dupper, data = res_plot9)
fitPhS <- lm(Pupper ~ Supper, data = res_plot9)

shapiro.test(fitPhD$residuals)
hist(fitPhD$residuals)
hist(fitPhS$residuals)
shapiro.test(fitPhS$residuals)

summary(fitPhD)
summary(fitPhS)

```


```{r}
res_plot9 <- res %>% 
  filter(!is.na(Pupper),  !is.na(Supper))

Pupper <- ggplot(res_plot9, aes(x = Supper, y = Pupper))+
  geom_point(aes(shape = lifestage), col = "black")+
  geom_point(data = res_plot9, aes(x = Dupper, y = Pupper, shape = lifestage), col = "grey", alpha = 0.5)+
  ylim(0,40)+
  xlim(0,40)+
  geom_smooth(method = lm, se = F, col = "black")+
  geom_smooth(data = res_plot9, aes(x = Dupper, y = Pupper), col = "grey", method = lm, se = F)+
  scale_shape_manual(values = c(1,16,8))+
  #scale_color_viridis_d()+
  #geom_segment(x = 5, xend = 30, y = 15, yend = 15, col = "black", linetype = 2)+
  theme_cowplot()+
  ylab(expression("P"[high]))+
  #annotate("text", x = 5, y = 37, label = "H", size = 5)+
   annotate("text", x = 1, y = 40, label = "F", size = 5)+
  annotate("text", x = 5, y = 5, label = expression("D"[high]), size = 5, col = "grey")+
   annotate("text", x = 5, y = 1, label = expression("S"[high]), size = 5, col = "black")+
  #xlab(expression("S(black)/D(grey)"[low]))+
  xlab(expression("S"[high]/"D"[high]))+ 
 # xlab(expression("S(black)/D(grey)"[high]))+
  theme(legend.position = c(0.1,0.4))
Pupper
```



















```{r}

detach("package:cowplot", unload = TRUE)
library(cowplot)
#plot_grid(optima, high, width,lows, ncol = 2 )
plot_grid(lows, optima, high, width,Ts,legend, ncol = 3 )
plot_grid(lows, optima, high, width,Ts,legend, Plow, Popt, Pupper, ncol = 3 )
plot_grid(lows, high, width,Ts, Plow,  Pupper,legend, ncol = 2 )
plot_grid(lows, Ts,Plow,  Phigh, ncol = 2 )

```
