---
title: "Hypotheses plots"
output: html_notebook
---

Performance. We calculated performance as a function of development rate and surivival. 
We expect development rate to have steeper slopes at tropical than at temperate latitudes due to "higher is better".

```{r}
# Gaussian --------------------------------------------------------------------------------------
# a is the height of the curve
# b is the position of the center of the peak
# c is the standard deviation
Gaus <- function(x, a, b, c) {
  output <- a * exp(1)^(-0.5 * (abs(x - b)/ c)^2)
  print(output)
}

```



Survival curve bounds development curve. However, Sopt is often smaller than Dopt. What factors influence variation in the distance between Dopt and Sopt?

```{r}
di <- tibble(Temperature = seq(14,34, by = 1),
             survival = c(0,0.1, 0.5, 0.75, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 0.75,0.5,0.1,0), 
             dr = Gaus(Temperature, 1, 30, 3), 
             drTrop = Gaus(Temperature, 1, 33, 6),
             performance = dr * survival, 
             dr2 = Gaus(Temperature, 1, 33, 4), 
             performance2 = dr2 * survival, 
             performance3 = drTrop * survival)
Popt <-   filter(di, performance == max(performance, na.rm = T))[["Temperature"]]
maxP <-   filter(di, performance == max(performance, na.rm = T))[["performance"]]
Plower <- filter(di, performance >= maxP/2)[1, 1][[1]]
Pupper <- max(filter(di, performance <= maxP/2)[, 1])

triangle <- data.frame(group = c(1,1,1), polygon.x = c(26.5,Popt,31.64), polygon.y = c(0.5,maxP,0.5))


ggplot(di, aes(x = Temperature, y = survival))+
  geom_line(col = "gray")+
  geom_line(aes(x = Temperature, y = dr), col = "grey")+
  geom_line(aes(x = Temperature, y = performance))+
  geom_hline(yintercept =  0.5, linetype = 2)+
  #geom_area(mapping = aes(Temperature = ifelse(Temperature >26.5 & Temperature< 31.64 & performance > 0.5, Temperature, 0)), fill = "red") +
  #geom_polygon(data = triangle, aes(x = polygon.x, y = polygon.y, group = group), alpha = 0.3)+
  ylab("performance, survival, dr")+
  theme_cowplot()
```


```{r}
Popt <-   filter(di, performance2 == max(performance2, na.rm = T))[["Temperature"]]
maxP <-   filter(di, performance2 == max(performance2, na.rm = T))[["performance2"]]
Plower <- filter(di, performance2 >= maxP/2)[1, 1][[1]]
Pupper <- max(filter(di, performance2 <= maxP/2)[, 1])

triangle2 <- data.frame(group = c(1,1,1), polygon.x = c(28.3,Popt,31.9), polygon.y = c(0.5,maxP,0.5))

ggplot(di, aes(x = Temperature, y = survival))+
  geom_line(col = "gray")+
  geom_line(aes(x = Temperature, y = dr2), col = "grey")+
  geom_line(aes(x = Temperature, y = performance2))+
  geom_hline(yintercept =  0.5, linetype = 2)+
  #geom_area(mapping = aes(Temperature = ifelse(Temperature >26.5 & Temperature< 31.64 & performance2 > 0.5, Temperature, 0)), fill = "red") +
  #geom_polygon(data = triangle2, aes(x = polygon.x, y = polygon.y, group = group), alpha = 0.3)+
  ylab("performance, survival, dr")+
  theme_cowplot()
```

```{r}
ggplot(di, aes(x = Temperature, y = survival))+
  geom_line(col = "gray")+
  geom_line(aes(x = Temperature, y = drTrop), col = "grey")+
  geom_line(aes(x = Temperature, y = performance3))+
  geom_hline(yintercept =  0.5, linetype = 2)+
  #geom_area(mapping = aes(Temperature = ifelse(Temperature >26.5 & Temperature< 31.64 & performance2 > 0.5, Temperature, 0)), fill = "red") +
  #geom_polygon(data = triangle2, aes(x = polygon.x, y = polygon.y, group = group), alpha = 0.3)+
  ylab("performance, survival, dr")+
  theme_cowplot()
```





```{r}
survival <- tibble(Temperature = c(12,13,14,  15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25, 26,27,28,29, 31,32,33,34,35,36,37,38,39), 
            Stemperate =  c(0,0.1, 0.5, 0.75, 1, 1,   1, 1,   1, 1,   1, 1, 1, 1, 1, 1, 1, 1, 0.75,0.5,0.1,0,NA,NA,NA, NA,NA), 
            Stropical =    c(NA,NA,NA,0, 0.1, 0.5, 0.75, 1,   1, 1,   1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 0.75,0.5,0.1,0, NA, NA), 
            Ss =    c(NA,NA, NA,NA,NA, NA, NA, NA, NA,0, 0.1, 0.5, 0.75,  1, 1, 1, 1, 1, 1, 1, 1, 0.75,0.5,0.1,0, NA,NA), 
            Short =    Gaus(Temperature, 1, 25, 5)) 

ggplot(survival, aes(x = Temperature, y = Stemperate))+
  geom_line(linetype = 2)+
  geom_line(data = survival, mapping= aes(x = Temperature, y = Stropical),linetype = 4)+
  geom_line(data = survival, mapping= aes(x = Temperature, y = Short), linetype = 3)+
  geom_line(data = survival, mapping= aes(x = Temperature, y = Ss), linetype = 1)+
  theme_cowplot()

```

For tropical (solid line), the valid development range is from 17 to 36. For temperate (dashed line), the valid development range is from 14 to 33


```{r}


Gaus <- function(x, a, b, c) {
  output <- a * exp(1)^(-0.5 * (abs(x - b)/ c)^2)
  print(output)
}

x <- seq(10,40,by = 1)
tropical <- tibble(Temperature = c(10,  14,  15, 20, 25, 33,32,33,37), 
                   survival =     c(0, 0.5, 0.75, 1, 1,  1, 0.75,0.5,0 )) 
# Choose the mean as 2.5 and standard deviation as 2. 

first <- tibble(ratetemp = Gaus(x, 1, 30, 5), 
                temp = x,
                ratetrop1 = Gaus(x, 1, 33, 5),
                ratetrop = ifelse(temp > 32, (ratetemp + ratetrop1 + ratetemp)/3, ratetrop1))

ggplot(first, aes(x = temp, y = ratetemp))+
  geom_line()+
  geom_line(aes(x = temp, y = ratetrop1), col = "orange")
```


```{r}
frankencurves <- tibble(Temperature = first$temp,
                        Dtemperate = ifelse(Temperature < 34, first$ratetemp, NA), 
                        Dtropical = ifelse(Temperature < 37, first$ratetrop1, NA))




```

```{r}
tropiOpt <- filter(frankencurves, Dtropical == max(Dtropical, na.rm = T))[[1,1]]
tropiRate <- filter(frankencurves, Dtropical == max(Dtropical, na.rm = T))[[1,3]]
tempeOpt <- filter(frankencurves, Dtemperate == max(Dtemperate, na.rm = T))[[1,1]]
tempeRate <- filter(frankencurves, Dtemperate == max(Dtemperate, na.rm = T))[[1,2]]
ggplot(frankencurves, aes(x = Temperature, y = Dtemperate))+
  geom_line(linetype = 2)+
  geom_line(data = frankencurves, mapping = aes(x = Temperature, y = Dtropical), linetype = 1)+
  #geom_line(data = frankencurves, mapping = aes(x = Temperature, y = temperate), linetype = 4, col = grey, alpha = 0.5)+
  #geom_line(data = frankencurves, mapping = aes(x = Temperature, y = tropical), linetype = 3, col = grey, alpha = 0.5)+
  #geom_line(data = first, mapping = aes(x = temp, y = ratetrop1), linetype = 3)+
  #geom_segment(x = tropiOpt, xend = tropiOpt, y = 0, yend = tropiRate)+
  #geom_segment(x = tempeOpt, xend = tempeOpt, y = 0, yend = tempeRate, linetype = 2)+
  theme_cowplot()+
  ylab("Standardized development rate")+
  xlab("Temperature")


```


```{r}


Temperature <-  seq(from = 0, to = 40, by = 1) 

hypot <- merge(survival, frankencurves, by = "Temperature")

hyp <- tibble(temp = hypot$Temperature, 
              S1 = hypot$Stropical, 
              S2 = hypot$Stemperate, 
              SS = hypot$Short,
              Ss = hypot$Ss,
              D1 = ifelse(is.na(S1), NA, ifelse(S1 > 0, hypot$Dtropical, NA)),
              D2 = ifelse(is.na(S2), NA, ifelse(S2 > 0, hypot$Dtemperate, NA)),
              
              P1 = S1 * D1, 
              P2 = S2 * D2, 
              P1s = SS * D1, 
              P2s = SS * D2, 
              P1ss = Ss *D1)
  
  
  
ggplot(hyp, aes(x = temp, y = S1))+
  geom_line(col = "grey")+
  geom_line(aes(x = temp, y = D1), linetype = 1, col = "grey")+
  geom_line(aes(x = temp, y = P1), linetype = 1, col = "black")+
  geom_line(aes(x = temp, y = D2), linetype = 2, col = "grey")+
  geom_line(aes(x = temp, y = P2), linetype = 2, col = "black")+
  geom_line(aes(x = temp, y = S2), linetype = 2, col = "grey")+
  #geom_line(aes(x = temp, y = S2), linetype = 1, col = "red")+
  #geom_line(aes(x = temp, y = D2), linetype = 1, col = "orange")+
  #geom_line(aes(x = temp, y = P2), linetype = 1, col = "orange")+
  theme_cowplot()
  


```

```{r}
ggplot(hyp, aes(x = temp, y = S2))+
  geom_line(col = "grey", linetype = 2)+
  geom_line(aes(x = temp, y = D2), linetype = 2, col = "grey")+
  geom_line(aes(x = temp, y = P2), linetype = 2, col = "black")+
  
  theme_cowplot()
```


```{r}
ggplot(hyp, aes(x = temp, y = Ss))+
  geom_line(col = "grey")+
  geom_line(aes(x = temp, y = D1), linetype = 1, col = "grey")+
  #geom_line(aes(x = temp, y = P1s), linetype = 2, col = "black")+
  geom_line(aes(x = temp, y = P1ss), linetype = 1, col = "black")+
  theme_cowplot()
```




```{r}
ggplot(hyp, aes(x = temp, y = S1))+
  geom_line(col = "grey")+
  geom_line(aes(x = temp, y = D1), linetype = 1, col = "grey")+
  #geom_line(aes(x = temp, y = P1s), linetype = 2, col = "black")+
  geom_line(aes(x = temp, y = P1), linetype = 1, col = "black")+
  theme_cowplot()

```

Is there a trade-off between cold and heat tolerance?
Cold tolerance = topt - 

```{r}

```

