---
title: "Hypotheses plots"
output: html_notebook
---
Figure 1. Thermal tolerance indicators
Figure 2. Hypotheses


```{r}
# load packages
library(tidyverse)
library(readxl)
library(tidyverse)
library(readxl)
library(cowplot)
library(psych) # to get geometric mean
library(lme4)
library(car)
library("dotwhisker")  # to make coefficient plots
library(sjPlot) #prediction plots
library(sjmisc)
library(broom)
```


```{r}
# Function to make toy data
# Gaussian --------------------------------------------------------------------------------------
# a is the height of the curve
# b is the position of the center of the peak
# c is the standard deviation
Gaus <- function(x, a, b, c) {
  output <- a * exp(1)^(-0.5 * (abs(x - b)/ c)^2)
  print(output)
}

```


# Figure 1. Thermal tolerance indicators and their relationships (definitions)

```{r}
# toy data
temperature <- seq(10,40, by = 1)
A_survival <- c(0, 0.25, 0.5, 0.75, 0.89,
                1, 1, 1, 1, 1, 1, 1, 1, 1, 1,
                1, 1, 1, 1, 1, 1, 1, 1, 1, 1,
                0.89, 0.75,0.5,0.25,0,NA)

A_survival <- c(0, 0.25, 0.5, 0.75, 0.89,
                1, 1, 1, 1, 1, 1, 1, 1, 1, 1,
                1, 1, 1, 1, 1, 1, 1,
                0.89, 0.75,0.5,0.25,0,NA, NA, NA, NA)

A_drGaus <- Gaus(temperature, 1, 30, 3)
# This code is get the values of the vector above, but separated by commas, so the curve can be edited:
paste(as.character(A_drGaus), sep="' '", collapse=", ")
A_drtrop <- c(NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, 
              NA, NA, NA, NA,
              0.00386592013947281, 0.0111089965382423, 0.0285655007845504, 0.0657285286165305, 0.135335283236613,               0.249352208777296, 0.411112290507187, 0.606530659712633, 0.800737402916808, 0.945959468906766,
              1, 0.945959468906766, 0.800737402916808, 0.606530659712633, 0, 
              NA, NA)

A_drtemp <- c(NA, NA, NA, NA, NA, 
              NA, NA, NA, NA, NA,
              0.00386592013947281, 0.0111089965382423, 0.0285655007845504, 0.0657285286165305, 0.135335283236613,               0.249352208777296, 0.411112290507187, 0.606530659712633, 0.800737402916808, 0.945959468906766, 
              1, 0.945959468906766, 0.800737402916808, 0.606530659712633, 0, 
              
              NA, NA, NA, NA, NA, NA)
B_dr <- c( NA,0.0168798841487899, 0.0251304889852744, 0.036658041953378, 0.0523931410698256, 0.0733696513683829, 0.100668899772898, 0.135335283236613, 0.178263979585048, 0.230066298993809, 0.290923807046148, 0.360447788597821, 0.437564737700269, 0.520450121020702, 0.606530659712633, 0.692569324205198, 0.774837428883249, 0.849365816568312, 0.912254076828545, 0.960005441285478, 0.989847803378734, 1,0.989847803378734, 0.960005441285478,0.90, 0.800737402916808, 0.606530659712633, 0, 
              NA, NA, NA)
Figure1 <- tibble(temperature = temperature, 
                  survival = A_survival, 
                  tropical = B_dr, 
                  temperate = A_drtemp, 
                  P_tropical = survival * tropical, 
                  P_temperate = survival * temperate, 
                  skew0 = A_drGaus)
Dopt <- filter(Figure1, tropical == 1)
Dlow <- filter(Figure1, tropical >= 0.5) [1,]
Dhigh <- filter(Figure1, tropical >= 0.5) [14,]
Dinterval <- rbind(Dlow, Dhigh)

```

Panel A (development rate)

```{r}
devrate <- ggplot(Figure1, aes(x = temperature, y = tropical))+
  geom_line(linetype = 3, size = 2)+
  
  geom_text(Dopt, mapping = aes(x = 30, y = 0.55, label = "Dwidth"), size = 5)+
  geom_text(Dopt, mapping = aes(x = 10, y = 0.95, label = "A"), size = 5)+
  
  geom_point(Dopt, mapping = aes(x = temperature, y = tropical), size = 4, shape = 1)+
  #geom_point(Dlow, mapping = aes(x = temperature, y = 0.5), size = 4, shape = 1)+
  #geom_point(Dhigh, mapping = aes(x = temperature, y = 0.5), size = 4, shape = 1)+
  geom_point(Dhigh, mapping = aes(x = 14.5, y = 0), size = 4, shape = 17)+
  
  geom_segment(x = Dlow$temperature, xend = Dhigh$temperature, y = 0.5, yend = 0.5, linetype = 3)+
  geom_segment(x = Dopt$temperature, xend = 14.5, y = 1, yend = 0)+
  
  theme_cowplot()+
  ylab("Development rate")+
  xlab("Temperature")+
  ylim(0.0,1)+
  theme(axis.text = element_text(size = 16))+
  #theme(axis.text.x= c("T0", "D-low", "D-opt", "D-high"))+
  scale_x_continuous(breaks = c(14.5, 23, 31, 36), labels = c(expression("T"[0]), expression("D"[low]), expression("D"[opt]), expression("D"[high])))

devrate
```


Panel B: Survival

```{r}
surv <- ggplot(Figure1, aes(x = temperature, y = tropical))+
  
  geom_line(Figure1, mapping = aes(x = temperature, y = survival), size = 2, linetype = 2)+
  
  geom_text(Dopt, mapping = aes(x = 24, y = 0.55, label = "Swidth"), size = 5)+
  geom_text(Dopt, mapping = aes(x = 10, y = 0.95, label = "B"), size = 5)+
  
  #geom_point(Figure1, mapping = aes(x = 12, y = 0.5), size = 4)+
  #geom_point(Figure1, mapping = aes(x = 34, y = 0.5), size = 4)+
  geom_point(Figure1, mapping = aes(x = 31, y = 1), size = 4)+
  
  geom_segment(x = 12, xend = 34, y = 0.5, yend = 0.5, linetype = 3)+
  #geom_segment(x = Dopt$temperature, xend = 14.5, y = 1, yend = 0)+
  #geom_point(Dopt, mapping = aes(x = temperature, y = 0), size = 4)+
  theme_cowplot()+
  ylab("Survival")+
  xlab("Temperature")+
  #ylim(0.04,1)+
  theme(axis.text = element_text(size = 16))+
  #theme(axis.text.x=element_blank())+
  #theme(axis.text.x= c("T0", "D-low", "D-opt", "D-high"))+
  scale_x_continuous(breaks = c(12, 31, 34), labels = c(expression("S"[low]),expression("S"[opt]), expression("S"[high])))
surv
```

Performance. 
We calculated performance as a function of development rate and surivival. If development rate has a fixed function, P-width (the length of the temperature interval that allows for 50% efficiency in dr and survival) will depend on the relationship between Dopt and Sopt. 

Panel C. Performance when Dopt < Sopt

```{r}
Figure1$s2 <- c(0, 0.25, 0.5, 0.75, 0.89,
                1, 1, 1, 1, 1, 1, 1, 1, 1, 1,
                1, 1, 1, 1, 1, 1, 1,1,1,1,1,
                0.89, 0.75,0.5,0.25,0)

Figure1$P2 <- Figure1$s2 * Figure1$tropical

per1 <- ggplot(Figure1, aes(x = temperature, y = tropical))+
  geom_line(Figure1, mapping = aes(x = temperature, y = P2), size = 3, col = "gray")+
  geom_line(linetype = 3, size = 1)+
  geom_line(Figure1, mapping = aes(x = temperature, y = s2), size = 1, linetype = 2)+
  
  geom_point(Figure1, mapping = aes(x = 35, y = 1), size = 4, col = "black")+
  geom_point(Figure1, mapping = aes(x = 31, y = 1), size = 4, col = "grey")+
  geom_point(Figure1, mapping = aes(x = 31, y = 1), size = 4, col = "black", shape = 1)+
  geom_segment(x = 23, xend = 36, y = 0.5, yend = 0.5, linetype = 3)+
  
  geom_text(Dopt, mapping = aes(x = 29, y = 0.55, label = "Pwidth"), size = 5)+
  geom_text(Dopt, mapping = aes(x = 10, y = 0.95, label = "C"), size = 5)+
  annotate("text", x = 31, y = 0.01, label = expression("D"[opt]))+
  annotate("text", x = 35, y = 0.01, label = expression("S"[opt]))+
  
  theme_cowplot()+
  ylab("Performance")+
  xlab("Temperature")+
  #ylim(0.04,1)+
  theme(axis.text = element_text(size = 16))+
  #theme(axis.text.x=element_blank())+
  #theme(axis.text.x= c("T0", "D-low", "D-opt", "D-high"))+
  scale_x_continuous(breaks = c(23,31, 36), labels = c(expression("P"[low]), expression("P"[opt]), expression("P"[high])))

per1

```




Panel D. Performance when Dopt = Popt

```{r}
per2 <- ggplot(Figure1, aes(x = temperature, y = tropical))+
  geom_line(Figure1, mapping = aes(x = temperature, y = P_tropical), size = 3, col = "gray")+
  geom_line(linetype = 3, size = 1)+
  geom_line(Figure1, mapping = aes(x = temperature, y = survival), size = 1, linetype = 2)+
  
  geom_text(Dopt, mapping = aes(x = 10, y = 0.95, label = "D"), size = 5)+
  geom_text(Dopt, mapping = aes(x = 29, y = 0.55, label = "Pwidth"), size = 5)+
  annotate("text", x = 31, y = 0.06, label = expression("D"[opt]))+
  annotate("text", x = 31, y = 0.01, label = expression("S"[opt]))+
  geom_text(Dopt, mapping = aes(x = 29, y = 0.55, label = "Pwidth"), size = 5)+
  
  #geom_point(Figure1, mapping = aes(x = 33.5, y = 0.5), size = 5, col = "grey")+
  #geom_point(Figure1, mapping = aes(x = 23, y = 0.5), size = 5, col = "grey")+
  geom_point(Figure1, mapping = aes(x = 31, y = 1), size = 5, col = "grey")+
  geom_point(Figure1, mapping = aes(x = 31, y = 1), size = 4, col = "black", shape = 1)+
  
  geom_segment(x = 23, xend = 33.5, y = 0.5, yend = 0.5, linetype = 3)+
  
  theme_cowplot()+
  ylab("Performance")+
  xlab("Temperature")+
  theme(axis.text = element_text(size = 12))+
  scale_x_continuous(breaks = c(23, 31, 33.5), labels = c(expression("P"[low]),expression("P"[opt]),expression("P"[high])))
per2

```



Panel E. Performance when Dopt > Sopt
```{r}

Figure1$s3 <-  c(0, 0.25, 0.5, 0.75, 0.89,
                1, 1, 1, 1, 1, 1, 1, 1, 1, 1,
                1, 1, 1,  
                0.89, 0.75,0.5,0.25,0,NA, NA, NA, NA, 
                NA, NA, NA, NA)
Figure1$d3 <- ifelse(Figure1$temperature <= 32, Figure1$tropical, NA)

Figure1$P3 <- Figure1$s3 *Figure1$tropical
per3 <- ggplot(Figure1, aes(x = temperature, y = d3))+
  geom_line(Figure1, mapping = aes(x = temperature, y = P3), size = 3, col = "grey")+
  geom_line(linetype = 3, size = 1, alpha = 1)+
  geom_line(Figure1, mapping = aes(x = temperature, y = s3), size = 1, linetype = 2, alpha = 1)+
  
  geom_point(Figure1, mapping = aes(x = 27, y = 1), size = 4, col = "black", shape = 16)+
  geom_point(Figure1, mapping = aes(x = 27, y = 0.849), size = 5, col = "grey", shape = 16)+
  geom_point(Figure1, mapping = aes(x = 31, y = 1), size = 4, col = "black", shape = 1)+
  
  geom_segment(x = 23, xend = 30, y = 0.5, yend = 0.5, linetype = 3)+
  geom_text(Dopt, mapping = aes(x = 27, y = 0.55, label = "Pwidth"), size = 5)+
  geom_text(Dopt, mapping = aes(x = 10, y = 0.95, label = "E"), size = 5)+
  
  annotate("text", x = 31, y = 0.01, label = expression("D"[opt]))+
  annotate("text", x = 27, y = 0.01, label = expression("S"[opt]))+
  theme_cowplot()+
  theme(axis.text = element_text(size = 16))+
  
  ylab("Performance")+
  xlab("Temperature")+
 # xlim(10,30)+
  
  #theme(axis.text.x=element_blank())+
  #theme(axis.text.x= c("T0", "D-low", "D-opt", "D-high"))+
  scale_x_continuous(limits = c(10,33),breaks = c(23, 27, 30), labels = c(expression("P"[low]), expression("P"[opt]),  expression("P"[high])))
per3
```


```{r}
# run this in .R document to get proper figure (no overlaping letters)

grid.arrange(
  arrangeGrob(
    devrate, 
    arrangeGrob(surv, nrow=1),
    ncol=2 ,widths=c(1.5,1.5)),
  arrangeGrob(per1, per2 ,per3 ,ncol=3, widths=rep(1,3)),
  nrow=2)
```


# Figure 2. Hypotheses

```{r}
hyp <- tibble(Dopt = rnorm(50, mean = 30, sd = 2), 
              Sopt = Dopt,
              Dhigh = rnorm(50, mean = 34, sd = 1), 
              Shigh = Dhigh -2, 
              Dwidth = rnorm(50, mean = 7, sd = 2),
              Swidth = Dwidth *2,
              latopt = Sopt/2)

opt <- ggplot(hyp, aes(x = Sopt, y = Dopt))+
  geom_line(size = 2)+
  
  ylab(expression("D"[opt]))+
  xlab(expression("S"[opt]))+
  geom_text(Dopt, mapping = aes(x = 24, y = 34.5, label = "A"), size = 5)+
  theme_cowplot()+
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), 
        axis.text.y = element_blank(), axis.ticks.y = element_blank())
opt

```


```{r}
high <- ggplot(hyp, aes(x = Shigh, y = Dhigh))+
  geom_line(size = 2)+
  
  ylab(expression("D"[high]))+
  xlab(expression("S"[high]))+
  geom_text(Dopt, mapping = aes(x = 28.5, y = 37, label = "B"), size = 5)+
  theme_cowplot()+
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), 
        axis.text.y = element_blank(), axis.ticks.y = element_blank())
high
```


```{r}
width <- ggplot(hyp, aes(x = Swidth, y = Dwidth))+
  geom_hline(yintercept = 4,size = 2)+
  theme_cowplot()+
  
  ylab(expression("D"[width]))+
  xlab(expression("S"[width]))+
  
  annotate("text", x = 0, y = 8, label = "C", size = 5)+
  ylim(0,8)+
  xlim(0, 8)+
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), 
        axis.text.y = element_blank(), axis.ticks.y = element_blank())
width
```



```{r}
widthlat <- ggplot(hyp, aes(x = Shigh, y = Dhigh -1, col = "larva"))+
  geom_line(data = hyp, mapping = aes(x = Shigh, y = Dhigh, col = "egg"), size = 1.3)+
  geom_line(data = hyp, mapping = aes(x = Shigh, y = Dhigh -.2, col = "pupa"), size = 1.3)+
  geom_line(size = 1.3)+
  #scale_color_discrete()+
  scale_color_manual(values = c("grey", "grey50", "black"))+
  theme_cowplot()+
  annotate("text", x = 28.5, y = 37, label = "D", size = 5)+
  ylab(expression(P[width]))+
  xlab("Absolute latitude")+
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), 
        axis.text.y = element_blank(), axis.ticks.y = element_blank(), 
        legend.title = element_blank(), legend.position = c(0.2,0.8))
widthlat
```




```{r}


lab <- expression(paste( S[opt],", ", D[opt],", ", P[opt]))
threelat <-  ggplot(hyp, aes(x = Dopt, y = Sopt))+
  geom_point(col = "white")+
  geom_segment(x = 26, xend = 34, y = 34-1.5, yend = 26-1.5, size = 1.3, col = "grey50")+
  geom_segment(x = 26, xend = 34, y = 34+1.5, yend = 26+1.5, size = 1.3, col = "black")+
  geom_segment(x = 26, xend = 34, y = 34+1, yend = 26+1, size = 1.3, col = "grey")+
  #geom_segment(x = 26, xend = 34, y = 34, yend = 26, size = 2, linetype = 1)+
  #geom_segment(x = 2, xend = 31, y = 0, yend = 1, col = "black")+
  theme_cowplot()+
  annotate("text", x = 25, y = 35, label = "E", size = 5)+
  ylab(lab)+
  xlab("Absolute latitude")+
  ylim(25, 35)+
  xlim(25, 35)+
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), 
     axis.text.y = element_blank(), axis.ticks.y = element_blank())


threelat
```

```{r}
skewhy <- ggplot(hyp, aes(x = Dopt, y = Sopt))+
  
  geom_segment(x = 26, xend = 34, y = 34, yend = 26, size = 1.3, col = "black")+
  theme_cowplot()+
  annotate("text", x = 25, y = 35, label = "E", size = 5)+
  ylab("Skew")+
  xlab(expression("P"[high]))+
  ylim(25, 35)+
  xlim(25, 35)+
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), 
     axis.text.y = element_blank(), axis.ticks.y = element_blank())
skewhy
```

```{r}
# small Dwidth, skew = 0
skew <- tibble(temperature = seq(0,20, by = 0.1), 
               ssmall = Gaus(temperature, 1, 10, 2), #2
               swide = Gaus(temperature, 1, 10, 5), # 4
               wide_r = Gaus(temperature, 1, 15, 0.5), 
               wide_l = Gaus(temperature, 1, 15, 8), 
               wide = ifelse(temperature <= 15, wide_l, wide_r), 
               small_r = Gaus(temperature, 1, 15, 0.5), 
               small_l = Gaus(temperature, 1, 15, 3.25), 
               small = ifelse(temperature <= 15, small_l, small_r))
inset1 <- ggplot(skew, aes(x = temperature, y = ssmall))+
  geom_line(size = 1, col = "grey")+
 
  geom_segment(x = 8, xend = 12, y = 0.5, yend = 0.5, col = "black", size = 1, linetype = 3)+ # opt line
  #geom_point(x = 31, y = 1, col = "black")+ # opt line
  #geom_segment(x = 23, xend = 31, y = 0.5, yend = 0.5,linetype = 1, col = "grey35", size = 2)+ # cold tolerance (9)
  #geom_segment(x = 31, xend = 36, y = 0.5, yend = 0.5,linetype = 1, col = "lightgrey", size = 2)+ # heat tolerance (6)
  #geom_segment(x = 23, xend = 36, y = 0.5, yend = 0.5, linetype = 3)+
  
  #geom_segment(x = 12, xend = 20, y = 0.9, yend = 0.9,linetype = 2)+
  #geom_segment(x = 12, xend = 17, y = 0.8, yend = 0.8,linetype = 3)+
  #geom_segment(x = 17, xend = 31-11, y = 0.7, yend = 0.7,linetype = 1, size = 3)+
  
  #geom_segment(x = 15, xend = 20, y = 0.8, yend = 0.8,linetype = 3)+
  #geom_segment(x = 12, xend = 15, y = 0.7, yend = 0.7,linetype = 1, size = 3)+
  
  #annotate("text", x = 10, y = 0.4, label = "Dwidth = x\nskew = 0")+
  #annotate("text", x = 34, y = 0.5, label = "heat \ntolerance")+
  #annotate("text", x = 14, y = 0.74, label = "skew = cold tolerance/ \n        Pwidth")+
  ylim(0.02, 1)+
  
  theme_cowplot()+
   ylab("")+
   xlab("")+
  theme(legend.position = "none")+
   theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), 
     axis.text.y = element_blank(), axis.ticks.y = element_blank(), 
     axis.line.x = element_blank(), axis.line.y = element_blank())

inset1

```

```{r}
# large Dwidth, skew = 0
inset2 <- ggplot(skew, aes(x = temperature, y = swide))+
  geom_line(size = 1, col = "grey")+
 
  geom_segment(x = 5, xend = 15, y = 0.5, yend = 0.5, col = "black", size = 1, linetype = 3)+ # opt line
  #geom_point(x = 31, y = 1, col = "black")+ # opt line
  #geom_segment(x = 23, xend = 31, y = 0.5, yend = 0.5,linetype = 1, col = "grey35", size = 2)+ # cold tolerance (9)
  #geom_segment(x = 31, xend = 36, y = 0.5, yend = 0.5,linetype = 1, col = "lightgrey", size = 2)+ # heat tolerance (6)
  #geom_segment(x = 23, xend = 36, y = 0.5, yend = 0.5, linetype = 3)+
  
  #geom_segment(x = 12, xend = 20, y = 0.9, yend = 0.9,linetype = 2)+
  #geom_segment(x = 12, xend = 17, y = 0.8, yend = 0.8,linetype = 3)+
  #geom_segment(x = 17, xend = 31-11, y = 0.7, yend = 0.7,linetype = 1, size = 3)+
  
  #geom_segment(x = 15, xend = 20, y = 0.8, yend = 0.8,linetype = 3)+
  #geom_segment(x = 12, xend = 15, y = 0.7, yend = 0.7,linetype = 1, size = 3)+
  
  #annotate("text", x = 10, y = 0.4, label = "Dwidth = 2x\nskew = 0")+
  #annotate("text", x = 34, y = 0.5, label = "heat \ntolerance")+
  #annotate("text", x = 14, y = 0.74, label = "skew = cold tolerance/ \n        Pwidth")+
 # ylim(0.01, 1)+
  theme_cowplot()+
  ylab("")+
  xlab("")+
  theme(legend.position = "none")+
  
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), 
     axis.text.y = element_blank(), axis.ticks.y = element_blank(), 
     axis.line.x = element_blank(), axis.line.y = element_blank())
inset2
```

```{r}
# large Dwidth, skew = 0.7
inset3 <- ggplot(skew, aes(x = temperature, y = wide))+
  geom_line(size = 1, col = "grey")+
 
  geom_segment(x = 6, xend = 15.5, y = 0.5, yend = 0.5, col = "black", size = 1, linetype = 3)+ # opt line
  #geom_point(x = 31, y = 1, col = "black")+ # opt line
  #geom_segment(x = 23, xend = 31, y = 0.5, yend = 0.5,linetype = 1, col = "grey35", size = 2)+ # cold tolerance (9)
  #geom_segment(x = 31, xend = 36, y = 0.5, yend = 0.5,linetype = 1, col = "lightgrey", size = 2)+ # heat tolerance (6)
  #geom_segment(x = 23, xend = 36, y = 0.5, yend = 0.5, linetype = 3)+
  
 # geom_segment(x = 15, xend = 15, y = 0, yend = 1,linetype = 1, size =2, col = "grey")+
  #geom_segment(x = 12, xend = 17, y = 0.8, yend = 0.8,linetype = 3)+
  #geom_segment(x = 17, xend = 31-11, y = 0.7, yend = 0.7,linetype = 1, size = 3)+
  
  #geom_segment(x = 15, xend = 20, y = 0.8, yend = 0.8,linetype = 3)+
  #geom_segment(x = 12, xend = 15, y = 0.7, yend = 0.7,linetype = 1, size = 3)+
  
  #annotate("text", x = 10, y = 0.4, label = "Dwidth = 2x\nskew = 0.7")+
  #annotate("text", x = 34, y = 0.5, label = "heat \ntolerance")+
  #annotate("text", x = 14, y = 0.74, label = "skew = cold tolerance/ \n        Pwidth")+
  ylim(0.01, 1)+
  theme_cowplot()+
  ylab("")+
  xlab("")+
  theme(legend.position = "none")+
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), 
     axis.text.y = element_blank(), axis.ticks.y = element_blank(), 
     axis.line.x = element_blank(), axis.line.y = element_blank())
inset3
```

```{r}
# small Dwidth, skew = 0.7
inset4 <- ggplot(skew, aes(x = temperature, y = small))+
  geom_line(size = 1, col = "grey")+
  geom_segment(x = 11.2, xend = 15.2, y = 0.5, yend = 0.5,linetype = 3, size =1, col = "black")+
  #geom_segment(x = 11.2, xend = 16.2, y = 0.5, yend = 0.5, col = "black", size = 1)+ # opt line
  #geom_point(x = 31, y = 1, col = "black")+ # opt line
  #geom_segment(x = 23, xend = 31, y = 0.5, yend = 0.5,linetype = 1, col = "grey35", size = 2)+ # cold tolerance (9)
  #geom_segment(x = 31, xend = 36, y = 0.5, yend = 0.5,linetype = 1, col = "lightgrey", size = 2)+ # heat tolerance (6)
  #geom_segment(x = 23, xend = 36, y = 0.5, yend = 0.5, linetype = 3)+
  
  #geom_segment(x = 12, xend = 20, y = 0.9, yend = 0.9,linetype = 2)+
  #geom_segment(x = 12, xend = 17, y = 0.8, yend = 0.8,linetype = 3)+
  #geom_segment(x = 17, xend = 31-11, y = 0.7, yend = 0.7,linetype = 1, size = 3)+
  
  #geom_segment(x = 15, xend = 20, y = 0.8, yend = 0.8,linetype = 3)+
  #geom_segment(x = 12, xend = 15, y = 0.7, yend = 0.7,linetype = 1, size = 3)+
  
  #annotate("text", x = 13.5, y = 0.4, label = "Dwidth = x\nskew = 0.7")+
  #annotate("text", x = 34, y = 0.5, label = "heat \ntolerance")+
  #annotate("text", x = 14, y = 0.74, label = "skew = cold tolerance/ \n        Pwidth")+
  ylim(0.01, 1)+
  theme_cowplot()+
  ylab("")+
  xlab("")+
  theme(legend.position = "none")+
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), 
     axis.text.y = element_blank(), axis.ticks.y = element_blank(), 
     axis.line.x = element_blank(), axis.line.y = element_blank())
  
inset4
```





```{r}
library(cowplot)
inside <- plot_grid(inset4, inset3, inset1, inset2, ncol = 2)
inside
```

```{r}
out <- tibble(skew = c(0, 0.7), 
              Dwidth = c(5, 10))
outside <- ggplot(out, aes(x = Dwidth, y = skew))+
  ylab(expression("Skew"))+
  
   xlab(expression(P[width]))+
  annotate("text", x = 0, y = 1, label = "F", size = 5)+
  theme_cowplot()+
  ylim(0,1)+
  xlim(0,10)+
  scale_x_continuous(limits = c(0,10),breaks = c( 2.5, 7.5), labels = c("small", "large"))+
  scale_y_continuous(limits = c(0,1),breaks = c( 0.25, 0.75), labels = c("0", "1"))
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), 
       axis.text.y = element_blank(), axis.ticks.y = element_blank())
outside
```

```{r}
panelf <- ggdraw(outside) +
  draw_plot(inset1, 0.05, 0.1, 0.5,.5) +
  draw_plot(inset2, 0.5, 0.1, 0.5,.5) +
  draw_plot(inset3, 0.45, 0.5, 0.5,.5) +
  draw_plot(inset4, 0, 0.5, 0.5,.5) +
  #draw_plot(inside, 0.5, 0.1, 0.45,.45) +
  #draw_plot(inside, 0.5, 0.1, 0.45,.45) +
  draw_plot_label(
    c("A", "B"),
    c(1, 0),
    c(0, 0),
    size = 12
  )

panelf
```


Run this in a regular .R file to prevent crowdine
```{r}
plot_grid(devrate, surv, per1, per2, per3, panelf, ncol = 3)
```



```{r}
plot_grid(opt, high, width,  widthlat,threelat,skewhy, ncol = 3)

```



```{r}
# small Dwidth, skew = 0
A_drGaus <- Gaus(temperature, 1, 30, 3)
Figure1$skew0 <- Gaus(temperature, 1, 30, 10)
inset <- ggplot(Figure1, aes(x = temperature, y = skew0))+
  geom_line(size = 3, col = "grey")+
  geom_segment(x = 30, xend = 30, y = 0, yend = 1, col = "black")+ # opt line
  geom_point(x = 31, y = 1, col = "black")+ # opt line
  geom_segment(x = 23, xend = 31, y = 0.5, yend = 0.5,linetype = 1, col = "grey35", size = 2)+ # cold tolerance (9)
  geom_segment(x = 31, xend = 36, y = 0.5, yend = 0.5,linetype = 1, col = "lightgrey", size = 2)+ # heat tolerance (6)
  geom_segment(x = 23, xend = 36, y = 0.5, yend = 0.5, linetype = 3)+
  
  #geom_segment(x = 12, xend = 20, y = 0.9, yend = 0.9,linetype = 2)+
  #geom_segment(x = 12, xend = 17, y = 0.8, yend = 0.8,linetype = 3)+
  #geom_segment(x = 17, xend = 31-11, y = 0.7, yend = 0.7,linetype = 1, size = 3)+
  
  #geom_segment(x = 15, xend = 20, y = 0.8, yend = 0.8,linetype = 3)+
  #geom_segment(x = 12, xend = 15, y = 0.7, yend = 0.7,linetype = 1, size = 3)+
  
  annotate("text", x = 27, y = 0.5, label = "cold \ntolerance")+
  annotate("text", x = 34, y = 0.5, label = "heat \ntolerance")+
  annotate("text", x = 14, y = 0.74, label = "skew = cold tolerance/ \n        Pwidth")+
  
  theme_cowplot()+
  ylab("Performance")+
  xlab("Temperature")+
  scale_x_continuous(breaks = c(23, 31, 36), labels = c(expression("P"[low]), expression("P"[opt]), expression("S"[opt])))
inset
```


Skew inset 1 (small skew)
```{r}
inset <- ggplot(Figure1, aes(x = temperature, y = P_tropical))+
  geom_line(size = 3, col = "grey")+
  geom_segment(x = 31, xend = 31, y = 0, yend = 1, col = "black")+ # opt line
  geom_point(x = 31, y = 1, col = "black")+ # opt line
  geom_segment(x = 23, xend = 31, y = 0.5, yend = 0.5,linetype = 1, col = "grey35", size = 2)+ # cold tolerance (9)
  geom_segment(x = 31, xend = 36, y = 0.5, yend = 0.5,linetype = 1, col = "lightgrey", size = 2)+ # heat tolerance (6)
  geom_segment(x = 23, xend = 36, y = 0.5, yend = 0.5, linetype = 3)+
  
  #geom_segment(x = 12, xend = 20, y = 0.9, yend = 0.9,linetype = 2)+
  #geom_segment(x = 12, xend = 17, y = 0.8, yend = 0.8,linetype = 3)+
  #geom_segment(x = 17, xend = 31-11, y = 0.7, yend = 0.7,linetype = 1, size = 3)+
  
  #geom_segment(x = 15, xend = 20, y = 0.8, yend = 0.8,linetype = 3)+
  #geom_segment(x = 12, xend = 15, y = 0.7, yend = 0.7,linetype = 1, size = 3)+
  
  annotate("text", x = 27, y = 0.5, label = "cold \ntolerance")+
  annotate("text", x = 34, y = 0.5, label = "heat \ntolerance")+
  annotate("text", x = 14, y = 0.74, label = "skew = cold tolerance/ \n        Pwidth")+
  
  theme_cowplot()+
  ylab("Performance")+
  xlab("Temperature")+
  scale_x_continuous(breaks = c(23, 31, 36), labels = c(expression("P"[low]), expression("P"[opt]), expression("S"[opt])))
inset
```


Large skew
```{r}
inset2 <- ggplot(Figure1, aes(x = temperature, y = P_tropical))+
  geom_line(size = 3, col = "grey")+
  geom_segment(x = 31, xend = 31, y = 0, yend = 1, col = "black")+ # opt line
  geom_point(x = 31, y = 1, col = "black")+ # opt line
  geom_segment(x = 23, xend = 31, y = 0.5, yend = 0.5,linetype = 1, col = "grey35", size = 2)+ # cold tolerance (9)
  geom_segment(x = 31, xend = 36, y = 0.5, yend = 0.5,linetype = 1, col = "lightgrey", size = 2)+ # heat tolerance (6)
  geom_segment(x = 23, xend = 36, y = 0.5, yend = 0.5, linetype = 3)+
  
  #geom_segment(x = 12, xend = 20, y = 0.9, yend = 0.9,linetype = 2)+
  #geom_segment(x = 12, xend = 17, y = 0.8, yend = 0.8,linetype = 3)+
  #geom_segment(x = 17, xend = 31-11, y = 0.7, yend = 0.7,linetype = 1, size = 3)+
  
  #geom_segment(x = 15, xend = 20, y = 0.8, yend = 0.8,linetype = 3)+
  #geom_segment(x = 12, xend = 15, y = 0.7, yend = 0.7,linetype = 1, size = 3)+
  
  annotate("text", x = 27, y = 0.5, label = "cold \ntolerance")+
  annotate("text", x = 34, y = 0.55, label = "heat")+
  #annotate("text", x = 16, y = 0.74, label = "skew = cold tolerance/ \n        Pwidth")+
  
  theme_cowplot()+
  ylab("Performance")+
  xlab("Temperature")+
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), 
        axis.text.y = element_blank(), axis.ticks.y = element_blank())
inset2
```


skew (proportion of tolerance that represents cold tolerance)
```{r}
skew1 <- ggplot(hyp, aes(x = Shigh, y = Dhigh))+
  geom_line(size = 2)+
  
  ylab(expression("skew"))+
  xlab(expression("Absolute latitude"))+
  annotate("text", x = 31, y = 35, label = "skew = cold tolerance/ \n        Pwidth")+
  theme_cowplot()+
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), 
       axis.text.y = element_blank(), axis.ticks.y = element_blank())
skew1
```

```{r}

skew <- ggdraw(skew1 + theme_half_open(12)) +
  draw_plot(inset2, 0.5, 0.1, 0.45,.45) +
  draw_plot_label(
    c("A", "B"),
    c(1, 0),
    c(0, 0),
    size = 12
  )

skew
```


```{r}
lifesta <- tibble(lifestage = c("egg", "larva", "pupa"), 
                  meanPw = c(10, 4, 10), 
                  min = c(8, 2, 8),
                  max = c(12, 6, 12))

life <- ggplot(lifesta, aes(x = lifestage, y = meanPw))+
  geom_pointrange(lifesta, mapping = aes(ymin = min, ymax= max))+
  theme_cowplot()+
  ylim(0,13)+
  annotate("text", x = 0.6, y = 13, label = "F", size = 5)+
  ylab(expression("P"[width]))+
  xlab("")+
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank(), axis.title.x = element_blank())
life  
```


```{r}
plot_grid(opt, high, width, threelat, widthlat,life, ncol = 3)

```

# Old code below:

Anna's plots

A. Survival is fixed, dr is adaptive with latitude


```{r}
FigureA <- tibble(temperature = temperature, 
                  survival = A_survival, 
                  tropical = A_drtrop, 
                  temperate = A_drtemp, 
                  P_tropical = survival * tropical, 
                  P_temperate = survival * temperate)

ggplot(FigureA, aes(x = temperature, y = survival))+
  geom_line(linetype = 2, size = 2)+
  #geom_line(aes(x = temperature, y = tropical, col = "tropical"))+
  geom_line(aes(x = temperature, y = P_temperate), col = "grey",  size = 2)+
  geom_line(aes(x = temperature, y = temperate), col = "black", linetype = 3, size = 2)+
  scale_color_viridis_d()+
  theme_cowplot()
```

```{r}
ggplot(FigureA, aes(x = temperature, y = survival))+
  geom_line(linetype = 2, size = 2)+
  
  geom_line(aes(x = temperature, y = P_tropical), col = "grey", size = 2)+
  geom_line(aes(x = temperature, y = tropical), col = "black", size = 2, linetype = 3)+
  scale_color_viridis_d()+
  theme_cowplot()
```

```{r}
# Figure B
B_drGaus <- Gaus(temperature, 1, 30, 7)
B_dr <- c( NA,0.0168798841487899, 0.0251304889852744, 0.036658041953378, 0.0523931410698256, 0.0733696513683829, 0.100668899772898, 0.135335283236613, 0.178263979585048, 0.230066298993809, 0.290923807046148, 0.360447788597821, 0.437564737700269, 0.520450121020702, 0.606530659712633, 0.692569324205198, 0.774837428883249, 0.849365816568312, 0.912254076828545, 0.960005441285478, 0.989847803378734, 1,0.989847803378734, 0.960005441285478,0.90, 0.800737402916808, 0.606530659712633, 0, 
              NA, NA, NA)
B_temperate <- c(0, 0.25, 0.5, 0.75, 0.89,
                1, 1, 1, 1, 1, 1, 1, 1, 1, 1,
                1, 
                0.89, 0.75,0.5,0.25,0,NA,
                NA, NA, NA, NA, NA, NA, NA, NA, NA)
B_tropical <- c(NA, NA, NA, NA, NA, NA, 
                NA, NA, NA,NA, 0, 0.25, 0.5, 0.75, 0.89,
                1, 1, 1, 1, 1, 1,1, 1, 1, 1, 1,
                0.89, 0.75,0.5,0.25,0)
FigureB <- tibble(temperature = temperature, 
                  dr= B_dr, 
                  temperate = B_temperate, 
                  tropical = B_tropical, 
                  P_temperate = dr * temperate, 
                  P_tropical = dr * tropical)

ggplot(FigureB, aes(x = temperature, y = dr))+
  geom_line(aes(x = temperature, y = P_tropical), col = "grey", size = 2)+
  geom_line(aes(x = temperature, y = tropical), col = "black", size = 2, linetype = 2)+
  geom_line(linetype = 3, size = 2)+
  
  
  scale_color_viridis_d()+
  theme_cowplot()
```

```{r}
ggplot(FigureB, aes(x = temperature, y = dr))+
  geom_line(aes(x = temperature, y = P_temperate), col = "grey", size = 2)+
  geom_line(aes(x = temperature, y = temperate), col = "black", size = 2, linetype = 2)+
  geom_line(linetype = 3, size = 2)+
  scale_color_viridis_d()+
  theme_cowplot()
```
Figure C
```{r}
dr_temperateC <- c(NA, NA, NA, NA,
              0.00386592013947281, 0.0111089965382423, 0.0285655007845504, 0.0657285286165305, 0.135335283236613,               0.249352208777296, 0.411112290507187, 0.606530659712633, 0.800737402916808, 0.945959468906766, 
              1, 0.945959468906766, 0.800737402916808, 0.606530659712633, 0, 
              NA, NA, NA, NA,NA,NA,
              NA, NA, NA, NA, NA, NA)

FigureC <- tibble(temperature = temperature, 
                  dr_tropical = FigureA$tropical,
                  dr_temperate = dr_temperateC,
                  su_tropical = FigureB$tropical,
                  su_temperate = FigureB$temperate, 
                  P_tropical = dr_tropical * su_tropical,
                  P_temperate = dr_temperate * su_temperate)

ggplot(FigureC, aes(x = temperature, y = dr_temperate))+
  geom_line(aes(x = temperature, y = P_temperate), col = "grey", size = 2)+
  geom_line(aes(x = temperature, y = su_temperate), col = "black", size = 2, linetype = 2)+
  geom_line(linetype = 3, size = 2)+
  scale_color_viridis_d()+
  theme_cowplot()


```

```{r}
FigureD <- tibble(temperature = temperature, 
                  dr = FigureB$dr, 
                  survival = FigureA$survival, 
                  P = dr * survival)
ggplot(FigureD, aes(x = temperature, y = dr))+
  geom_line(aes(x = temperature, y = P), col = "grey", size = 2)+
  geom_line(aes(x = temperature, y = survival), col = "black", size = 2, linetype = 2)+
  geom_line(linetype = 3, size = 2)+
  scale_color_viridis_d()+
  theme_cowplot()
```


```{r}

ggplot(FigureC, aes(x = temperature, y = dr_tropical))+
  geom_line(aes(x = temperature, y = P_tropical), col = "grey", size = 2)+
  geom_line(aes(x = temperature, y = su_tropical), col = "black", size = 2, linetype = 2)+
  geom_line(linetype = 3, size = 2)+
  scale_color_viridis_d()+
  theme_cowplot()

```


Is there a trade-off between cold and heat tolerance?

```{r}
Tolerance <- tibble(cold= rnorm(50, mean = 12, sd = 2), 
                    hot1 = rnorm(50, mean = 2, sd = 2), 
                    hot2 = (cold/2)-3,
                    hot3 = 10 -(cold/2))

ggplot(Tolerance, aes(x = hot1, y = cold)) +
  geom_point(size = 2)+
  theme_cowplot()+
  geom_hline(yintercept = 12)+
  ylim(0,20)+
  xlim(0,20)+
  ylab("cold tolerance")+
  xlab("heat tolerance")
  
  
  
```

```{r}
ggplot(Tolerance, aes(x = hot2, y = cold)) +
  geom_point(size = 2)+
  theme_cowplot()+
  geom_hline(yintercept = 12)+
  ylim(0,20)+
  xlim(0,20)+
  ylab("cold tolerance")+
  xlab("heat tolerance")
```



```{r}
ggplot(Tolerance, aes(x = hot3, y = cold)) +
  geom_point(size = 2)+
  theme_cowplot()+
  geom_hline(yintercept = 12)+
  ylim(0,20)+
  xlim(0,20)+
  ylab("cold tolerance")+
  xlab("heat tolerance")
```

