---
title: "Methods figure"
output: html_notebook
---

```{r}
# Load packages
library(tidyverse)
library(readxl)
library(cowplot)
```
 
Import data
```{r}
#Anna's Laptop
Data <- read_xlsx("~/Desktop/PhysiologyDatabaseVersion5.xlsx", sheet = "WorkingTable", na = c("NA", ""))

SetQuality <- read_csv("~/Desktop/GitHub/ThermalStressCompilation/SetCharacterization.csv")
```


```{r}
# Mariana's desktop
Data <- read_xlsx("/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/PhysiologyDatabaseVersion5.xlsx", 
                  sheet = "WorkingTable", na = c("NA", "")) 
SetQuality <- read_csv("SetCharacterization.csv")
# cols 1 to 7 correspond to dt, 8 to 13 to survival, 14 to development time

```

```{r}
# Mariana's laptop
Data <- read_xlsx("/Users/mar/Desktop/Projects/ThermalPerformance/PhysiologyDatabaseVersion5.xlsx", 
                  sheet = "WorkingTable", na = c("NA", "")) 
SetQuality <- read_csv("SetCharacterization.csv")
# cols 1 to 7 correspond to dt, 8 to 13 to survival, 14 to development time
```

```{r}
# get lists of complete sets for development and survival
d_complete_sets <- SetQuality %>% # to get a list of complete sets for dt
  filter(just.rise == FALSE) %>% 
  select(set) %>% 
  distinct()

s_complete_sets <- SetQuality %>% # to get a list of complete sets for survival
  filter(just.rise1 == FALSE) %>% 
  select(set) %>% 
  distinct()
```

Remove parasitoids and convert character to factor
```{r}
unique(Data$status)
Ana <- Data[Data$status != "parasitoid",]
rm(Data)

# convert character to factor
Ana <- Ana %>%
  mutate_if(is.character, factor)
```

# Methods
We compiled a database including records of development time and/or survival  of Lepidoptera growing at constant temperatures. The full table includes `r length(unique(Ana$set))` sets of `r length(unique(Ana$sp))` species from `r length(unique(Ana$family))` families from `r length(unique(Ana$locality))` localities. See supplementary information for a full list of species.
Using linear interppolation we calculated the following variables in all complete sets (curves including both a rise and a fall) for development rate (n = `r length(d_complete_sets$set)`) and for survival (n = `r length(s_complete_sets$set)`)

dTopt: temperature that maximizes development rate

max_sTopt: maximum temperature that maximizes survival, 

min_sTopt: minimum temperature maximizing survival

dBreadth: interval of temperatures permiting >= 50% development efficiency

sBreadth: interval of temperatures permiting >= 50% survival

P: survival*development (fitness)


```{r}
# Woring set and functions

set_w <- filter(Ana, set == 175) # working set

# Functions to make a smooth curve 
# 1 Development rate
fit_dr <- function(df){
  glm(dr ~ temp + I(temp^2), data = df, family = quasibinomial)
}
# 2. Model: survival ~ temperature 
fit_survival <- function(df) {
  glm(survival/100 ~ temp + I(temp^2), data = df, family = quasibinomial)
}

# 3. Calculate predicted values over the full temperature range
get_predicted <- function(model){
  pred <- tibble(temp = c(1:40), 
                 predicted = predict.glm(model, newdata = tibble(temp = c(1:40)), type = "response"))
}

#  T0 from GDD model
Data <- set_w
get_t0 <- function(Data){
  d_opt <- max(Data$dr) # maximum development rate value
  dTopt <- filter(Data, dr == d_opt)[["temp"]] 
  Data <- filter(Data, temp <= dTopt, dr > 0) # to include only linear section
	  fit <- lm(dr ~ temp, data = Data) # linear regression
    T0 <- as.numeric(- coef(fit)[1]/coef(fit)[2]) # T0 is the x intercept
    slope_f <- as.numeric(coef(fit)[2])
    R2 <- as.numeric(summary(fit)$r.squared) #extract r2
    P <- as.numeric(summary(fit)$coefficients[2, 4]) #extract P
    shap <- as.numeric(shapiro.test(fit$residuals)[2])
    Data$G <- Data$dt * (Data$temp - as.numeric(as.character(T0)))
    mean_G <- mean(Data$G, na.rm = T)
    mean_temp <- mean(Data$temp, na.rm = T)
    Output <- tibble(T0 = T0, 
                     slope = slope_f, 
                     N_temp = nrow(Data),
                     spread = max(Data$temp) - min(Data$temp),
                     r2= R2, P = P, shapiro_testP = shap, G = mean_G, 
                  meantemp = mean_temp) 
                  #plot = plot(dr~temp, data = Data, main = paste(Data$sp[1], Data$set[1]))
    
}


# Apply functions to set 636
hypofit <- fit_dr(set_w)

hyposu <- fit_survival(set_w)

dtpredicted <- get_predicted(hypofit)

supredicted <- get_predicted(hyposu)

predicted <- left_join(dtpredicted, supredicted, by = "temp")

names(predicted) <- c("temp", "dr", "survival")

rm(hypofit, hyposu, dtpredicted, supredicted)

```

```{r}
# Figure 1A: development rate, linear regression and T0
d_opt <- max(set_w$dr) # optimum development rate
dTopt <- filter(set_w, dr == d_opt)[["temp"]] # temperature that maximizes development rate
gs <- "grey66"
t0table <- get_t0(set_w)
text <- tibble(temp = 7.5,
               dr = 0.02)
linear_3 <- filter(set_w, temp <= dTopt, dr > 0)
model <- lm(dr~temp, data = linear_3)
linear_predicted <-  tibble(temp = c(0:40),
  predicted_dr = predict.lm(model, newdata = tibble(temp = c(0:40)), type = "response"))
t0 <- t0table[[1]]
Figure1A <- ggplot(data = linear_3, aes(x = temp, y = dr)) + 
  geom_rect(xmin = t0, xmax = max(linear_3$temp), ymin = 0, ymax = Inf, linetype = 3,alpha = 0.03, fill = gs)+
  geom_line(data = linear_predicted, aes(x = temp, y = predicted_dr))+
  geom_point(data = linear_3, mapping = aes(x = temp, y = dr))+
  geom_point(data = set_w, mapping = aes(x = temp, y = dr), shape = 1)+
   geom_text(data = text, aes(x = temp+15, y = dr, label = paste("T0 =",round(t0, digits = 2), sep = " ")))+
  #geom_hline(yintercept = 0, col = gs)+
  
  #geom_vline(xintercept  = 3.761635, linetype = 4)+
  theme_cowplot()+
   ylim(0, 0.35)+
   xlim(5, 43)+
  labs(x = "Temperature (°C)", y = "Development rate")
Figure1A
```
Figure 1A. Development rate (1/development time) of  `r set_w$sp [[1]]` `r set_w$lifestage [[1]]`s kept at constant temperatures. Line indicates a linear regression including only the treatments in which development rate has a linear relationship with temperature (solid points). T0 is estimated as the x-intercept of this relationship (9.7 C). The shaded region indicates the interval of temperatures at which development is assumed to occur.

```{r}
# Figure 1B: Development rate, sub and supra optimum breadth. 
 
d_50 <- d_opt/2 # value of development rate at 50% efficiency
out <- seq(from = 0, to = 40, by = 0.1)
interpolation <- as.data.frame(approx(x = set_w$temp, y = set_w$dr, xout = out, method = "linear", rule = 1))
names(interpolation) <- c("temp", "dr")
d_sub_min <- head(filter(interpolation, dr >= d_50), 1)[[1]]
d_sub_max <- tail(filter(interpolation, dr >= d_50), 1)[[1]]
cd <- "#E69F00"
gd <- "grey20"
cs <- "#56B4E9"
label_text1 <- tibble(x = 25,
                    y = 0.12,
                label = "Development interval")

Figure1B <- ggplot(data = predicted, aes(x = temp, y = dr)) + 
  geom_rect(ymin = -Inf, ymax = +Inf, xmin = d_sub_min, xmax = d_sub_max, alpha = 0.03, fill = gs)+
  geom_point(data = set_w, mapping = aes(x = temp, y = dr))+
  geom_line(data = set_w, mapping = aes(x = temp, y = dr))+
  geom_text(data = label_text1, aes(x = x + 7.5, y = y ,label ="D-interval"))+
  geom_segment(x = d_sub_min, xend = d_sub_max, y = d_50, yend = d_50, linetype = 3)+
  #geom_segment(x = dTopt, xend = sub_max, y = d_50, yend = d_50, linetype = 2)+
   ylim(0, 0.35)+
  geom_text(data = label_text1, mapping = (aes(x = x+10, y = y + 0.29,label ="Dopt")))+
   xlim(5, 43)+
  theme_cowplot()+
  
  labs(x = "Temperature", y = "Development rate")
Figure1B
```
Figure 1B. By linear interpolation we estimated the interval of temperatures that allow for development rate at 50% of the maximum efficiency `r d_opt` at `r dTopt` (0.22 at 32C)
```{r}
# Figure 1C: Survival. Breadth and cold and hot optima

out <- seq(from = 0, to = 40, by = 0.1)
s_interpolation <- as.data.frame(approx(x = set_w$temp, y = set_w$survival, xout = out, method = "linear", rule = 1))
names(s_interpolation) <- c("temp", "survival")
su_min <- head(filter(s_interpolation, survival >= 50), 1)[[1]]
su_max <- tail(filter(s_interpolation, survival >= 50), 1)[[1]]
label_text <- tibble(x = 25,
                    y = 0.6,
                label = "Survival interval")

Figure1C <- ggplot(data = set_w,aes(x = temp, y = survival/100)) + 
  geom_rect(ymin = -Inf, ymax = +Inf, xmin = su_min, xmax = su_max, alpha = 0.03, fill = gs)+
  geom_point(shape = 2)+
  geom_line(data = set_w, mapping = aes(x = temp, y = survival/100))+
  geom_segment(x = su_min, xend = su_max, y = 0.5, yend = 0.5, linetype = 3)+
  geom_segment(x  = 32.5, xend  = 32.5, y = 0.5, yend = 0.5, linetype = 2)+
  theme_cowplot()+
  geom_text(data = label_text, aes(x = x, y = y ,label ="S-interval"))+
  geom_text(data = label_text, aes(x = x - 16 , y = y + 0.4 ,label ="minSopt"))+
  geom_text(data = label_text, aes(x = x + 16, y = y + 0.4 ,label ="maxSopt"))+
  #ylim(0, 0.35)+
   xlim(5, 43)+
  labs(x = "Temperature", y = "Survival")
Figure1C
```
Figure 1C. Survival of eggs at `r set_w$sp [[1]]` `r set_w$lifestage [[1]]`s kept at constant temperatures. Blue interval corresponds to treatments that allowed for at least 50% survival
```{r}
# Figure 1D: Performance: in gray survival and development rate curves, in bold the resulting performance curve (d*p). 

set_w$P <- set_w$dr * (set_w$survival/100)
P_opt <- max(set_w$P)

P50 <- P_opt/2
PTopt <- filter(set_w, P ==P_opt)[["temp"]]
out <- seq(from = 0, to = 40, by = 0.1)
Pinterpolation <- as.data.frame(approx(x = set_w$temp, y = set_w$P, xout = out, method = "linear", rule = 1))
names(Pinterpolation) <- c("temp", "P")
P_min <- head(filter(Pinterpolation, P >= P50), 1)[["temp"]]
P_max <- tail(filter(Pinterpolation, P >= P50), 1)[["temp"]]
label_sub <- tibble(temp = 28,
                    P = 0.25,
                label = "suboptimum")
label_supra <- tibble(temp = 37,
                    P = 0.25,
                label = "supraoptimum")
PoptPoint <- filter(Pinterpolation, P == P_opt)
PlastPoint <- tail(filter(Pinterpolation, P > 0),1)[["P"]]
Figure1D <- ggplot(data = set_w,aes(x = temp, y = P/0.4)) + 
  geom_line()+
  geom_point(data = set_w, mapping = aes(x = temp, y = survival/100), shape = 2, col= "darkgrey")+
  geom_point(data = set_w, mapping = aes(x = temp, y = dr/0.4), col = "darkgrey")+
  geom_line(data = set_w, mapping = aes(x = temp, y = survival/100), linetype = 2, col = "darkgrey")+
  geom_line(data = set_w, mapping = aes(x = temp, y = dr/0.4),linetype = 2, col = "darkgrey")+
  
  geom_segment(x = P_min, xend = PTopt, y = P50/0.4, yend = P50/0.4,linetype = 3)+
  #geom_text(data = label_sub, aes(x = temp, y = P + 0.01,label ="sub-Popt"))+
  
  geom_segment(x = PTopt, xend = P_max, y = P50/0.4, yend = P50/0.4, linetype = 3)+
  #geom_text(data = label_supra, aes(x = temp, y = P +.01,label ="supra\nPopt"))+
  geom_text(data = label_sub, aes(x = temp + 4, y = P - .15,label ="P-interval"))+
  
  theme_cowplot()+
  geom_rect(ymin = -Inf, ymax = +Inf, xmin = P_min, xmax = PTopt, alpha = 0.03, fill = gd)+
  geom_rect(ymin = -Inf, ymax = +Inf, xmin = PTopt, xmax = P_max, alpha = 0.06, fill = gd)+
  geom_segment(x  = PTopt, xend  = PTopt, y = 0, yend = +Inf, linetype = 1, col = "darkgrey")+
  
  geom_point(x  = PTopt, y = P_opt/0.4)+
  geom_text(data = PoptPoint-2, aes(x = temp-3, y = 0.75,label ="Popt"))+
  
  scale_y_continuous(sec.axis = sec_axis(~.* 0.4, name = "P, Development rate"))+
   #ylim(0, 0.35)+
   xlim(5, 43)+
  labs(x = "Temperature", y = "Survival")
Figure1D
```
Figure 1D. Performance (P, solid line) is the result of multipliying survival (triangles, dashed line) by development rate (points, dashed line). Shaded region denotes the interval of temperatures that allows for at least 50% survival and 50% development efficiency. This interval can be split into sub-optimal temperatures (below Popt, ligth grey) and supra-optimal temperatures, (above Popt, dark grey) 
Performance of `r unique(set_w$sp)` (`r unique(set_w$family)`), from `r unique(set_w$locality)`

```{r}
#Multiplot
plot_grid(Figure1B, Figure1C, Figure1A,Figure1D,labels = c('A', 'B', 'C', 'D'))
```


Other descriptive stats of the data set

```{r}
Temps_by_set <- Ana %>% 
  select(set, temp) %>% 
  group_by(set) %>% 
  tally(sort = T)
```

Each set is a unit including `r min(Temps_by_set$n)` to `r max (Temps_by_set$n)`  constant temperature treatments (mode = `r which.max(tabulate(Temps_by_set$n))`) and the corresponding values of development time and/or survival.


```{r}
non_redundant_sets <- Ana %>% 
  select(id, sp, lifestage,locality) %>%
  distinct()  
  
  
  
```

There are multiple sets per population in this database because performance was often evaluated under a variety of experimental conditions including different host plants or daylengths, in addition, data was sometimes reported by sex. We considered sets from the same locality and ontogenetic stage that were reported in the same study (publication) to be redundant and thus combined them by taking the geometric mean at each temperature treatment. After this correction we had `r nrow(non_redundant_sets)` non-redundant sets. 
