---
title: "Analyze performance variables"
output: html_notebook
---
## Contents
1. Import data and housekeeping
2. Popt 
3. P-width

```{r}
# load packages
library(tidyverse)
library(readxl)
library(cowplot)
library(psych) # to get geometric mean
library(lme4)
library(car)
library("dotwhisker")  # to make coefficient plots
library(sjPlot) #prediction plots
library(sjmisc)
library(broom)
```

1. Import file
```{r}
# Mariana's desktop

responses <- read_csv("/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/ThermalResponsesII.csv")

```


```{r}
# Mariana's laptop
responses <- read_csv("/Users/mar/Desktop/Projects/ThermalPerformance/ThermalResponsesII.csv")
```

```{r}
# Anna's Laptop

responses <- read_csv("~/Desktop/ThermalResponsesII.csv")
```

```{r}
# convert character to factor
responses <- responses %>%
  mutate_if(is.character, factor)
# discard non-pertinent data

responses_table <- responses %>% 
  filter(lifestage == "egg"| lifestage == "larva"| lifestage == "pupa", quality!= "inferred", quality != "combination", !is.na(Popt)) %>% 
  select(Popt, Plower, Pupper, Pwidth, lifestage, sp, lat) %>% 
  mutate(absLat = abs(lat))
responses_table$lifestage <- factor(responses_table$lifestage)
responses_table$sp <- factor(responses_table$sp)

```




```{r}
# 1. Subset to P metrics and predictors only
P.set <- with(responses_table, cbind(Popt, Pupper, Pwidth, Plower, lifestage, absLat, sp))

# Full models, with random intercept, and then random slopes

## random slopes: need to input dataset as dataframe

fit1_fullrandom <- lmer(Popt ~ lifestage + absLat + lifestage:absLat + (lifestage|sp), data = as.data.frame(P.set)) #

summary(fit1_fullrandom)

fit_red1random <- lmer(Popt ~ lifestage + absLat + (lifestage|sp), REML = F, data = as.data.frame(P.set))

anova(fit1_fullrandom, fit_red1random, test.statistic = "Chisq")
## use full model




summary(fit1_fullrandom)
Anova(fit1_fullrandom)
P.set <- as.data.frame(P.set)
meanP <- P.set %>% 
  select(Popt, absLat) %>% 
  ungroup() %>% 
  filter(absLat > 30 & absLat < 50) %>% 
  summarise(mean = mean(Popt, na.rm = T), 
            sd = sd(Popt, na.rm = T))
#plot.design(Dopt ~ lifestage *  absLat + (lifestage|sp) , data = D.set)
dw_plot(fit1_fullrandom) +
  #geom_vline(xintercept= mean(D.set$Dopt),lty=2)+
  geom_vline(xintercept= 0,lty=3)+
  theme_cowplot()+
  scale_color_viridis_d()+
  theme(legend.position = "none")
```
 
```{r}

#fit model without lifestage in the random effect
fit1_fullrandomsp <- lmer(Popt ~ lifestage + absLat + lifestage:absLat + (1|sp), data = as.data.frame(P.set)) #
dw_plot(fit1_fullrandomsp) +
  #geom_vline(xintercept= mean(D.set$Dopt),lty=2)+
  geom_vline(xintercept= 0,lty=3)+
  theme_cowplot()+
  scale_color_viridis_d()+
  theme(legend.position = "none")
```
 
 
 P-width
```{r}
# get mean at temperate latitudes
meanPwidth <- P.set %>% 
  select(Pwidth, absLat) %>% 
  ungroup() %>% 
  filter(absLat > 30 & absLat < 50) %>% 
  summarise(mean = mean(Pwidth, na.rm = T), 
            sd = sd(Pwidth, na.rm = T))

# Full models, with random intercept, and then random slopes
fit2_full <- lmer(Pwidth ~ lifestage + absLat + lifestage:absLat + (1|sp), data = as.data.frame(P.set)) #

fit2_reduced <- lmer(Pwidth ~ lifestage + absLat  + (1|sp), data = as.data.frame(P.set)) #

anova(fit2_full, fit2_reduced, test.statistic = "chisq")
fit2_reducedML <- lmer(Pwidth ~ lifestage + absLat  + (1|sp),  REML = F, data = as.data.frame(P.set)) #
summary(fit2_reducedML)

dw_plot(fit2_reducedML) +
  #geom_vline(xintercept= mean(D.set$Dopt),lty=2)+
  geom_vline(xintercept= 0,lty=3)+
  theme_cowplot()+
  scale_color_viridis_d()+
  theme(legend.position = "none")
```
 
Upper limit of D-width
 
```{r}
# Full models, with random intercept, and then random slopes
fit3_fullrandom <- lmer(Pupper ~ lifestage + absLat + lifestage:absLat + (lifestage|sp), data = as.data.frame(P.set)) #
# this model has too many random effects
fit3_full <- lmer(Pupper ~ lifestage + absLat + lifestage:absLat + (1|sp), data = as.data.frame(P.set)) #


fit3_reduced <- lmer(Pupper ~ lifestage + absLat  + (1|sp), data = as.data.frame(P.set)) #

anova(fit3_full, fit3_reduced, test.statistic = "chisq")
 # use full model
fit3_fullML <- lmer(Pupper ~ lifestage + absLat + lifestage:absLat + (1|sp),  REML = F, data = as.data.frame(P.set))

summary(fit3_fullML)

dw_plot(fit3_fullML) +
  #geom_vline(xintercept= mean(D.set$Dopt),lty=2)+
  geom_vline(xintercept= 0,lty=3)+
  theme_cowplot()+
  scale_color_viridis_d()+
  theme(legend.position = "none")
```

 
 
Lower limit of D-width
 
```{r}
# Full models, with random intercept, and then random slopes
fit4_fullrandom <- lmer(Plower ~ lifestage + absLat + lifestage:absLat + (lifestage|sp), data = as.data.frame(P.set)) #
# this model failed to converge
fit4_full <- lmer(Plower ~ lifestage + absLat + lifestage:absLat + (1|sp), data = as.data.frame(P.set)) #


fit4_reduced <- lmer(Plower ~ lifestage + absLat  + (1|sp), data = as.data.frame(P.set)) #

anova(fit4_full, fit4_reduced, test.statistic = "chisq")

# use reduced model
fit4_reducedML <- lmer(Plower ~ lifestage + absLat + lifestage:absLat + (1|sp),  REML = F, data = as.data.frame(P.set))

summary(fit4_reducedML)

dw_plot(fit4_reducedML) +
  #geom_vline(xintercept= mean(D.set$Dopt),lty=2)+
  geom_vline(xintercept= 0,lty=3)+
  theme_cowplot()+
  scale_color_viridis_d()+
  theme(legend.position = "none")
```

 
```{r}
## Omit NA rows for PCA. need dataframe for input
P.set.nona <- as.data.frame(na.omit(P.set))
Ppc <- princomp(~Popt + Pupper + Pwidth, data = P.set.nona)
summary(Ppc)
```

```{r}
loadings(Ppc)
```


```{r}
names(Ppc)
## Bind PC scores to dataset

P.set.pcs <- cbind(P.set.nona, Ppc$scores)
head(P.set.pcs)

## Now model Comp.1 (PC scores on 1st PC)
fit1_red1 <- lmer(Comp.1 ~ lifestage + absLat + (1|sp), data = P.set.pcs)
summary(fit1_red1)

## test simpler models
fit1_red2 <- lmer(Comp.1 ~ lifestage  + (1|sp), data = P.set.pcs)
fit1_red3 <- lmer(Comp.1 ~ absLat + (1|sp), data = P.set.pcs)
fit1_red4 <- lmer(Comp.1 ~ (1|sp), data = P.set.pcs)
anova(fit1_red1,fit1_red2, fit1_red3,fit1_red4, test.statistic = "Chisq")

## estimate parameters using ML rather than REML (default)
fit1_red3.ML <- lmer(Comp.1 ~  absLat  + (1|sp), data = P.set.pcs, REML = F)
summary(fit1_red3.ML)

dw_plot(fit1_red3.ML) +
  #geom_vline(xintercept= mean(D.set$Dopt),lty=2)+
  geom_vline(xintercept= 0,lty=3)+
  theme_cowplot()+
  scale_color_viridis_d()+
  theme(legend.position = "none")


```
 The first component explained 68% of variation, the second component explained 26%
 Popt and Pupper were positively correlated. 
 Popt and Pwidth were negatively correlated?
 
 
 
```{r}
# PCA analysis for D and S metrics
PCA_responses_table <- responses %>% 
  filter(lifestage == "egg"| lifestage == "larva"| lifestage == "pupa", quality!= "inferred", quality != "combination", !is.na(Popt)) %>% 
  mutate(absLat = abs(lat))
responses_table$lifestage <- factor(responses_table$lifestage)
responses_table$sp <- factor(responses_table$sp)
DS.set <- with(PCA_responses_table, cbind(Dopt, Dupper, Dwidth, maxSopt, Supper, Swidth, lifestage, absLat, sp))

DS.set.nona <- as.data.frame(na.omit(DS.set))
DSpc <- princomp(~Dopt + Dupper + Dwidth + maxSopt + Supper + Swidth, data = DS.set.nona)
```

```{r}
# PCA analysis for D and S metrics
summary(DSpc)
```

```{r}
# PCA analysis for D and S metrics
loadings(DSpc)
```

```{r}
# PCA analysis for D and S metrics
names(DSpc)

## Bind PC scores to dataset

DS.set.pcs <- cbind(DS.set.nona, DSpc$scores)
head(P.set.pcs)

## Now model Comp.1 (PC scores on 1st PC)
fit5_red1 <- lmer(Comp.1 ~ lifestage + absLat  + (1|sp), data = DS.set.pcs)
summary(fit5_red1)

## test simpler models
fit5_red2 <- lmer(Comp.1 ~ lifestage  + (1|sp), data = DS.set.pcs)
fit5_red3 <- lmer(Comp.1 ~ absLat + (1|sp), data = DS.set.pcs)
fit5_red4 <- lmer(Comp.1 ~ (1|sp), data = DS.set.pcs)
anova(fit5_red1,fit5_red2, fit5_red3,fit5_red4, test.statistic = "Chisq")

## estimate parameters using ML rather than REML (default)
fit5_red3.ML <- lmer(Comp.1 ~  absLat  + (1|sp), data = DS.set.pcs, REML = F)
summary(fit5_red3.ML)

dw_plot(fit5_red3.ML) +
  #geom_vline(xintercept= mean(D.set$Dopt),lty=2)+
  geom_vline(xintercept= 0,lty=3)+
  theme_cowplot()+
  scale_color_viridis_d()+
  theme(legend.position = "none")


```
 
 
 
 
 
```{r}
names(responses_table)

plot(T0 ~ Popt, data = responses_table)

```
 
 

```{r}
meanPupper <- P.set %>% 
  select(Pupper, absLat) %>% 
  ungroup() %>% 
  filter(absLat > 30 & absLat < 50) %>% 
  summarise(mean = mean(Pupper, na.rm = T), 
            sd = sd(Pupper, na.rm = T))
```

```{r}
meanPlower <- P.set %>% 
  select(Plower, absLat) %>% 
  ungroup() %>% 
  filter(absLat > 30 & absLat < 50) %>% 
  summarise(mean = mean(Plower, na.rm = T), 
            sd = sd(Plower, na.rm = T))
```





# Old code below____________________________________________________________


To assess variation in thermal responses across latitude and life stages, we fitted linear mixed models, including each thermal response scaled and life and laitude as fixed factors and species as a random factor. We discarded "eggtoemergence" category in the subsequent analyses because it is a combination of the other three.

## Popt: temperature that maximizes development rate
Popt ranged from `r min(responses_table$Popt, na.rm = T)` to `r max(responses_table$Popt, na.rm = T)`, mean =  `r mean(responses_table$Popt, na.rm = T)`, SD = `r sd(responses_table$Popt, na.rm = T)`
Latitude ranged from `r min(responses_table$absLat, na.rm = T)` to `r max(responses_table$absLat, na.rm = T)`


```{r}
# 1. Full model
# Scale response variable
#responses_table$Popt_s <- scale(responses_table$PTopt)[,1]

fit1_full <- lmer(Popt ~ -1 + lifestage * absLat + (1|sp), data = responses_table) #

# this model cannot be done because we have 134 observations and the number of random effects would be 171
#fit1_full_random <- lmer(Popt ~ -1 + lifestage * absLat + (lifestage|sp), data = responses_table) #
# this model is singular
#fit1_full_family <- lmer(Popt ~ -1 + lifestage * absLat + (1|sp) + (1|family), data = responses_table) #

summary(fit1_full)

fit1_full_aic <- AIC(fit1_full)
fit1_fullpvals <- Anova(fit1_full)
dw_plot(fit1_full) +
  geom_vline(xintercept = mean(responses_table$Popt, na.rm = T),lty = 2)+
  geom_vline(xintercept = 0,lty = 3)+
  theme_cowplot()+
  scale_color_viridis_d()+
  theme(legend.position = "none")
# Figure 1. effects and 95% CI of the full model
```
 Full model: lmer(Popt ~ -1 + lifestage * absLat + (1|sp), data = responses_table)
 Full model's AIC: `r fit1_full_aic`


```{r}
plot_model(fit1_full, type = "int")+
  theme_cowplot()+
  scale_color_viridis_d()
```

```{r}
# 1. Full model
# Scale response variable
#responses_table$Pwidth_s<- scale(responses_table$Pwidth)[,1]

fit2_full <- lmer(Pwidth ~ -1 + lifestage * absLat + (1|sp), data = responses_table) #

summary(fit2_full)

fit2_full_aic <- AIC(fit2_full)
fit2_fullpvals <- Anova(fit2_full)
dw_plot(fit2_full) +
  geom_vline(xintercept = mean(responses_table$Pwidth, na.rm = T),lty=2)+
  geom_vline(xintercept = 0,lty = 3)+
  theme_cowplot()+
  scale_color_viridis_d()+
  theme(legend.position = "none")
# Figure 1. effects and 95% CI of the full model
```
 Full model: lmer(P-width_s ~ -1 + lifestage * absLat + (1|sp), data = responses_table)
 Full model's AIC: `r fit2_full_aic`

```{r}
# Plot model predictions
# type = c("est", "re", "eff", "pred", "int", "std", "std2", "slope", "resid", "diag")
plot_model(fit2_full, type = "int")+
  theme_cowplot()+
  scale_color_viridis_d()

```



```{r}
# 1. Full model
# Scale response variable
#responses_table$Plower_s <- scale(responses_table$Plower)[,1]

fit3_full <- lmer(Plower ~ -1 + lifestage * absLat + (1|sp), data = responses_table) #

summary(fit3_full)

fit3_full_aic <- AIC(fit3_full)
fit3_fullpvals <- Anova(fit3_full)
dw_plot(fit3_full) +
  geom_vline(xintercept = mean(responses_table$Plower, na.rm = T),lty = 2)+
  geom_vline(xintercept = 0,lty = 3)+
  theme_cowplot()+
  scale_color_viridis_d()+
  theme(legend.position = "none")
#  effects and 95% CI of the full model

```


 Full model: lmer(lower ~ -1 + lifestage * absLat + (1|sp), data = responses_table)
 Full model's AIC: `r fit3_full_aic`

```{r}
# Plot model predictions
# type = c("est", "re", "eff", "pred", "int", "std", "std2", "slope", "resid", "diag")
plot_model(fit3_full, type = "int")+
  theme_cowplot()+
  scale_color_viridis_d()
```


```{r}
# 1. Full model
# Scale response variable
#responses_table$upper_s <- scale(responses_table$Imax)[,1]

fit4_full <- lmer(Pupper ~ -1 + lifestage * absLat + (1|sp), data = responses_table) #

summary(fit4_full)

fit4_full_aic <- AIC(fit4_full)
fit4_fullpvals <- Anova(fit4_full)
dw_plot(fit4_full) +
  geom_vline(xintercept = mean(responses_table$Pupper, na.rm = T), lty=2)+
  geom_vline(xintercept = 0 ,lty = 3)+
  theme_cowplot()+
  scale_color_viridis_d()+
  theme(legend.position = "none")
#  effects and 95% CI of the full model

```


 Full model: lmer(upper ~ -1 + lifestage * absLat + (1|sp), data = responses_table)
 Full model's AIC: `r fit4_full_aic`

```{r}
# Plot model predictions
# type = c("est", "re", "eff", "pred", "int", "std", "std2", "slope", "resid", "diag")
plot_model(fit4_full, type = "int")+
  theme_cowplot()+
  scale_color_viridis_d()
```

ll model: lmer(Popt_s ~ -1 + lifestage * absLat + (1|sp), data = responses_table)
 Full model's AIC: `r fit1_full_aic`

```{r}
# #2. Reduced model
# fit1_reduced <- lmer(Popt_s ~  -1 + lifestage + absLat + (1|sp), data = responses_table) #
# 
# # fit1Noctuids <- lmer(Popt ~ -1 + lifestage + absLat + (1|sp), data = filter(ToptAnalysis, family == "Noctuidae"))
# # fit1Not_noctuids <- lmer(Popt ~ -1 + lifestage + absLat + (1|sp), data = filter(ToptAnalysis, family != "Noctuidae"))
# tidy(fit1_reduced)
# fit1_summary <- summary(fit1_reduced)
# fit1_reduced_aic <- AIC(fit1_reduced)
# #Anova(fit1_reduced)
# dw_plot(fit1_reduced, conf.level = 0.95) + 
#   geom_vline(xintercept=0,lty=2)+
#   theme_cowplot()+
#   scale_color_viridis_d()+
#   theme(legend.position = "none")#+
#   #scale_y_discrete(labels = c('sd Observation.Residual','sd species', 
#                              # 'Latitude', 'Life stage (pupa)','Life stage (larva)'))
# 

```
Figure 2. Popt was higher at lower latitudes
Lots of variation among species!
