---
title: "Performance"
output: html_notebook
---
## Sections
1. Import data
2. Calculate P (survival * development rate)
3. Select sets with at least 4 data points and a complete curve (rise, peak, fall) for egg, larva, pupa, and egg to adult.
4. Consolidate redundant sets.
5. Obtain predicted performance (P) in the range from 0 to 40 C using linear interpolation
6. Calculate Popt (temperatures that maximizes performance)
7. Calculate P-interval (min, max, length)
8. Visualization
9. Analyses

We used r (tydiverse) for data wrangling, (ggplot2) for data visualization, (psych) for obtaining geometric mean, (readxl) to import data files and (cowplot) to format ggplots.

```{r}
# load packages
library(tidyverse)
library(readxl)
library(cowplot)
library(psych)
```

Import data 
```{r}
# Mariana's desktop
Data <- read_xlsx("/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/PhysiologyDatabaseVersion5.xlsx", 
                  sheet = "WorkingTable", na = c("NA", ""))
```

```{r}
# Mariana's laptop
Data <- read_xlsx("/Users/mar/Desktop/Projects/ThermalPerformance/PhysiologyDatabaseVersion5.xlsx", 
                  sheet = "WorkingTable", na = c("NA", "")) 

```


```{r}
# Anna's Laptop
Data <- read_xlsx("~/Desktop/PhysiologyDatabaseVersion5.xlsx", 
                  sheet = "WorkingTable", na = c("NA", "")) 

```


```{r}
# Discard parasitoids and change character to factor 
unique(Data$status)
Ana <- Data[Data$status != "parasitoid",]
rm(Data)

# convert character to factor
Ana <- Ana %>%
  mutate_if(is.character, factor)
```

# 2. Calculate P (only sets with survival and development time)

```{r}
Ana$P <- Ana$dr * (Ana$survival/100)
```


# 3. Select sets with at least 4 data points and a complete curve (rise, peak, fall) for egg, larva, pupa, and egg to adult.

To be included in performance analyses, sets should have at least 4 temperature treatments reporting survival and form a complete curve (rise, peak, fall). We restricted analyses to four life stages: egg, larva, pupa and egg to adult.

```{r}
# Function to determine whether a set is:
# "Complete" (has low, peak and high values)
# "Incomplete" (only low and peak)
# "No data" (no P data available)
Ana$set2 <- Ana$set
is_complete <- function(dat){
  dat <- dat[!is.na(dat$P) & !is.na(dat$temp),] # Remove treatments with no performance data
 if(nrow(dat) > 3 ){
  
   y_max <- max(dat$P, na.rm = T) # get maximum value for performance
  x_max <- dat[dat$P == y_max, "temp"][[1]][1]# get temperature that maximizes performance 
  
  species <- unique(dat$sp)
  lifestage <- unique(dat$lifestage)
  seti <- unique(dat$set2)
  print(paste("set:",seti))
  #print(species)
  #print(lifestage)
  #print(x_max)
  #print(y_max)
  cold <- filter(dat, temp < x_max)  
  hot <- filter(dat, temp > x_max)  
  result = nrow(cold[!is.na(cold$P),]) && nrow(hot[!is.na(hot$P),])
  if(result){
    print("Complete")
  } else {
    print("Incomplete")
  }
 } else{
   print("Less than 4 data points")
 }
  
  
}

```


```{r}
# Apply is_complete() to all sets
SetAssessment <- Ana %>% 
  group_by(set) %>% 
  nest() %>% 
  mutate(curve = map(data,is_complete)) %>% 
  select(set, curve) %>% 
  unnest() 

```


From a total of `r length(unique(Ana$set))`, only  `r nrow(filter(SetAssessment, curve == "Complete"))` had a complete curve for performance.
```{r}
# Select sets of life stages of interest: egg, larva, pupa, egg to emergence
complete_sets <- SetAssessment %>% 
  filter(curve == "Complete") 
list_complete <- (unique(complete_sets$set))
life_stages <- c("egg", "larva", "pupa", "eggtoemergence")
Ana_performance <- Ana %>% 
  filter(set %in% list_complete) %>% 
  filter(lifestage %in% life_stages)
egg_performance <- Ana_performance %>%
  filter(lifestage == "egg")
larva_performance <- Ana_performance %>%
  filter(lifestage == "larva")
pupa_performance <- Ana_performance %>%
  filter(lifestage == "pupa")
eggtoemergence_performance <- Ana_performance %>%
  filter(lifestage == "eggtoemergence")

Ana_performance$lifestage <- factor(Ana_performance$lifestage, levels = c("egg", "larva", "pupa", "eggtoemergence"))
```

There are `r length(unique(Ana_performance$set))` sets of the relevant lifestages (`r length(unique(Ana_performance$sp))` species). Of those, `r length(unique(egg_performance$set))` correspond to egg (N = `r length(unique(egg_performance$sp))` species), `r length(unique(larva_performance$set))` to larva (N = `r length(unique(larva_performance$sp))` species), `r length(unique(pupa_performance$set))`to pupa (N = `r length(unique(pupa_performance$sp))` species) and `r length(unique(eggtoemergence_performance$set))` to eggtoadult (N = `r length(unique(eggtoemergence_performance$sp))`)

3. Consolidate redundant sets
Redundant sets occur beacuse researchers varied experimental conditions such as host plant, humidity and daylength, or because data were reported for males and females separately. We considered multiple measurements reported in the same publication that were obtained at the same temperature, of the same ontogenetic stage of individuals of the same population (sp and locality) to be redundant, and we obtained the geometric mean of all treatments (host, sex, daylength, humidity,etc) at the same temperature for analyses. 
Consolidated sets kept the number (unique set identifier) of the first set they include.


```{r}
# make a new column that unifies redundant sets
Ana_performance <- Ana_performance %>% 
  mutate(redundancy_id = as.numeric(factor(paste(sp, locality, lifestage, id))))

#length(unique(Ana_performance$set)) # there were 156 sets
#length(unique(Ana_performance$redundancy_id)) # there were 137 non-redundant sets

# Function to get the geometric mean of development rate (dr_c), and mean (dr2)
mean_redundant <- function(dat){
newset <- dat$set [[1]]
dat %>%
  select(temp, sp, P) %>% 
  group_by(temp) %>% 
  summarise(performance_c = geometric.mean(P),
            performance2 = mean(P), 
            newset = newset,
            n_sets = length(P)) 
}

# Apply function to Ana_performance

Ana_performance_consolidated <- Ana_performance %>% 
  group_by(redundancy_id) %>% 
  nest() %>% 
  mutate(consolidated = map(data, mean_redundant)) %>% 
  select(redundancy_id, consolidated) %>% 
  unnest(col = consolidated)# Function

# re-name variables. Note that consolidated sets summarize multiple sets and got assigned the number (set id value) of the first set.
names(Ana_performance_consolidated) <- c("redundancy_id", "temp", "P", "P2", "set", "n_sets")
```


4. Obtain predicted performance in the range from 0 to 40C using linear interpolation. 

```{r}
#create curves: 1- Make interpolation function:
interpolate <- function(dat) {
  out <- seq(from = 0, to = 40, by = 0.1)
  as.data.frame(approx(x = dat$temp, y = dat$P, xout = out, method = "linear", rule = 1))
  
}

Predicted_Ana_performance <-  Ana_performance_consolidated %>% 
  group_by(set) %>% 
  nest() %>% 
  mutate(predicted = map(data,interpolate)) %>% 
  select(set, predicted) %>% 
  unnest(cols = predicted)
names(Predicted_Ana_performance) <- c("set", "temp", "P")
```



6. Calculate Popt (temperatures that maximizes performance)

```{r}
# function to get optimum

get_Popt <- function(dat){
  dat <- dat[!is.na(dat$P) & !is.na(dat$temp),] # Remove treatments with no dr data
   y_max <- max(dat$P, na.rm = T) # get maximum value for dr
  x_max_list <- dat[dat$P == y_max, "temp"][[1]]# get temperature that maximizes dr   
  if(length(x_max_list) > 1){
    Popt <- tibble(PTopt = x_max_list,
                   Propt = y_max, 
                   Opt = "optima")
  } else{
    Popt <- tibble(PTopt = x_max_list,
                   Propt = y_max, 
                   Opt = "optimum")
  }
} 


# apply function to all sets
Popts <- Predicted_Ana_performance %>% 
  group_by(set) %>% 
  nest() %>% 
  mutate(Popt = map(data,get_Popt)) %>% 
  select(set, Popt)

# get a table with only one optimum temperature
singles <- Popts %>% 
  unnest(cols = c(Popt)) %>% 
  filter(Opt == "optimum")


# get a table with sets with multiple optima
multiple <- Popts %>% 
  unnest(cols = c(Popt)) %>% 
  filter(Opt == "optima")
# Only one optimum,yay!!
```



```{r}

# 1- Get values from "singles" table:


add_info <- Ana_performance %>% 
  select(set, sp, family, lifestage, lat, lon, locality, quality) %>% 
  distinct()

Popt_table <- inner_join(singles, add_info, by = "set")

```

#7. Calculate P-interval (min, max, length)


```{r}


get_interval <- function(dat){
  dat <- dat[!is.na(dat$P) & !is.na(dat$temp),] # Remove treatments with no P data
  y_max <- max(dat$P) # get maximum value for dr
  x_max <- dat[dat$P == y_max, "temp"][[1]][[1]]# get temperature that maximizes P
  cold <- filter(dat, temp < x_max)  
  hot <- filter(dat, temp > x_max)  
  limit <- y_max/2
  print(limit)
  if (min(cold$P) > limit){
    print("cold incomplete")
    y_mincold <- NA
    x_min_cold <- NA
    grade_cold <- "incomplete"
  } else{
    y_mincold <- min(filter(cold, P >= y_max/2)[["P"]])
    x_min_cold <- min(filter(cold, P >= y_max/2)[["temp"]])
    grade_cold <- "complete"
  }
  
  if(min(hot$P) > limit){
    print("heat incomplete")
    y_minhot <- NA
    x_min_hot <- NA 
    grade_hot <- "incomplete"
  } else {
    y_minhot <- min(filter(hot, P >= y_max/2)[["P"]])
    x_min_hot <- max(filter(hot, P >= y_max/2)[["temp"]])
    grade_hot <- "complete"
  }
  
  
  output <- tibble(Imin = x_min_cold, 
                   Imax = x_min_hot, 
                   Ilength = Imax - Imin, 
                   gradecold = grade_cold, 
                   gradehot = grade_hot)
  
  
}




```


```{r}
# Apply get interval to everything
# apply function to all sets
Interval <- Predicted_Ana_performance %>% 
  group_by(set) %>% 
  nest() %>% 
  mutate(interval = map(data,get_interval)) %>% 
  select(set, interval) %>% 
  unnest(cols = interval)
Interval_table1 <- inner_join(Interval, add_info, by = "set")
Interval_table1$lifestage <- factor(Interval_table1$lifestage, levels = c("egg", "larva", "pupa", "eggtoemergence"))
Interval_table_cold <- filter(Interval_table1, gradecold == "complete")
Interval_table_hot <- filter(Interval_table1, gradehot == "complete")
Interval_table_both <- filter(Interval_table1, gradecold == "complete", gradehot == "complete")
rm(Interval_table1)
```


```{r}
# clean-up so only relevant tables for analysis remain in workspace
#rm(add_info) # columns with sp, lifestage, and lat lon
#rm(Dopts) #list of set and otimum temperatures
#rm(Interval) # table missing columns relevant for analysis
#rm(multiple)

#rm(singles)
#rm(two_optima_table)
#rm(complete_sets)
```

Description of tables relevant for analysis


8. Visualization

Temperature that allows for development rate at maximum efficiency

```{r}
ggplot(Popt_table, aes(x = lifestage, y = PTopt))+
  geom_boxplot()+
  theme_cowplot()+
  ylab("Popt (C)")+
  xlab("Life stage")+
  geom_jitter(mapping = aes(x = lifestage, y = PTopt, col = family))+
  ylim(10,40)
  
```

```{r}

ggplot(filter(Popt_table, quality != "inferred"), aes(x = abs(lat), y = PTopt, col = family))+
  geom_point()+
  theme_cowplot()+
  ylab("Popt")+
  xlab("Absolute latitude")+
  facet_wrap(.~lifestage)
```

```{r}
ggplot(Interval_table_both, aes(x = lifestage, y = Ilength))+
  geom_boxplot()+
  theme_cowplot()+
  ylab("P-interval length")+
  geom_jitter(mapping = aes(x = lifestage, y = Ilength, col = family))+
  ylim(0,40)

```

Lenght of the temperature interval that allows for development rates at >=50% efficiency.  


```{r}
# Summary statistics
summary <-Interval_table_both %>% 
  group_by(lifestage) %>% 
  summarise(median = median(Ilength, na.rm = T), q0.25 =quantile(Ilength,probs =0.25, na.rm = T), q0.75 =quantile(Ilength,probs =0.75, na.rm = T), N = n())#, c(0.25, 0.5, 0.75), type = 1,na.rm = T))

summary
```
 
 
 Median length of the favorable interval (allows for development to occur at 50% efficiency) was `r summary[summary$lifestage == "egg", "median"][[1]] `C for eggs (N = `r summary[summary$lifestage == "egg", "N"][[1]] `), `r summary[summary$lifestage == "larva", "median"][[1]] `C for larvae (N = `r summary[summary$lifestage == "egg", "median"][[1]] `), `r summary[summary$lifestage == "pupa", "N"][[1]] `C for pupae (N = `r summary[summary$lifestage == "egg", "N"][[1]] `), and `r summary[summary$lifestage == "eggtoemergence", "median"][[1]] ` for the whole lifecycle (egg to emergence, N = `r summary[summary$lifestage == "egg", "N"][[1]] `)
 
```{r}

ggplot(filter(Interval_table_both, quality != "inferred"), aes(x = abs(lat), y = Ilength, col = family))+
  geom_point()+
  theme_cowplot()+
  ylab("D-interval length")+
  
  facet_wrap(.~lifestage)
```

```{r}
ggplot(Interval_table_cold, aes(x = lifestage, y = Imin))+
  geom_boxplot()+
  theme_cowplot()+
  ylab("P-interval minimum (C)")+
  geom_jitter(mapping = aes(x = lifestage, y = Imin, col = family))+
  ylim(0,40)
```
Minimum temperature that allows for development at >= 50% efficiency

```{r}
ggplot(filter(Interval_table_cold, quality != "inferred"), aes(x = abs(lat), y = Imin), col = "quality")+
  geom_point()+
  theme_cowplot()+
  ylab("D-interval minimum")+
  #ylim(10,40)+
  facet_wrap(.~lifestage)
```





```{r}
ggplot(Interval_table_hot, aes(x = lifestage, y = Imax))+
 geom_boxplot()+
  theme_cowplot()+
  ylab("D-interval max")+
  geom_jitter(mapping = aes(x = lifestage, y = Imax, col = family))+
  ylim(10,40)
```
Maximum temperature that allows for development at >= 50% efficiency.

```{r}
ggplot(filter(Interval_table_cold, quality != "inferred"), aes(x = abs(lat), y = Imin), col = "quality")+
  geom_point()+
  theme_cowplot()+
  ylab("D-interval maximum")+
  facet_wrap(.~lifestage)
```



7. Perform analyses

Popt. For analysis I discarded "eggtoemergence" because it integrates patterns of eggs, larvae and pupae. I also discarded sets in which the locality was "inferred" instead of reported by authors.

```{r}
library(lme4)
library(car)
hist(Popt_table$PTopt)
shapiro.test(Popt_table$PTopt)
PoptAnalysis <- filter(Popt_table, lifestage != "eggtoemergence", quality!= "inferred")

PoptAnalysis$absLat <- abs(PoptAnalysis$lat)

fit1 <- lmer(PTopt ~ lifestage * absLat + (1|sp), data = PoptAnalysis) #
summary(fit1)
Anova(fit1) # 
plot(fit1) #

```

Optimum performance of all life stages occurred between `r min(PoptAnalysis$PTopt, na.rm = T)` and `r max(PoptAnalysis$PTopt, na.rm = T)`C. The mean tempeature allowing for optimum development was `r mean(PoptAnalysis$PTopt, na.rm = T)` sd = `r sd(PoptAnalysis$PTopt, na.rm = T)`. N = `r length(unique(PoptAnalysis$set))` consolidated sets of `r length(unique(PoptAnalysis$sp))` species.

Absoulte latitude, but not lifestage explained variation in Popt.


```{r}

library(effects)
library(ggeffects)
PoptAnalysis$group <- PoptAnalysis$lifestage
pred <- ggpredict(fit1, terms = c("absLat","lifestage"))
pred2 <- Effect(c("absLat","lifestage"),fit1)
#plot(pred2)
ggplot(pred, aes(x = x, y = predicted, colour = group, fill = group)) +
  geom_line()+
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, col = group), alpha = 0.1)+
  geom_point(data = PoptAnalysis, mapping = aes(x = abs(lat), y = PTopt, col =group ))+
  scale_color_viridis_d()+
  scale_fill_viridis_d()+
  ylab("Popt")+
  xlab("Absolute latitude")+
  #ylim(15, 40)+
  theme_cowplot()#+
  #facet_grid(.~group)




```
Observed (points) and predicted (line, shaded interval) variation of Popt (temperature that allows for optimum performance) according to latitude and lifestage



```{r}
# P-Interval lower limit
IntervalAnalysis_cold <- filter(Interval_table_cold, lifestage != "eggtoemergence", quality!= "inferred")
IntervalAnalysis_cold$absLat <- abs(IntervalAnalysis_cold$lat)

IntervalAnalysis_cold$absLat <- abs(IntervalAnalysis_cold$lat)


fit2 <- lmer(Imin ~ lifestage * absLat + (1|sp), data = IntervalAnalysis_cold) # 
summary(fit2)
Anova(fit2) # 
plot(fit2) #
```

The lower limit of P-interval ranged from `r min(IntervalAnalysis_cold$Imin, na.rm = T)` and `r max(IntervalAnalysis_cold$Imin, na.rm = T)`C. The mean tempeature allowing for optimum performance was `r mean(IntervalAnalysis_cold$Imin, na.rm = T)` sd = `r sd(PoptAnalysis$PTopt, na.rm = T)`. N = `r length(unique(IntervalAnalysis_cold$set))` consolidated sets of `r length(unique(IntervalAnalysis_cold$sp))` species.



```{r}
# P-interval, lower limit


library(effects)
library(ggeffects)
pred <- ggpredict(fit2, terms = c("absLat","lifestage"))
IntervalAnalysis_cold$group <- IntervalAnalysis_cold$lifestage

ggplot(pred, aes(x = x, y = predicted, colour = group, fill = group)) +
  geom_line()+
   geom_ribbon(aes(ymin = conf.low, ymax = conf.high, col = group), alpha = 0.1)+
  scale_color_viridis_d()+
  scale_fill_viridis_d()+
  ylab("P-interval lower limit")+
  xlab("Absolute latitude")+
 # ylim(5, 20)+
  geom_point(data = IntervalAnalysis_cold, mapping = aes(x = abs(lat), y = Imin, col =group ), alpha = 0.4)+
  theme_cowplot()




```

Observed (points) and predicted (line, shaded interval) variation of the minimum temperature that allows for >= 50% performance efficiency according to latitude and lifestage.

```{r}
# P-interval upper limit
IntervalAnalysis_hot <- filter(Interval_table_hot, lifestage != "eggtoemergence", quality!= "inferred")
IntervalAnalysis_hot$absLat <- abs(IntervalAnalysis_hot$lat)
IntervalAnalysis_hot$absLat <- abs(IntervalAnalysis_hot$lat)
hist(IntervalAnalysis_hot$Imax)
shapiro.test((IntervalAnalysis_hot$Imax))
fit3 <- lmer(Imax ~ lifestage * absLat + (1|sp), data = IntervalAnalysis_hot) # 
summary(fit3)
Anova(fit3) # 
plot(fit3) #
```

The upper limit of P-interval ranged from `r min(IntervalAnalysis_hot$Imax, na.rm = T)` and `r max(IntervalAnalysis_hot$Imax, na.rm = T)`C. The mean tempeature allowing for optimum development was `r mean(IntervalAnalysis_hot$Imax, na.rm = T)` sd = `r sd(IntervalAnalysis_hot$Imax, na.rm = T)`. N = `r length(unique(IntervalAnalysis_hot$set))` consolidated sets of `r length(unique(IntervalAnalysis_hot$sp))` species.The maximum temperature vary across life stages but not latitudes.

```{r}
pred <- ggpredict(fit3, terms = c("absLat","lifestage"))
IntervalAnalysis_hot$group <- IntervalAnalysis_hot$lifestage
ggplot(pred, aes(x = x, y = predicted, colour = group, fill = group)) +
  theme_cowplot()+
  geom_line()+
   geom_ribbon(aes(ymin = conf.low, ymax = conf.high, col = group), alpha = 0.1)+
  geom_point(data = IntervalAnalysis_hot, mapping = aes(x = absLat, y = Imax, col = lifestage))+
  ylab("P-interval upper limit")+
  scale_color_viridis_d(alpha = 0.5)+
  xlab("Absolute latitude")+
  #geom_jitter(mapping = aes(x = lifestage, y = maxTDopt, col = family))+
  ylim(20,40)
```

Observed (points) variation of the maximum temperature that allows for performance at >= 50% efficiency. Neither absolute latitude nor lifestage explained the observed variation.
This temperature is where D-interval starts

```{r}
#Ilength
IntervalAnalysis_both <- filter(Interval_table_both, lifestage != "eggtoemergence", quality!= "inferred")
IntervalAnalysis_both$absLat <- abs(IntervalAnalysis_both$lat)


hist(IntervalAnalysis_both$Ilength)
shapiro.test(IntervalAnalysis_both$Ilength)
fit4 <- lmer(Ilength ~ lifestage + absLat + (1|sp), data = IntervalAnalysis_both) # 
summary(fit4)
Anova(fit4) # 
plot(fit4) #
```

```{r}
IntervalAnalysis_both$group <- IntervalAnalysis_both$lifestage
ggplot(IntervalAnalysis_both, aes(x = absLat, y = Ilength, col = lifestage))+
 geom_point()+
  theme_cowplot()+
  ylab("P-interval length")+
  scale_color_viridis_d(alpha = 0.5)+
  xlab("Absolute latitude")#+
  #geom_jitter(mapping = aes(x = lifestage, y = maxTDopt, col = family))+
  #ylim(0,40)

```

Length of the interval of temperatures that allowed for development at >=50% efficiency. interval length varied from min `r min(IntervalAnalysis_both$Ilength, na.rm = T)`  to max `r max(IntervalAnalysis_both$Ilength, na.rm = T)`, mean = `r mean(IntervalAnalysis_both$Ilength, na.rm = T)`, SD = `r sd(IntervalAnalysis_both$Ilength, na.rm = T)` N = `r length(unique(IntervalAnalysis_both$set))` sets,  `r length(unique(IntervalAnalysis_both$sp))`species


```{r}
Intervar <- select(IntervalAnalysis_both, set, Ilength, Imin, Imax)

allvalues <- full_join(Intervar,PoptAnalysis, by = "set")
ggplot(allvalues, aes(x = absLat, y = Propt, col = lifestage))+
 geom_point()+
  theme_cowplot()+
  ylab("Maximum performance rate")+
  scale_color_viridis_d(alpha = 0.5)+
  xlab("Absolute latitude")
  
  

```



