---
title: "Calculate survival responses"
output: html_notebook
---

We used r (tydiverse) for data wrangling, (ggplot2) for data visualization, (psych) for obtaining geometric mean, (readxl) to import data and (cowplot) to format ggplots.

```{r}
# load packages
library(tidyverse)
library(readxl)
library(cowplot)
library(psych)
```

Import data 
```{r}
# Mariana's desktop
Data <- read_xlsx("/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/PhysiologyDatabaseVersion5.xlsx", 
                  sheet = "WorkingTable", na = c("NA", ""))
```

```{r}
# Mariana's laptop
Data <- read_xlsx("/Users/mar/Desktop/Projects/ThermalPerformance/PhysiologyDatabaseVersion5.xlsx", 
                  sheet = "WorkingTable", na = c("NA", "")) 

```


```{r}
# Anna's Laptop
Data <- read_xlsx("~/Desktop/PhysiologyDatabaseVersion5.xlsx", 
                  sheet = "WorkingTable", na = c("NA", "")) 

```


```{r}
# Discard parasitoids and change character to factor 
unique(Data$status)
Ana <- Data[Data$status != "parasitoid",]
rm(Data)

# convert character to factor
Ana <- Ana %>%
  mutate_if(is.character, factor)
```

# 2. Select sets 
To be included in survival analyses, sets should have at least 4 temperature treatments reporting survival and form a complete curve (rise, peak, fall). We restricted analyses to four life stages: egg, larva, pupa and egg to adult.

```{r}
# Function to determine whether a set is:
# "Complete" (has low, peak and high values)
# "Incomplete" (only low and peak)
# "No data" (no survival data available)
Ana$set2 <- Ana$set
is_complete <- function(dat){
  dat <- dat[!is.na(dat$survival) & !is.na(dat$temp),] # Remove treatments with no survival data
 if(nrow(dat) > 3 ){
  
   y_max <- max(dat$survival, na.rm = T) # get maximum value for survival
  x_max <- dat[dat$survival == y_max, "temp"][[1]][1]# get temperature that maximizes survival 
  
  species <- unique(dat$sp)
  lifestage <- unique(dat$lifestage)
  seti <- unique(dat$set2)
  print(paste("set:",seti))
  #print(species)
  #print(lifestage)
  #print(x_max)
  #print(y_max)
  cold <- filter(dat, temp < x_max)  
  hot <- filter(dat, temp > x_max)  
  # if(sum(cold$survival >=0) && sum(hot$survival >=0)){
  #     print("Complete")
  #   } else {(print("Incomplete"))}
  result = nrow(cold[!is.na(cold$survival),]) && nrow(hot[!is.na(hot$survival),])
  if(result){
    print("Complete")
  } else {
    print("Incomplete")
  }
 } else{
   print("Less than 4 data points")
 }
  
  
}


```


```{r}
# Apply is_complete() to all sets
SetAssessment <- Ana %>% 
  group_by(set) %>% 
  nest() %>% 
  mutate(curve = map(data,is_complete)) %>% 
  select(set, curve) %>% 
  unnest() 

```

From a total of `r length(unique(Ana$set))`, only  `r nrow(filter(SetAssessment, curve == "Complete"))` had a complete curve for survival. There were `r nrow(filter(SetAssessment, curve == "Incomplete"))` incomplete sets and `r nrow(filter(SetAssessment, curve == "Less than 4 data points"))` with less than four data points.

```{r}
# Select sets of life stages of interest: egg, larva, pupa, egg to emergence
complete_sets <- SetAssessment %>% 
  filter(curve == "Complete") 
list_complete <- (unique(complete_sets$set))
life_stages <- c("egg", "larva", "pupa", "eggtoemergence")
Ana_survival <- Ana %>% 
  filter(set %in% list_complete) %>%  # this is to get only complete sets
  filter(lifestage %in% life_stages) # and only the life stages of interest
egg_survival <- Ana_survival %>%
  filter(lifestage == "egg")
larva_survival <- Ana_survival %>%
  filter(lifestage == "larva")
pupa_survival <- Ana_survival %>%
  filter(lifestage == "pupa")
eggtoemergence_survival <- Ana_survival %>%
  filter(lifestage == "eggtoemergence")

Ana_survival$lifestage <- factor(Ana_survival$lifestage, levels = c("egg", "larva", "pupa", "eggtoemergence"))
```
There are `r length(unique(Ana_survival$set))` sets of the relevant lifestages (`r length(unique(Ana_survival$sp))` species). Of those, `r length(unique(egg_survival$set))` correspond to egg (N = `r length(unique(egg_survival$sp))` species), `r length(unique(larva_survival$set))` to larva (N = `r length(unique(larva_survival$sp))` species), `r length(unique(pupa_survival$set))`to pupa (N = `r length(unique(pupa_survival$sp))` species) and `r length(unique(eggtoemergence_survival$set))` to eggtoadult (N = `r length(unique(eggtoemergence_survival$sp))`)

3. Consolidate redundant sets
Redundant sets occur beacuse researchers varied experimental conditions such as host plant, humidity and daylength, or because data were reported for males and females separately. We considered multiple measurements reported in the same publication that were obtained at the same temperature, of the same ontogenetic stage of individuals of the same population (sp and locality) to be redundant, and we obtained the geometric mean of all treatments (host, sex, daylength, humidity,etc) at the same temperature for analyses. 
Consolidated sets kept the number (unique set identifier) of the first set they include.


```{r}
# make a new column that unifies redundant sets
Ana_survival <- Ana_survival %>% 
  mutate(redundancy_id = as.numeric(factor(paste(sp, locality, lifestage, id))))

#length(unique(Ana_survival$set)) # there were 156 sets
#length(unique(Ana_survival$redundancy_id)) # there were 137 non-redundant sets

# Function to get the geometric mean of development rate (dr_c), and mean (dr2)
mean_redundant <- function(dat){
newset <- dat$set [[1]]
dat %>%
  select(temp, sp, survival) %>% 
  group_by(temp) %>% 
  summarise(survival_c = geometric.mean(survival),
            survival2 = mean(survival), 
            newset = newset,
            n_sets = length(survival)) 
}

# Apply function to Ana_survival

Ana_survival_consolidated <- Ana_survival %>% 
  group_by(redundancy_id) %>% 
  nest() %>% 
  mutate(consolidated = map(data, mean_redundant)) %>% 
  select(redundancy_id, consolidated) %>% 
  unnest(col = consolidated)# Function

# re-name variables. Note that consolidated sets summarize multiple sets and got assigned the number (set id value) of the first set.
names(Ana_survival_consolidated) <- c("redundancy_id", "temp", "survival", "survival2", "set", "n_sets")
```


4. Obtain predicted survival in the range from 0 to 40C using linear interpolation. 

```{r}
#create curves: 1- Make interpolation function:
interpolate <- function(dat) {
  out <- seq(from = 0, to = 40, by = 0.1)
  as.data.frame(approx(x = dat$temp, y = dat$survival, xout = out, method = "linear", rule = 1))
  
}

Predicted_Ana_survival <-  Ana_survival_consolidated %>% 
  group_by(set) %>% 
  nest() %>% 
  mutate(predicted = map(data,interpolate)) %>% 
  select(set, predicted) %>% 
  unnest(cols = predicted)
names(Predicted_Ana_survival) <- c("set", "temp", "survival")
```


5. Calculate minSopt and  maxSopt
```{r}
# function to get optimum
dat <- filter(Ana_survival, set == 3)
get_Sopt <- function(dat){
  dat <- dat[!is.na(dat$survival) & !is.na(dat$temp),] # Remove treatments with no survival data
   y_max <- max(dat$survival, na.rm = T) # get maximum value for survival
  x_max_list <- tail(dat[dat$survival == y_max, "temp"], 1)[[1]]# 
  x_min_list <- head(dat[dat$survival == y_max, "temp"], 1)[[1]]# 
    
  Soptima <- tibble(maxSopt = x_max_list,
                   minSopt = x_min_list)

} 


# apply function to all sets
Sopts <- Predicted_Ana_survival %>% 
  group_by(set) %>% 
  nest() %>% 
  mutate(Sopt = map(data,get_Sopt)) %>% 
  select(set, Sopt) %>% 
  unnest(cols = c(Sopt))

add_info <- Ana_survival %>% 
  select(set, sp, family, lifestage, lat, lon, locality, quality) %>% 
  distinct()

Sopts_table <- inner_join(Sopts, add_info, by = "set")

```

6. Calculate interval (min 50% dr, max 50% dr, length)

```{r}

get_interval <- function(dat){
  dat <- dat[!is.na(dat$survival) & !is.na(dat$temp),] # Remove treatments with no survival data
  y_max <- max(dat$survival) # get maximum value for survival
  x_max <- dat[dat$survival == y_max, "temp"][[1]][[1]]# get temperature that maximizes survival
  cold <- filter(dat, temp < x_max)  
  hot <- filter(dat, temp > x_max)  
  limit <- y_max/2
  if (min(cold$survival) > limit){
    print("cold incomplete")
    y_mincold <- NA
    x_min_cold <- NA
    grade_cold <- "incomplete"
  } else{
    y_mincold <- min(filter(cold, survival >= y_max/2)[["survival"]])
    x_min_cold <- min(filter(cold, survival >= y_max/2)[["temp"]])
    grade_cold <- "complete"
  }
  
  if(min(hot$survival) > limit){
    print("heat incomplete")
    y_minhot <- NA
    x_min_hot <- NA 
    grade_hot <- "incomplete"
  } else {
    y_minhot <- min(filter(hot, survival >= y_max/2)[["survival"]])
    x_min_hot <- max(filter(hot, survival >= y_max/2)[["temp"]])
    grade_hot <- "complete"
  }
  
  
  output <- tibble(Imin = x_min_cold, 
                   Imax = x_min_hot, 
                   Ilength = Imax - Imin, 
                   gradecold = grade_cold, 
                   gradehot = grade_hot)
  
  
}


```


```{r}

# apply function to all sets

Interval <- Predicted_Ana_survival %>% 
  group_by(set) %>% 
  nest() %>% 
  mutate(interval = map(data,get_interval)) %>% 
  select(set, interval) %>% 
  unnest(cols = interval)

add_info <- Ana_survival %>% 
  select(set, sp, family, lifestage, lat, lon, locality, quality) %>% 
  distinct()
Interval_table1 <- inner_join(Interval, add_info, by = "set")
Interval_table1$lifestage <- factor(Interval_table1$lifestage, levels = c("egg", "larva", "pupa", "eggtoemergence"))
Interval_table_cold <- filter(Interval_table1, gradecold == "complete")
Interval_table_hot <- filter(Interval_table1, gradehot == "complete")
Interval_table_both <- filter(Interval_table1, gradecold == "complete", gradehot == "complete")
rm(Interval_table1)
```
Note there are three "Interval tables": Interval_table_cold: table with values for the lower end of S-Interval (some incomplete curves)
Interval_table_hot: table with values for the upper end of S-Interval (some incomplete curves)
Interval_table_both: table with values for both lower and upper ends of S-Interval (only sets with complete curves)

Save tables for analysis: 
Sopts_table: maxSopt, minSopt

Interval_table_cold: table with values for the lower end of S-Interval (some incomplete curves)
Interval_table_hot: table with values for the upper end of S-Interval (some incomplete curves)
Interval_table_both: table with values for both lower and upper ends of S-Interval (only sets with complete curves)

```{r}
# 
# write_csv(Sopts_table,"/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/Sopt.csv")
# write_csv(Sopts_table,"/Users/mar/Desktop/Projects/ThermalPerformance/Sopt.csv")
# write_csv(Interval_table_cold,"/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/llS-width.csv")
# write_csv(Interval_table_cold,"/Users/mar/Desktop/Projects/ThermalPerformance/llS-width.csv")
# write_csv(Interval_table_hot,"/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/ulS-width.csv")
# write_csv(Interval_table_hot,"/Users/mar/Desktop/Projects/ThermalPerformance/ulS-width.csv")
# write_csv(Interval_table_both,"/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/S-width.csv")
# write_csv(Interval_table_both,"/Users/mar/Desktop/Projects/ThermalPerformance/S-width.csv")

```

Merge all survival variables tables in one table

```{r}
names(Interval_table_cold) 
```

```{r}
merge1 <- select(Interval_table_cold, set, Slower = Imin)
merge2 <- select(Interval_table_hot, set, Supper = Imax)
merge3 <- select(Interval_table_both, set, Swidth = Ilength)

Survival1 <- full_join(Sopts_table, merge1, by = "set")
Survival2 <- full_join(Survival1, merge2, by = "set")
Survival <- full_join(Survival2, merge3, by = "set")
head(Survival)

```


```{r}
# write_csv(Survival,"/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/Survival.csv")
```


