---
title: "Analyze survival variables"
output: html_notebook
---
## Contents
1. Import data and housekeeping
2. Models with lifestage within species (lifestage|sp) as a random effect
3. Models with species (1|sp) as a random effect
4. Correlation coefficients of survival variables
5. PCA

```{r}
library(tidyverse)
library(readxl)
library(cowplot)
library(psych) # to get geometric mean
library(lme4)
library(car)
library("dotwhisker")  # to make coefficient plots
library(sjPlot) #prediction plots
library(sjmisc)
library(broom)
library(effects)
library(ggeffects)
library(Hmisc)

```

# 1. Import data and housekeeping
```{r}
# Mariana's desktop

responses <- read_csv("/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/ThermalResponsesII.csv")

```


```{r}
# convert character to factor
responses <- responses %>%
  mutate_if(is.character, factor)
# discard non-pertinent data

responses_table <- responses %>% 
  filter(lifestage == "egg"| lifestage == "larva"| lifestage == "pupa", quality!= "inferred", quality != "combination", !is.na(maxSopt)) %>% 
  select(Sopt = maxSopt, Slower, Supper, Swidth, lifestage, sp, lat, family) %>% 
  mutate(absLat = abs(lat))
responses_table$lifestage <- factor(responses_table$lifestage)
responses_table$sp <- factor(responses_table$sp)

```

## Sopt: temperature that maximizes development rate
Sopt ranged from `r min(responses_table$Sopt, na.rm = T)` to `r max(responses_table$Sopt)`, mean =  `r mean(responses_table$Sopt)`, SD = `r sd(responses_table$Sopt)`
Latitude ranged from `r min(responses_table$absLat, na.rm = T)` to `r max(responses_table$absLat, na.rm = T)`

#2. Models with lifestage within species (lifestage|sp) as a random effect

```{r}
# Subset to survival metrics and predictors only
S.set <- with(responses_table, cbind(Sopt, Supper, Swidth, Slower, lifestage, absLat, sp))

# Full models, with random intercept, and then random slopes
## random slopes: need to input dataset as dataframe

fit1_fullrandom <- lmer(Sopt ~ lifestage + absLat + lifestage:absLat + (lifestage|sp), data = as.data.frame(S.set)) #

fit1_redrandom <- lmer(Sopt ~ lifestage + absLat + (lifestage|sp), data = as.data.frame(S.set)) #

summary(fit1_fullrandom)
summary(fit1_redrandom)

anova(fit1_fullrandom, fit1_redrandom, test.statistic = "Chisq")
# This model fails to converge:
fit_redrandom <- lmer(Sopt ~ lifestage + absLat + (lifestage|sp), REML = F, data = as.data.frame(S.set))
# This model does not fail to converge:
fit_redrandom <- lmer(Sopt ~ lifestage + absLat + (lifestage|sp), REML = T, data = as.data.frame(S.set))
# Why??



## IF we  use log-likelihood instead of restricted maximum likelihood:
summary(fit1_fullrandom)
Anova(fit1_fullrandom)
S.set <- as.data.frame(S.set)
meanS <- S.set %>% 
  select(Sopt, absLat) %>% 
  ungroup() %>% 
  filter(absLat > 30 & absLat < 50) %>% 
  summarise(mean = mean(Sopt, na.rm = T), 
            sd = sd(Sopt, na.rm = T))
#plot.design(Dopt ~ lifestage *  absLat + (lifestage|sp) , data = D.set)
dw_plot(fit1_fullrandom) +
  #geom_vline(xintercept= mean(D.set$Dopt),lty=2)+
  geom_vline(xintercept= 0,lty=3)+
  theme_cowplot()+
  scale_color_viridis_d()+
  theme(legend.position = "none")
#type = c("est", "re", "eff", "pred", "int", "std", "std2", "slope", "resid", "diag")

# lifestage is numeric in S.set, so this plot does not work
plot_model(fit1_fullrandom, type = "int")+
  theme_cowplot()+
  scale_color_viridis_d()
```

# 3. Models with species (1|sp) as a random effect
 
```{r}
# models with species (1|species) as a random effect
# 1. Subset to S metrics and predictors only

S.set <- select(responses_table, Sopt, Supper, Swidth, Slower, lifestage, absLat, sp, family)
S.noctu <- S.set %>% 
  filter(family == "Noctuidae")
S.allelse <- S.set %>% 
  filter(family != "Noctuidae")
  

# Full model

full <- lmer(Sopt ~ lifestage + absLat + lifestage:absLat + (1|sp), data = as.data.frame(S.set)) #
latlife <- lmer(Sopt ~ absLat + lifestage + (1|sp), data = as.data.frame(S.set)) #
lat <- lmer(Sopt ~ absLat + (1|sp), data = as.data.frame(S.set)) #
life <- lmer(Sopt ~ lifestage + (1|sp), data = as.data.frame(S.set)) #

anova(full, latlife, lat, life, test.statistic = "Chisq")

latlifeI <- lmer(Sopt ~ -1 + absLat + lifestage + (1|sp), data = as.data.frame(S.set)) #


latlifeI <- lmer(Sopt ~ -1 + absLat + lifestage + (1|sp), REML = F, data = as.data.frame(S.set)) #
latlifeN <- lmer(Sopt ~ -1 + absLat + lifestage + (1|sp), REML = F, data = as.data.frame(S.noctu)) #
latlifeothers <- lmer(Sopt ~ -1 + absLat + lifestage + (1|sp), REML = F, data = as.data.frame(S.allelse)) #



Anova(full)
summary(latlifeI)
summary(latlifeI)
```
 
```{r}
# get data for plot
latlifeI_effects <- tidy(latlifeI)
Esizes_wide <- tibble(o_mean = mean(S.set$Sopt),
                   lat = filter(latlifeI_effects, term == "absLat")[[2]], 
                   lat10 = filter(latlifeI_effects, term == "absLat")[[2]]*10, 
                   egg = filter(latlifeI_effects, term == "lifestageegg")[[2]] - o_mean, 
                   larva = filter(latlifeI_effects, term == "lifestagelarva")[[2]] - o_mean, 
                   pupa = filter(latlifeI_effects, term == "lifestagepupa")[[2]] - o_mean, 
                   SD_sp = filter(latlifeI_effects, term == "sd_(Intercept).sp")[[2]])

names(Esizes_wide)

Esizes <- pivot_longer(Esizes_wide, c("o_mean", "lat", "lat10", "egg", "larva", "pupa", "SD_sp"), names_to = "effect", values_to = "degrees")
Esizes$effect <- factor(Esizes$effect, levels = c("lat", "lat10", "egg", "larva", "pupa", "SD_sp"))

ggplot(filter(Esizes, effect != "o_mean"), aes(x = effect, y = degrees))+
  geom_bar(stat = "identity")+
  theme_cowplot()+
  ylim(-2,6)


```
 
 
```{r}
# Plot data and predictions
library(effects)
library(ggeffects)

S.set$group <- S.set$lifestage
S.noctu$group <- S.noctu$lifestage
S.allelse$group <- S.allelse$lifestage
pred <- ggpredict(latlifeI, terms = c("absLat","lifestage"))
prednoctu <- ggpredict(latlifeN, terms = c("absLat","lifestage"))
predelse <- ggpredict(latlifeothers, terms = c("absLat","lifestage"))


ggplot(predelse, aes(x = x, y = predicted, colour = group, fill = group)) +
  geom_line()+
   geom_ribbon(aes(ymin = conf.low, ymax = conf.high, col = group), alpha = 0.1)+
  scale_color_viridis_d()+
  scale_fill_viridis_d()+
  ylab("Sopt")+
  xlab("Absolute latitude")+
 # ylim(5, 20)+
  geom_point(data = S.set, mapping = aes(x = absLat, y = Sopt, col =group ), alpha = 0.6)+
  theme_cowplot()


```
 
 
```{r}
#type = c("est", "re", "eff", "pred", "int", "std", "std2", "slope", "resid", "diag")
plot_model(latlifeI, type = "pred", terms = c("lifestage", "absLat"))+
  theme_cowplot()+
  scale_color_viridis_d()
```
 
# 4. Correlation coefficients of survival variables

```{r}
names(S.set)
cor(S.set[1:4])
rcorr(as.matrix(S.set[1:4])) # correlation coefficients and then P values
```


 
# 5. PCA of survival variables

```{r}
## Omit NA rows for PCA. need dataframe for input
S.set <- with(responses_table, cbind(Sopt, Supper, Swidth, Slower, lifestage, absLat, sp))
S.set.nona <- as.data.frame(na.omit(S.set))
Spc <- princomp(~Sopt + Supper + Swidth, data = S.set.nona)
summary(Spc)
```


```{r}
loadings(Spc)
```


```{r}
names(Spc)
## Bind PC scores to dataset

S.set.pcs <- cbind(S.set.nona, Spc$scores)
head(S.set.pcs)

## Now model Comp.1 (PC scores on 1st PC)
fit1_red1 <- lmer(Comp.1 ~ lifestage + absLat + (1|sp), data = S.set.pcs)
summary(fit1_red1)

## test simpler models
fit1_red2 <- lmer(Comp.1 ~ lifestage  + (1|sp), data = S.set.pcs)
fit1_red3 <- lmer(Comp.1 ~ absLat + (1|sp), data = S.set.pcs)
fit1_red4 <- lmer(Comp.1 ~ (1|sp), data = S.set.pcs)
anova(fit1_red1,fit1_red2, fit1_red3,fit1_red4, test.statistic = "Chisq")

## estimate parameters using ML rather than REML (default)
fit1_red3.ML <- lmer(Comp.1 ~  absLat  + (1|sp), data = S.set.pcs, REML = F)
summary(fit1_red3.ML)

dw_plot(fit1_red3.ML) +
  #geom_vline(xintercept= mean(D.set$Dopt),lty=2)+
  geom_vline(xintercept= 0,lty=3)+
  theme_cowplot()+
  scale_color_viridis_d()+
  theme(legend.position = "none")


```


# 6. Supper

```{r}
S.set <- data.frame(S.set)
ggplot(S.set, aes(x = Supper, fill = lifestage))+
  geom_histogram()+
  facet_grid(.~lifestage)
hist(S.set$Supper)
median(S.set$Supper, na.rm = T)
```



```{r}
# models with species (1|species) as a random effect
# 1. Subset to S metrics and predictors only

S.set <- select(responses_table, Sopt, Supper, Swidth, Slower, lifestage, absLat, sp)

# Full model

full <- lmer(Supper ~ lifestage + absLat + lifestage:absLat + (1|sp), data = as.data.frame(S.set)) #
latlife <- lmer(Supper ~ absLat + lifestage + (1|sp), data = as.data.frame(S.set)) #
lat <- lmer(Supper ~ absLat + (1|sp), data = as.data.frame(S.set)) #
life <- lmer(Supper ~ lifestage + (1|sp), data = as.data.frame(S.set)) #

anova(full, latlife, lat, life, test.statistic = "Chisq")

latlifeI <- lmer(Supper ~ -1 + absLat + lifestage + (1|sp), data = as.data.frame(S.set)) #


latlifeI <- lmer(Supper ~ -1 + absLat + lifestage + (1|sp), REML = F, data = as.data.frame(S.set)) #

Anova(full)
summary(latlife)
summary(latlifeI)
```
 
```{r}
# get data for plot
latlifeI_effects <- tidy(latlifeI)
Esizes_wide <- tibble(o_mean = mean(S.set$Supper, na.rm = T),
                   lat = filter(latlifeI_effects, term == "absLat")[[2]], 
                   lat10 = filter(latlifeI_effects, term == "absLat")[[2]]*10, 
                   egg = filter(latlifeI_effects, term == "lifestageegg")[[2]] - o_mean, 
                   larva = filter(latlifeI_effects, term == "lifestagelarva")[[2]] - o_mean, 
                   pupa = filter(latlifeI_effects, term == "lifestagepupa")[[2]] - o_mean, 
                   SD_sp = filter(latlifeI_effects, term == "sd_(Intercept).sp")[[2]])

names(Esizes_wide)

Esizes <- pivot_longer(Esizes_wide, c("o_mean", "lat", "lat10", "egg", "larva", "pupa", "SD_sp"), names_to = "effect", values_to = "degrees")
Esizes$effect <- factor(Esizes$effect, levels = c("lat", "lat10", "egg", "larva", "pupa", "SD_sp"))

ggplot(filter(Esizes, effect != "o_mean"), aes(x = effect, y = degrees))+
  geom_bar(stat = "identity")+
  theme_cowplot()


```
 
 
```{r}
# Plot data and predictions
S.set$group <- S.set$lifestage
pred <- ggpredict(latlifeI, terms = c("absLat","lifestage"))


ggplot(pred, aes(x = x, y = predicted, colour = group, fill = group)) +
  geom_line()+
   geom_ribbon(aes(ymin = conf.low, ymax = conf.high, col = group), alpha = 0.1)+
  scale_color_viridis_d()+
  scale_fill_viridis_d()+
  ylab("Supper")+
  xlab("Absolute latitude")+
 # ylim(5, 20)+
  geom_point(data = S.set, mapping = aes(x = absLat, y = Supper, col =group ), alpha = 0.6)+
  theme_cowplot()


```
 
 
```{r}
#type = c("est", "re", "eff", "pred", "int", "std", "std2", "slope", "resid", "diag")
plot_model(latlifeI, type = "pred", terms = c("lifestage", "absLat"))+
  theme_cowplot()+
  scale_color_viridis_d()
```
 



```{r}
S.set <- data.frame(S.set)
meanSwidth <- S.set %>% 
  select(Swidth, absLat) %>% 
  ungroup() %>% 
  filter(absLat > 30 & absLat < 50) %>% 
  summarise(mean = mean(Swidth, na.rm = T), 
            sd = sd(Swidth, na.rm = T))
```


```{r}
meanSlower <- S.set %>% 
  select(Slower, absLat) %>% 
  ungroup() %>% 
  filter(absLat > 30 & absLat < 50) %>% 
  summarise(mean = mean(Slower, na.rm = T), 
            sd = sd(Slower, na.rm = T))
```
