---
title: "Analyze survival variables"
output: html_notebook
---
## Contents
1. Import data and housekeeping

2. Slow (Slower)
2.1  Model selection
2.2  Implement phylogenetic correction

3. Shigh (Dupper)
3.1  Model selection
3.2  Implement phylogenetic correction

4. Swidth
4.1  Model selection
4.2  Phylogenetic correction

5. Sopt
5.1  Model selection
5.2  Phylogenetic correction

# 1. Import data and housekeeping
```{r}
 
# 
#library(INLA) # install.packages("INLA", repos=c(getOption("repos"), #INLA="https://inla.r-inla-download.org/R/stable"), dep=TRUE)
library(tidyverse)
library(readxl)
library(cowplot)
library(psych) # to get geometric mean
library(lme4)
library(car)
library("dotwhisker")  # to make coefficient plots
library(sjPlot) #prediction plots
library(sjmisc)
library(broom)
library(effects)
library(ggeffects)
library(Hmisc)
library(phyr)
library(ape)
library(nlme)
library(stringr)
library(phylotools)
library(phytools)
library(rr2)
```


```{r}
# Mariana's desktop

responses <- read_csv("/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/ThermalResponsesTable.csv")

tree <- read.tree("GTRG_Unpartitioned_constrain.raxml.bestTree.tre")

```


```{r}
# convert character to factor
responses <- responses %>%
  mutate_if(is.character, factor)
# discard non-pertinent data

responses_table <- responses %>% 
  filter(lifestage == "egg"| lifestage == "larva"| lifestage == "pupa", 
         #quality!= "inferred", 
         quality != "combination", 
         !is.na(maxSopt)) %>% 
  select(set, Sopt = maxSopt, Slower, Supper, Swidth, lifestage, sp, lat, family) %>% 
  mutate(absLat = abs(lat))
responses_table$lifestage <- factor(responses_table$lifestage)
responses_table$sp <- factor(responses_table$sp)
S.set <- select(responses_table, set, Sopt, Supper, Swidth, Slower, lifestage, absLat, sp, family)

# save a list of sets included in analyses
S.sets <- S.set %>% 
  select(set)
# write_csv(S.sets,"/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/sets_survival.csv")

```



#2 Slow 
(Slower)
Slow ranged from `r min(S.set$Slower, na.rm = T)` to `r max(S.set$Slower, na.rm = T)`, mean =  `r mean(S.set$Slower, na.rm = T)`, SD = `r sd(S.set$Slower, na.rm = T)`
Latitude ranged from `r min(S.set$absLat, na.rm = T)` to `r max(S.set$absLat, na.rm = T)`
## 2.1 Model selection

```{r}
Slow_i <- lmer(Slower ~ lifestage + absLat + lifestage:absLat + (1|sp), data = S.set) #
Slow_a <- lmer(Slower ~ lifestage + absLat + lifestage + (1|sp), data = S.set)

anova(Slow_i, Slow_a, test.statistic = "Chisq") # keep interaction model
```


```{r}
Slow_i_ML <- lmer(Slower ~ -1 + absLat + lifestage + lifestage:absLat + (1|sp), REML = F, data = S.set) #

#Anova(Slow_i_ML)
summary(Slow_i_ML)

```
 
 
```{r}
# Plot data and predictions
S.set$group <- S.set$lifestage
predSlow <- ggpredict(Slow_i_ML, terms = c("absLat","lifestage"))


Slow_plot <- ggplot(predSlow, aes(x = x, y = predicted, colour = group, fill = group)) +
  geom_line()+
   geom_ribbon(aes(ymin = conf.low, ymax = conf.high, col = group), alpha = 0.1)+
  scale_fill_viridis_d(option = "plasma", begin = 0, end = 0.8)+
  scale_color_viridis_d(option = "plasma", begin = 0, end = 0.8)+
  ylab(expression(S[low]))+
   annotate("text", x = 0, y = 30, label = "X", size = 5)+
  
  xlab("Absolute latitude")+
  #ylim(0, 40)+
  geom_point(data = S.set, mapping = aes(x = absLat, y = Slower, col =group ), alpha = 0.6, shape = 16)+
  theme_cowplot()+
  theme(legend.position = "none")

Slow_plot
```
 
 
```{r}
#type = c("est", "re", "eff", "pred", "int", "std", "std2", "slope", "resid", "diag")
# plot_model(Slow_i_ML, type = "pred", terms = c("lifestage", "absLat"))+
#  theme_cowplot()+
#  scale_color_viridis_d()
```
 
## 2.2 Phylogenetic correction
###2.2.1 Adjust data table
Change species' names of the following species to reflect substitutions due to lack of data and to correct mispellings

Substitutions:
 In tree                  In dataset    
Ephestia columbiella for     Ephestia calidella #### Not in analyses because source population was "inferred"
Episimus tyrius for          Episimus utilis 
Euzopherodes allocrossa for Euzopherodes vapidella ### No development rate data available
Marmara arbutiella for      Marmara gulosa
Protodeltote albidula for   Naranga aenescens

Mispellings: 
Hyphantria cunea instead of   Hypantria cunea
Ameyelois transitella instead of Amyelois transitella,
```{r}
S.set <- S.set
S.set$original_species <- S.set$sp
S.set$sp <- ifelse(S.set$original_species == "Episimus utilis", "Episimus tyrius", ifelse(S.set$original_species == "Marmara gulosa", "Marmara arbutiella", ifelse(S.set$original_species == "Naranga aenescens", "Protodeltote albidula", ifelse(S.set$original_species == "Hypantria cunea", "Hyphantria cunea",  ifelse(S.set$original_species == "Ameyelois transitella", "Amyelois transitella", ifelse(S.set$original_species == "Utethesia ornatrix", "Utetheisa ornatrix",              ifelse(S.set$original_species == "Euzopherodes vapidella", "Euzopherodes allocrossa", ifelse(S.set$original_species == "Ephestia calidella","Ephestia columbiella", as.character(S.set$original_species)))))))))

Sspecies <- tibble(sp = unique(S.set$sp))
tree_species <-  tibble(sp = as.character(paste(word(tree$tip.label, sep = "_", 4), word(tree$tip.label, sep = "_", 5))))
V_species <- unique(S.set$sp)
# setdiff(Sspecies, tree_species)
# all species in S.set are in the tree. 

# setdiff(tree_species, Sspecies) # the tree has 64 species more than S.set

```
Trim the tree so it only includes the species in the P.set. First the tips have to be relabeled so they include only the species names
```{r}
rename <- as.data.frame(tibble(oldname = tree$tip.label, 
                 newname = tree_species$sp))
tree_relabel <- sub.taxa.label(tree, rename)
S.tree <- keep.tip(tree_relabel, V_species)
plotTree(S.tree)
```




```{r}
Slow_phylo <- pglmm(Slower ~ -1 + lifestage + absLat + lifestage:absLat + (1|sp__),bayes= F, data = S.set, cov_ranef = list(sp = S.tree), REML = F)

Slow_nophylo <- pglmm(Slower ~ -1 + lifestage + absLat + lifestage:absLat +  (1|sp),bayes= F, data = S.set, cov_ranef = list(sp = S.tree), REML = F)


summary(Slow_phylo)
```

Assess the effect of phylogeny
```{r}
R2(Slow_phylo,Slow_nophylo, phy = S.tree) # 
 
```


Extract effect sizes
```{r}
SlowFixedEffects <- tibble(Response = "Slow", Term = row.names(Slow_phylo$B), Effect = Slow_phylo$B[1:6], 
SE = Slow_phylo$B.se[1:6])

SlowRandomEffects <- tibble(Response = "Slow", Term = c("sp", "sp__", "residual"), Variance = c(Slow_phylo$s2r, Slow_phylo$s2resid) , 
SD = sqrt(Variance))
```


Consolidate effects and save table
```{r}
SFixedEffects <- rbind(SoptFixedEffects, SwidthFixedEffects, ShighFixedEffects, SlowFixedEffects)
SRandomEffects <- rbind(SoptRandomEffects, SwidthRandomEffects, ShighRandomEffects, SlowRandomEffects)
# write_csv(SFixedEffects,"/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/SFixedEffects.csv")
# write_csv(SRandomEffects,"/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/SRandomEffects.csv")
```


## 3 Shigh (Supper)
Supper ranged from `r min(S.set$Supper, na.rm = T)` to `r max(S.set$Supper)`, mean =  `r mean(S.set$Supper)`, SD = `r sd(S.set$Supper)`
Latitude ranged from `r min(S.set$absLat, na.rm = T)` to `r max(S.set$absLat, na.rm = T)`
### 3.1 Model selection

```{r}
Shigh_i <- lmer(Supper ~ lifestage + absLat + lifestage:absLat + (1|sp), data = S.set) #
Shigh_a<- lmer(Supper ~ lifestage + absLat + lifestage + (1|sp), data = S.set)

anova(Shigh_i, Shigh_a, test.statistic = "Chisq") # keep aditive model
```

```{r}
Shigh_a_ML <- lmer(Supper ~ -1 + absLat + lifestage + (1|sp), REML = F, data = S.set) #

#Anova(Shigh_a_ML)
summary(Shigh_a_ML)
```
 
 
```{r}
# Plot data and predictions
S.set$group <- S.set$lifestage
predSh <- ggpredict(Shigh_i, terms = c("absLat","lifestage"))


ggplot(predSh, aes(x = x, y = predicted, colour = group, fill = group)) +
  geom_line()+
   geom_ribbon(aes(ymin = conf.low, ymax = conf.high, col = group), alpha = 0.1)+
  scale_fill_viridis_d(option = "plasma", begin = 0, end = 0.8)+
  scale_color_viridis_d(option = "plasma", begin = 0, end = 0.8)+
  ylab(expression(S[high]))+#ylab("Supper")+
  xlab("Absolute latitude")+
  ylim(0, 40)+
  geom_point(data = S.set, mapping = aes(x = absLat, y = Supper, col =group ), alpha = 0.6, shape = 1)+
  theme_cowplot()+
  theme(legend.position = "none")

```
 
### 3.2 Phylogenetic correction

```{r}
Shigh_phylo <- pglmm(Supper ~ -1 + lifestage + absLat +  (1|sp__),bayes= F, data = S.set, cov_ranef = list(sp = S.tree), REML = F)

Shigh_nophylo <- pglmm(Supper ~ -1 + lifestage + absLat +  (1|sp),bayes= F, data = S.set, cov_ranef = list(sp = S.tree), REML = F)
```

Assess effect of phylogeny
```{r}
R2(Shigh_phylo,Shigh_nophylo, phy = S.tree) # 
 
```

```{r}
summary(Shigh_phylo)
```

Extract effect sizes
```{r}
ShighFixedEffects <- tibble(Response = "Shigh", Term = row.names(Shigh_phylo$B), Effect = Shigh_phylo$B[1:4], 
SE = Shigh_phylo$B.se[1:4])



ShighRandomEffects <- tibble(Response = "Shigh", Term = c("sp", "sp__", "residual"), Variance = c(Shigh_phylo$s2r, Shigh_phylo$s2resid) , 
SD = sqrt(Variance))

```


##4 Swidth
Swidth ranged from `r min(S.set$Swidth, na.rm = T)` to `r max(S.set$Swidth)`, mean =  `r mean(S.set$Swidth, na.rm = T)`, SD = `r sd(S.set$Swidth, na.rm = T)`
Latitude ranged from `r min(S.set$absLat, na.rm = T)` to `r max(S.set$absLat, na.rm = T)`

###4.1 Model selection

```{r}
Swidth_i <- lmer(Swidth ~ lifestage + absLat + lifestage:absLat + (1|sp), data = S.set) #
Swidth_a<- lmer(Swidth ~ lifestage + absLat + lifestage + (1|sp), data = S.set)

anova(Swidth_i, Swidth_a, test.statistic = "Chisq") # keep additive model
```

```{r}
Swidth_a_ML <- lmer(Swidth ~ -1 + absLat + lifestage +  (1|sp), REML = F, data = S.set) #

#Anova(Swidth_a_ML)
summary(Swidth_a_ML)

```
 

 
```{r}
# Plot data and predictions
S.set$group <- S.set$lifestage
predsw <- ggpredict(Swidth_i, terms = c("absLat","lifestage"))

ggplot(predsw, aes(x = x, y = predicted, colour = group, fill = group)) +
  geom_line()+
   geom_ribbon(aes(ymin = conf.low, ymax = conf.high, col = group), alpha = 0.1)+
  scale_fill_viridis_d(option = "plasma", begin = 0, end = 0.8)+
  scale_color_viridis_d(option = "plasma", begin = 0, end = 0.8)+
  ylab(expression(S[width]))+ #ylab("Swidth")+
  xlab("Absolute latitude")+
  annotate("text", x = 0, y = 40, label = "X", size = 5)+
  ylim(0, 40)+
  geom_point(data = S.set, mapping = aes(x = absLat, y = Swidth, col =group ), alpha = 0.6, shape = 0)+
  theme_cowplot()+
  theme(legend.position = "none")

```
 
 
```{r}
# type = c("est", "re", "eff", "pred", "int", "std", "std2", "slope", "resid", "diag")
# plot_model(Swidth_a_ML, type = "pred", terms = c("lifestage", "absLat"))+
#  theme_cowplot()+
#  scale_color_viridis_d()
```
 
### 3.2 Phylogenetic correction

```{r}
Swidth_phylo <- pglmm(Swidth ~ -1 + lifestage + absLat +  (1|sp__), bayes = F, data = S.set, cov_ranef = list(sp = S.tree), REML = F)

Swidth_nophylo <- pglmm(Swidth ~ -1 + lifestage + absLat +  (1|sp), bayes = F, data = S.set, cov_ranef = list(sp = S.tree), REML = F)

```

```{r}

R2(Swidth_phylo,Swidth_nophylo) # if this does not work unload and reload rr2

```
Extract effect sizes
```{r}
SwidthFixedEffects <- tibble(Response = "Swidth", Term = row.names(Swidth_phylo$B), Effect = Swidth_phylo$B[1:4], 
SE = Swidth_phylo$B.se[1:4])

SwidthRandomEffects <- tibble(Response = "Swidth", Term = c("sp", "sp__", "residual"), Variance = c(Swidth_phylo$s2r,Swidth_phylo$s2resid) , 
SD = sqrt(Variance))

```



All in one figures


```{r}

PanelC <- ggplot(predSlow, aes(x = x, y = predicted, colour = group, fill = group)) +
  geom_line()+
  #geom_line(data = predsw, mapping = aes(x = x, y = predicted, colour = group, fill = group))+
  geom_line(data = predSh, mapping = aes(x = x, y = predicted, colour = group, fill = group))+
   #geom_ribbon(aes(ymin = conf.low, ymax = conf.high, col = group), alpha = 0.1)+
  scale_color_viridis_d(option = "plasma", begin = 0, end = 0.8)+
  scale_fill_viridis_d()+
  #scale_shape_manual(values = c(1,16,8))+
 #ylab(expression(S[low]-S[high]))+
  annotate("text", x = 59, y = 15, label = expression(S[low]))+
  annotate("text", x = 59, y = 35, label = expression(S[high]))+
  
# ylab("Temperature")+
 ylab("")+
  #xlab("Absolute latitude")+
  xlab("")+
  ylim(0, 40)+
  annotate("text", x = 0, y = 40, label = "C", size = 5)+
  geom_point(data = S.set, mapping = aes(x = absLat, y = Slower, col = group), alpha = 0.6, shape =16)+
  #geom_point(data = S.set, mapping = aes(x = absLat, y = Swidth, col = group), alpha = 0.6, shape = 6)+
  geom_point(data = S.set, mapping = aes(x = absLat, y = Supper, col = group), alpha = 0.6, shape = 1)+
  theme_cowplot()+
  theme(legend.title = element_blank(), legend.position = "none")
PanelC
```


```{r}
PanelD <- ggplot(predsw, aes(x = x, y = predicted, colour = group, fill = group)) +
  geom_line()+
  #geom_line(data = predso, mapping = aes(x = x, y = predicted), colour = "grey")+
  #geom_line(data = predSh, mapping = aes(x = x, y = predicted, colour = group, fill = group))+
   #geom_ribbon(aes(ymin = conf.low, ymax = conf.high, col = group), alpha = 0.1)+
  scale_color_viridis_d(option = "plasma", begin = 0, end = 0.8)+
  scale_fill_viridis_d()+
  #scale_shape_manual(values = c(1,16,8))+
  annotate("text", x = 59, y = 10, label = expression(S[width]))+
 ylab("Temperature")+
  xlab("Absolute latitude")+
  ylim(0, 30)+
  annotate("text", x = 0, y = 30, label = "D", size = 5)+
  #geom_point(data = S.set, mapping = aes(x = absLat, y = Slower, col = group), alpha = 0.6, shape =16)+
  #geom_point(data = S.set, mapping = aes(x = absLat, y = Sopt), alpha = 0.6, shape = 1, col = "grey")+
  geom_point(data = S.set, mapping = aes(x = absLat, y = Swidth, col = group), alpha = 0.6, shape = 0)+
  theme_cowplot()+
  theme(legend.title = element_blank(), legend.position = "none")
PanelD
```
Save data to make figures:
predSh
predSlow
predsw
S.set
```{r}
#write_csv(predSh,"/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/Output/predSh.csv")
#write_csv(predSlow,"/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/Output/predSlow.csv")
#write_csv(S.set,"/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/Output/Sset.csv")
#write_csv(predsw,"/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/Output/predsw.csv")
```


 
 # Old code below
 ********************************************
 
 
 



## Sopt: temperature that maximizes development rate
Sopt ranged from `r min(S.set$Sopt, na.rm = T)` to `r max(S.set$Sopt)`, mean =  `r mean(S.set$Sopt)`, SD = `r sd(S.set$Sopt)`
Latitude ranged from `r min(S.set$absLat, na.rm = T)` to `r max(S.set$absLat, na.rm = T)`,  N =`r length(!is.na(S.set$Sopt))`

# 2.1 Model selection

```{r}
Sopt_i <- lmer(Sopt ~ -1 + lifestage + absLat + lifestage:absLat + (1|sp), data = S.set) #singular model
Sopt_a <- lmer(Sopt ~ -1 + lifestage + absLat + (1|sp), data = S.set) #

anova(Sopt_i, Sopt_a, test.statistic = "Chisq") # keep additive model
```

```{r}
Sopt_a_ML <- lmer(Sopt ~ -1 + absLat + lifestage  + (1|sp), data = S.set)

summary(Sopt_a_ML)
```
According to this model, Sopt increased across development, as it was 28.6 for eggs, 30.2 for larvae and 30.5 for pupae. Sopt decreased 1.5 C per every 10 degrees latitude. Variation among species was small: 1.8

 
 
```{r}
# Plot data and predictions
library(effects)
library(ggeffects)

S.set$group <- S.set$lifestage
predso <- ggpredict(Sopt_i_ML, terms = c("absLat","lifestage"))



Sopt_plot <- ggplot(predso, aes(x = x, y = predicted, colour = group, fill = group, shape = group)) +
  geom_line()+
   geom_ribbon(aes(ymin = conf.low, ymax = conf.high, col = group), alpha = 0.1)+
  scale_color_viridis_d()+
  scale_fill_viridis_d()+
  scale_shape_manual(values = c(1,16,8))+
  ylab(expression(S[opt]))+
  annotate("text", x = 0, y = 40, label = "B", size = 5)+
  xlab("Absolute latitude")+
  ylim(10, 40)+
  geom_point(data = S.set, mapping = aes(x = absLat, y = Sopt, col =group ), alpha = 0.6)+
  theme_cowplot()+
  theme(legend.position = "none")

Sopt_plot
```
 
 
```{r}
# #type = c("est", "re", "eff", "pred", "int", "std", "std2", "slope", "resid", "diag")
# plot_model(Sopt_a_ML, type = "pred", terms = c("lifestage", "absLat"))+
#   theme_cowplot()+
#   scale_color_viridis_d()
```

# Fit phylogenetic correction

```{r}
# With phylogenetic correction
Sopt_phylo <- pglmm(Sopt ~ -1 + lifestage + absLat + (1|sp__), bayes= F, data = S.set, cov_ranef = list(sp = S.tree), REML =F)

# Without phylogenetic correction
Sopt_nophylo <- pglmm(Sopt ~ -1 + lifestage + absLat  +  (1|sp), bayes= F, data = S.set, cov_ranef = list(sp = S.tree), REML =F)

```



Asses the effect of phylogeny on Sopt
```{r}
R2(Sopt_phylo,Sopt_nophylo, phy = S.tree) # if this does not work unload and reload rr2
```

Extract effect sizes for plot
```{r}
SoptFixedEffects <- tibble(Response = "Sopt", Term = row.names(Sopt_phylo$B), Effect = Sopt_phylo$B[1:4], 
SE = Sopt_phylo$B.se[1:4])

SoptRandomEffects <- tibble(Response = "Sopt", Term = c("sp", "sp__", "residual"), Variance = c(Sopt_phylo$s2r,Sopt_phylo$s2resid) , 
SD = sqrt(Variance))

```



# 4. Correlation coefficients of survival variables

```{r}
names(S.set)
cor(S.set[1:4])
rcorr(as.matrix(S.set[1:4])) # correlation coefficients and then P values
```


 
# 5. PCA of survival variables

```{r}
## Omit NA rows for PCA. need dataframe for input
S.set <- with(responses_table, cbind(Sopt, Supper, Swidth, Slower, lifestage, absLat, sp))
S.set.nona <- as.data.frame(na.omit(S.set))
Spc <- princomp(~Sopt + Supper + Swidth, data = S.set.nona)
summary(Spc)
```


```{r}
loadings(Spc)
```


```{r}
names(Spc)
## Bind PC scores to dataset

S.set.pcs <- cbind(S.set.nona, Spc$scores)
head(S.set.pcs)

## Now model Comp.1 (PC scores on 1st PC)
fit1_red1 <- lmer(Comp.1 ~ lifestage + absLat + (1|sp), data = S.set.pcs)
summary(fit1_red1)

## test simpler models
fit1_red2 <- lmer(Comp.1 ~ lifestage  + (1|sp), data = S.set.pcs)
fit1_red3 <- lmer(Comp.1 ~ absLat + (1|sp), data = S.set.pcs)
fit1_red4 <- lmer(Comp.1 ~ (1|sp), data = S.set.pcs)
anova(fit1_red1,fit1_red2, fit1_red3,fit1_red4, test.statistic = "Chisq")

## estimate parameters using ML rather than REML (default)
fit1_red3.ML <- lmer(Comp.1 ~  absLat  + (1|sp), data = S.set.pcs, REML = F)
summary(fit1_red3.ML)

# dw_plot(fit1_red3.ML) +
#   #geom_vline(xintercept= mean(D.set$Dopt),lty=2)+
#   geom_vline(xintercept= 0,lty=3)+
#   theme_cowplot()+
#   scale_color_viridis_d()+
#   theme(legend.position = "none")


```

