---
title: "Analyze survival variables"
output: html_notebook
---
## Contents
1. Import data and housekeeping

```{r}
# load packages
library(tidyverse)
library(readxl)
library(cowplot)
library(psych) # to get geometric mean
library(lme4)
library(car)
library("dotwhisker")  # to make coefficient plots
library(sjPlot) #prediction plots
library(sjmisc)
library(broom)
```
1. Import file
```{r}
# Mariana's desktop

responses <- read_csv("/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/ThermalResponsesII.csv")

```


```{r}
# convert character to factor
responses <- responses %>%
  mutate_if(is.character, factor)
# discard non-pertinent data

responses_table <- responses %>% 
  filter(lifestage != "eggtoemergence", quality!= "inferred", quality != "combination") %>% 
  mutate(absLat = abs(lat))
```


```{r}
names(responses_table)
# 1. Subset to D metrics and predictors only
S.set <- with(responses_table, cbind(maxSopt, Swidth, Supper, lifestage, absLat, sp))

# Full models, with random intercept, and then random slopes

## random slopes: need to input dataset as dataframe

fit1_fullrandom <- lmer(maxSopt ~ lifestage + absLat + lifestage:absLat + (lifestage|sp), data = as.data.frame(S.set)) #

summary(fit1_fullrandom)

fit_red1random <- lmer(maxSopt ~ lifestage + absLat + (lifestage|sp), data = as.data.frame(S.set))

anova(fit1_fullrandom, fit_red1random, test.statistic = "Chisq")
## interaction term is worse fit by AIC ro BIC, use reduced model

# fit better model with ML
fit1_randomML <- lmer(maxSopt ~ lifestage + absLat + (lifestage|sp), REML = F, data = as.data.frame(S.set))
summary(fit1_randomML)
Anova(fit1_randomML)
S.set <- as.data.frame(S.set)

dw_plot(fit1_randomML) +
  #geom_vline(xintercept= mean(D.set$Dopt),lty=2)+
  geom_vline(xintercept= 0,lty=3)+
  theme_cowplot()+
  scale_color_viridis_d()+
  theme(legend.position = "none")


```



```{r}
## Omit NA rows for PCA. need dataframe for input
S.set.nona <- as.data.frame(na.omit(S.set))
Spc <- princomp(~maxSopt + Supper + Swidth, data = S.set.nona)
summary(Spc)
loadings(Spc)
names(Spc)

## Bind PC scores to dataset

S.set.pcs <- cbind(S.set.nona, Spc$scores)
head(S.set.pcs)

## Now model Comp.1 (PC scores on 1st PC)
fit1_red1 <- lmer(Comp.1 ~ lifestage + absLat + (lifestage|sp), data = as.data.frame(S.set.pcs))
# not working! 
```












Initially, there were `r length(unique(Sopt$set))` sets with values for Sopt, `r length(unique(Swidth$set))` sets with values for D-width. For `r length(unique(lower$set))` sets only the lower limit of D-width could be calculated and in  for `r length(unique(upper$set))` cases only the upper limit could be calculated.


```{r}
# discard non-pertinent data
unique(Sopt$quality)
Sopt_table <- Sopt %>% 
  filter(lifestage != "eggtoemergence", quality!= "inferred", quality != "combination") %>% 
  mutate(absLat = abs(lat))

lower_table <- lower %>% 
  filter(lifestage != "eggtoemergence", quality!= "inferred", quality != "combination") %>% 
  mutate(absLat = abs(lat))

upper_table <- upper %>% 
  filter(lifestage != "eggtoemergence", quality!= "inferred", quality != "combination") %>% 
  mutate(absLat = abs(lat))

Swidth_table <- Swidth %>% 
  filter(lifestage != "eggtoemergence", quality!= "inferred", quality != "combination") %>% 
  mutate(absLat = abs(lat))
#rm(Sopt, lower, upper, Swidth)

```
To assess variation in thermal responses across latitude and life stages, we fitted a linear mixed models, including each thermal response scaled and life and laitude as fixed factors and species as a random factor. We discarded "eggtoemergence" category in the subsequent analyses because it is a combination of the other three.

We first fitted a full model including latitude, life stage, and their interaction as fixed effects and species as a random effect. We also fitted a reduced model without the interaction term and used AIC to select between them. 

## Sopt: temperature that maximizes development rate
Sopt ranged from `r min(Sopt_table$maxSopt)` to `r max(Sopt_table$maxSopt)`, mean =  `r mean(Sopt_table$maxSopt)`, SD = `r sd(Sopt_table$maxSopt)`
Latitude ranged from `r min(Sopt_table$absLat, na.rm = T)` to `r max(Sopt_table$absLat, na.rm = T)`


```{r}
# 1. Full model
# Scale response variable
Sopt_table$Sopt_s <- scale(Sopt_table$maxSopt)[,1]

fit1_full <- lmer(Sopt_s ~ -1 + lifestage * absLat + (1|sp), data = Sopt_table) #

# this model cannot be done because we have 134 observations and the number of random effects would be 171
#fit1_full_random <- lmer(Sopt_s ~ -1 + lifestage * absLat + (lifestage|sp), data = Sopt_table) #
# this model is singular
#fit1_full_family <- lmer(Sopt_s ~ -1 + lifestage * absLat + (1|sp) + (1|family), data = Sopt_table) #

#summary(fit1_full)

fit1_full_aic <- AIC(fit1_full)
fit1_fullpvals <- Anova(fit1_full)
# dw_plot(fit1_full) +
#   geom_vline(xintercept=0,lty=2)+
#   theme_cowplot()+
#   scale_color_viridis_d()+
#   theme(legend.position = "none")
# Figure 1. effects and 95% CI of the full model
```
 Full model: lmer(Sopt_s ~ -1 + lifestage * absLat + (1|sp), data = Sopt_table)
 Full model's AIC: `r fit1_full_aic`

```{r}
#2. Reduced model
fit1_reduced <- lmer(Sopt_s ~  -1 + lifestage + absLat + (1|sp), data = Sopt_table) #

# fit1Noctuids <- lmer(Sopt ~ -1 + lifestage + absLat + (1|sp), data = filter(ToptAnalysis, family == "Noctuidae"))
# fit1Not_noctuids <- lmer(Sopt ~ -1 + lifestage + absLat + (1|sp), data = filter(ToptAnalysis, family != "Noctuidae"))
tidy(fit1_reduced)
fit1_summary <- summary(fit1_reduced)
fit1_reduced_aic <- AIC(fit1_reduced)
#Anova(fit1_reduced)
dw_plot(fit1_reduced, conf.level = 0.95) + 
  geom_vline(xintercept=0,lty=2)+
  theme_cowplot()+
  scale_color_viridis_d()+
  theme(legend.position = "none")#+
  #scale_y_discrete(labels = c('sd Observation.Residual','sd species', 
                             # 'Latitude', 'Life stage (pupa)','Life stage (larva)'))


```
Figure 2. Sopt was higher at lower latitudes
Lots of variation among species!
