---
title: "Sample size"
output: html_notebook
---
Extract sample size from sets included in analyses

1. Get ids of sets used in analyses
2. Sample size by treatment by set (list)
3. Sample size by set and number of temperature treatments (table)


```{r}
# load packages
library(tidyverse)
library(readxl)
library(cowplot)
library(psych)
```



```{r}
# table with sets inclduded in analyses
responses <- read_csv("/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/ThermalResponsesII.csv")


# table with relevant columns to extract sample size by treatment
alldata <- read_xlsx("/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/PhysiologyDatabaseVersion5.xlsx", 
                  sheet = "WorkingTable", na = c("NA", ""))
# get set ids
responses_table <- responses %>% 
  filter(lifestage == "egg"| lifestage == "larva"| lifestage == "pupa") %>% 
  mutate(lifestage = factor(lifestage))

sets_table <- responses_table %>% select(sp, family, set) %>% distinct()
set_list <- unique(sets_table$set)

# extract sets in set_list from alldata and keep only relevant columns for sample size estimation
data <- alldata %>% 
  filter(set %in% set_list) %>% 
  select(set, temp, lifestage, n, c1, idaid, c2, c3)


```

  `r length(set_list)` were included in analyses. 


```{r}
# extract N per treatment
# 1. Get list of sets with problems
problem1 <- data[is.na(as.numeric(as.character(data$n))),] #this instruction first changes non-numeric entries to NA and keeps only rows with NA values in sample size
problem_n_list <- unique(problem1$set)
length(unique(problem_n_list)) #154
proper_list <- setdiff(set_list, problem_n_list) #154 sets with problematic sample sizes
length(proper_list) # 237
154+237

#2. Get table of sets with no reporting problems 
proper <- data %>% 
  filter(set %in% proper_list) %>% 
  mutate(n = as.numeric(n))

# 3. Get table of sets with reporting problems
problem <- filter(data, set %in% problem_n_list)

# list problematic values
length(unique(problem1$n))
unique(problem1$n)

```
Replace character sample size with best guess given reports in paper or with NA when no info was reported ("nd")

```{r}
se <- filter(problem, n == "41,61,23")# there were three replicates, so add values
se2 <- filter(problem, n == "34,53,37")# there were three replicates, so add values
se3 <- filter(problem, n == "14,82,36")# there were three replicates, so add values
se4 <- filter(problem, n == "17,11,3")# there were three replicates, so add values
se5 <- filter(problem, n == "44,59,36,40")# there were four replicates, so add values

# Make 25 modifications to data table (NA will not be modified)
problem$n_approx <- ifelse(problem$n == "100-200", 150,
                     ifelse(problem$n == "65-80", 40,
                        ifelse(problem$n == "nd", "NA",
                        ifelse(problem$n == "10to20", 15,
                        ifelse(problem$n == "17,11,3", 41,
                           ifelse(problem$n == "14,82,36", 132,
                           ifelse(problem$n == "44,59,36,40",179,
                           ifelse(problem$n == "41,61,23", 125,
                               ifelse(problem$n == "34,53,37", 124,
                               ifelse(problem$n == "5 replications, 50 to 60 eggs", 275, 
                               ifelse(problem$n == "3 replicates of 50 each", 150, 
                                   ifelse(problem$n == ">30", 30, 
                                   ifelse(problem$n == "~18", 18,
                                   ifelse(problem$n == "8 groups of 50", 400, 
                                      ifelse(problem$n =="8 groups of 15", 120,
                                      ifelse(problem$n == "over 30 individuals per treatment", 30,        
                                      ifelse(problem$n == "~200", 200, 
                                          ifelse(problem$n == "15 pairs of adults", 30,   
                                          ifelse(problem$n == "~20 inds", 20,
                                          ifelse(problem$n == "at least 15", 15, 
                                              ifelse(problem$n == "~2000", 2000, 
                                              ifelse(problem$n == "~60", 60,                                                                                                     ifelse(problem$n == "150 eggs", 150,                                                                                                  ifelse(problem$n == "~20", 20,
                                                 ifelse(problem$n == "~100",100, problem$n)))))))))))))))))))))))))

length(unique(problem$n_approx))
unique(problem$n_approx)
problem$n_approx <- as.numeric(problem$n_approx)

```

2. Sample size by temperature treatment (multiple values per set)

```{r}
# get sample size by temperature for proper sets
proper_treatment1 <-  proper %>% 
  select(set, temp, n) 
ntreats <- proper_treatment1 %>% 
  group_by(set) %>% 
  summarise(treatments = n())
proper_treatment <- left_join(proper_treatment1, ntreats, by = "set")

# get sample size by temperature for problem sets
problem_treatment1 <-  problem %>% 
  select(set, temp, n = n_approx) 
ntreats <- problem_treatment1 %>% 
  group_by(set) %>% 
  summarise(treatments = n())
problem_treatment <- left_join(problem_treatment1, ntreats, by = "set")
# merge tables

n_by_treatment <- rbind(proper_treatment, problem_treatment)
unique(n_by_treatment$n)

```


3. Sample size by set 

```{r}
n_by_set <- n_by_treatment %>% 
  group_by(set) %>% 
  summarise(n_mean_inds = mean(n, na.rm = T),n_total_inds = sum(n, na.rm = T), n_temps = mean(treatments, na.rm = T), max = max(temp,na.rm = T), min =  min(temp, na.rm = T)) %>% 
  mutate(spread = max-min)
```

Print tables
```{r}
#write_csv(n_by_treatment,"/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/NbyTreatment.csv")
#write_csv(n_by_set,"/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/NbySet.csv")
```

Consider consolidated sets? This estimate is conservative compared to one accounting for consolidation. 

