---
title: "Survival"
output: html_notebook
---
## Sections
1. Import data
2. Select sets with at least 4 data points and a complete curve (rise, peak, fall) for egg, larva, pupa, and egg to adult.
3. Consolidate redundant sets.
4. Obtain predicted survival in the range from 0 to 40 C using linear interpolation
5. Calculate minSopt and maxSopt (min and max temperatures that maximize survival)
6. Calculate S-interval (min, max, length)
7. Visualization
8. Analyses

We used r (tydiverse) for data wrangling, (ggplot2) for data visualization, (psych) for obtaining geometric mean

```{r}
# load packages
library(tidyverse)
library(readxl)
library(cowplot)
library(psych)
```

Import data 
```{r}
# Mariana's desktop
Data <- read_xlsx("/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/PhysiologyDatabaseVersion5.xlsx", 
                  sheet = "WorkingTable", na = c("NA", ""))
```

```{r}
# Mariana's laptop
Data <- read_xlsx("/Users/mar/Desktop/Projects/ThermalPerformance/PhysiologyDatabaseVersion5.xlsx", 
                  sheet = "WorkingTable", na = c("NA", "")) 

```


```{r}
# Anna's Laptop
Data <- read_xlsx("~/Desktop/PhysiologyDatabaseVersion5.xlsx", 
                  sheet = "WorkingTable", na = c("NA", "")) 

```


```{r}
# Discard parasitoids and change character to factor 
unique(Data$status)
Ana <- Data[Data$status != "parasitoid",]
rm(Data)

# convert character to factor
Ana <- Ana %>%
  mutate_if(is.character, factor)
```

# 2. Select sets 
To be included in survival analyses, sets should have at least 4 temperature treatments reporting survival and form a complete curve (rise, peak, fall). We restricted analyses to four life stages: egg, larva, pupa and egg to adult.

```{r}
# Function to determine whether a set is:
# "Complete" (has low, peak and high values)
# "Incomplete" (only low and peak)
# "No data" (no survival data available)
Ana$set2 <- Ana$set
is_complete <- function(dat){
  dat <- dat[!is.na(dat$survival) & !is.na(dat$temp),] # Remove treatments with no survival data
 if(nrow(dat) > 3 ){
  
   y_max <- max(dat$survival, na.rm = T) # get maximum value for survival
  x_max <- dat[dat$survival == y_max, "temp"][[1]][1]# get temperature that maximizes survival 
  
  species <- unique(dat$sp)
  lifestage <- unique(dat$lifestage)
  seti <- unique(dat$set2)
  print(paste("set:",seti))
  #print(species)
  #print(lifestage)
  #print(x_max)
  #print(y_max)
  cold <- filter(dat, temp < x_max)  
  hot <- filter(dat, temp > x_max)  
  # if(sum(cold$survival >=0) && sum(hot$survival >=0)){
  #     print("Complete")
  #   } else {(print("Incomplete"))}
  result = nrow(cold[!is.na(cold$survival),]) && nrow(hot[!is.na(hot$survival),])
  if(result){
    print("Complete")
  } else {
    print("Incomplete")
  }
 } else{
   print("Less than 4 data points")
 }
  
  
}


```


```{r}
# Apply is_complete() to all sets
SetAssessment <- Ana %>% 
  group_by(set) %>% 
  nest() %>% 
  mutate(curve = map(data,is_complete)) %>% 
  select(set, curve) %>% 
  unnest() 

```

From a total of `r length(unique(Ana$set))`, only  `r nrow(filter(SetAssessment, curve == "Complete"))` had a complete curve for development rate.
```{r}
# Select sets of life stages of interest: egg, larva, pupa, egg to emergence
complete_sets <- SetAssessment %>% 
  filter(curve == "Complete") 
list_complete <- (unique(complete_sets$set))
life_stages <- c("egg", "larva", "pupa", "eggtoemergence")
Ana_survival <- Ana %>% 
  filter(set %in% list_complete) %>% 
  filter(lifestage %in% life_stages)
egg_survival <- Ana_survival %>%
  filter(lifestage == "egg")
larva_survival <- Ana_survival %>%
  filter(lifestage == "larva")
pupa_survival <- Ana_survival %>%
  filter(lifestage == "pupa")
eggtoemergence_survival <- Ana_survival %>%
  filter(lifestage == "eggtoemergence")

Ana_survival$lifestage <- factor(Ana_survival$lifestage, levels = c("egg", "larva", "pupa", "eggtoemergence"))
```
There are `r length(unique(Ana_survival$set))` sets of the relevant lifestages (`r length(unique(Ana_survival$sp))` species). Of those, `r length(unique(egg_survival$set))` correspond to egg (N = `r length(unique(egg_survival$sp))` species), `r length(unique(larva_survival$set))` to larva (N = `r length(unique(larva_survival$sp))` species), `r length(unique(pupa_survival$set))`to pupa (N = `r length(unique(pupa_survival$sp))` species) and `r length(unique(eggtoemergence_survival$set))` to eggtoadult (N = `r length(unique(eggtoemergence_survival$sp))`)

3. Consolidate redundant sets
Redundant sets occur beacuse researchers varied experimental conditions such as host plant, humidity and daylength, or because data were reported for males and females separately. We considered multiple measurements reported in the same publication that were obtained at the same temperature, of the same ontogenetic stage of individuals of the same population (sp and locality) to be redundant, and we obtained the geometric mean of all treatments (host, sex, daylength, humidity,etc) at the same temperature for analyses. 
Consolidated sets kept the number (unique set identifier) of the first set they include.


```{r}
# make a new column that unifies redundant sets
Ana_survival <- Ana_survival %>% 
  mutate(redundancy_id = as.numeric(factor(paste(sp, locality, lifestage, id))))

#length(unique(Ana_survival$set)) # there were 156 sets
#length(unique(Ana_survival$redundancy_id)) # there were 137 non-redundant sets

# Function to get the geometric mean of development rate (dr_c), and mean (dr2)
mean_redundant <- function(dat){
newset <- dat$set [[1]]
dat %>%
  select(temp, sp, survival) %>% 
  group_by(temp) %>% 
  summarise(survival_c = geometric.mean(survival),
            survival2 = mean(survival), 
            newset = newset,
            n_sets = length(survival)) 
}

# Apply function to Ana_survival

Ana_survival_consolidated <- Ana_survival %>% 
  group_by(redundancy_id) %>% 
  nest() %>% 
  mutate(consolidated = map(data, mean_redundant)) %>% 
  select(redundancy_id, consolidated) %>% 
  unnest(col = consolidated)# Function

# re-name variables. Note that consolidated sets summarize multiple sets and got assigned the number (set id value) of the first set.
names(Ana_survival_consolidated) <- c("redundancy_id", "temp", "survival", "survival2", "set", "n_sets")
```


4. Obtain predicted survival in the range from 0 to 40C using linear interpolation. 

```{r}
#create curves: 1- Make interpolation function:
interpolate <- function(dat) {
  out <- seq(from = 0, to = 40, by = 0.1)
  as.data.frame(approx(x = dat$temp, y = dat$survival, xout = out, method = "linear", rule = 1))
  
}

Predicted_Ana_survival <-  Ana_survival_consolidated %>% 
  group_by(set) %>% 
  nest() %>% 
  mutate(predicted = map(data,interpolate)) %>% 
  select(set, predicted) %>% 
  unnest(cols = predicted)
names(Predicted_Ana_survival) <- c("set", "temp", "survival")
```


5. Calculate minSopt and  maxSopt
```{r}
# function to get optimum
dat <- filter(Ana_survival, set == 3)
get_Sopt <- function(dat){
  dat <- dat[!is.na(dat$survival) & !is.na(dat$temp),] # Remove treatments with no survival data
   y_max <- max(dat$survival, na.rm = T) # get maximum value for survival
  x_max_list <- tail(dat[dat$survival == y_max, "temp"], 1)[[1]]# 
  x_min_list <- head(dat[dat$survival == y_max, "temp"], 1)[[1]]# 
    
  Soptima <- tibble(maxSopt = x_max_list,
                   minSopt = x_min_list)

} 


# apply function to all sets
Sopts <- Predicted_Ana_survival %>% 
  group_by(set) %>% 
  nest() %>% 
  mutate(Sopt = map(data,get_Sopt)) %>% 
  select(set, Sopt) %>% 
  unnest(cols = c(Sopt))

add_info <- Ana_survival %>% 
  select(set, sp, family, lifestage, lat, lon, locality, quality) %>% 
  distinct()

Sopts_table <- inner_join(Sopts, add_info, by = "set")

```

6. Calculate interval (min 50% dr, max 50% dr, length)

```{r}

get_interval <- function(dat){
  dat <- dat[!is.na(dat$survival) & !is.na(dat$temp),] # Remove treatments with no survival data
  y_max <- max(dat$survival) # get maximum value for survival
  x_max <- dat[dat$survival == y_max, "temp"][[1]][[1]]# get temperature that maximizes survival
  cold <- filter(dat, temp < x_max)  
  hot <- filter(dat, temp > x_max)  
  limit <- y_max/2
  if (min(cold$survival) > limit){
    print("cold incomplete")
    y_mincold <- NA
    x_min_cold <- NA
    grade_cold <- "incomplete"
  } else{
    y_mincold <- min(filter(cold, survival >= y_max/2)[["survival"]])
    x_min_cold <- min(filter(cold, survival >= y_max/2)[["temp"]])
    grade_cold <- "complete"
  }
  
  if(min(hot$survival) > limit){
    print("heat incomplete")
    y_minhot <- NA
    x_min_hot <- NA 
    grade_hot <- "incomplete"
  } else {
    y_minhot <- min(filter(hot, survival >= y_max/2)[["survival"]])
    x_min_hot <- max(filter(hot, survival >= y_max/2)[["temp"]])
    grade_hot <- "complete"
  }
  
  
  output <- tibble(Imin = x_min_cold, 
                   Imax = x_min_hot, 
                   Ilength = Imax - Imin, 
                   gradecold = grade_cold, 
                   gradehot = grade_hot)
  
  
}


```


```{r}

# apply function to all sets

Interval <- Predicted_Ana_survival %>% 
  group_by(set) %>% 
  nest() %>% 
  mutate(interval = map(data,get_interval)) %>% 
  select(set, interval) %>% 
  unnest(cols = interval)

add_info <- Ana_survival %>% 
  select(set, sp, family, lifestage, lat, lon, locality, quality) %>% 
  distinct()
Interval_table1 <- inner_join(Interval, add_info, by = "set")
Interval_table1$lifestage <- factor(Interval_table1$lifestage, levels = c("egg", "larva", "pupa", "eggtoemergence"))
Interval_table_cold <- filter(Interval_table1, gradecold == "complete")
Interval_table_hot <- filter(Interval_table1, gradehot == "complete")
Interval_table_both <- filter(Interval_table1, gradecold == "complete", gradehot == "complete")
rm(Interval_table1)
```
Interval_table_cold: table with values for the lower end of S-Interval (some incomplete curves)
Interval_table_hot: table with values for the upper end of S-Interval (some incomplete curves)
Interval_table_both: table with values for both lower and upper ends of S-Interval (only sets with complete curves)

```{r}
# clean-up so only relevant tables for analysis remain in workspace
# rm(add_info) # columns with sp, lifestage, and lat lon
# rm(Sopts) #list of set and otimum temperatures
# rm(Interval) # table missing columns relevant for analysis
# rm(multiple)
# rm(SetAssessment)
# rm(singles)
# rm(two_optima_table)
# rm(complete_sets)
```
Ana: original table
Ana_survival: table with complete sets of relevant lifestages

Interval_table: Imin, Imax and Ilength


6. Plot

```{r}
ggplot(Sopts_table, aes(x = lifestage, y = maxSopt))+
  geom_boxplot()+
  theme_cowplot()+
  ylab("maxSopt")+
  #geom_jitter(mapping = aes(x = lifestage, y = maxTDopt, col = family))+
  ylim(0,40)
  
```
Highest temperature that allows for optimum survival by lifestage

```{r}
ggplot(filter(Sopts_table, quality != "inferred"), aes(x = abs(lat), y = maxSopt))+
  geom_point()+
  theme_cowplot()+
  ylab("maxSopt")+
  xlab("Absolute latitude")+
  #ylim(10,40)+
  facet_wrap(.~lifestage)
```
Highest temperature that allows for optimum survival by absolute latitude

```{r}
ggplot(Sopts_table, aes(x = lifestage, y = minSopt))+
  geom_boxplot()+
  theme_cowplot()+
  ylab("minSopt")+
  #geom_jitter(mapping = aes(x = lifestage, y = maxTDopt, col = family))+
  ylim(0,40)
  
```
Minimum temperature that allows for optimum survival by life stage

```{r}
ggplot(filter(Sopts_table, quality != "inferred"), aes(x = abs(lat), y = minSopt))+
  geom_point()+
  theme_cowplot()+
  ylab("minSopt")+
  xlab("Absolute latitude")+
  #ylim(10,40)+
  facet_wrap(.~lifestage)
```

```{r}
ggplot(Interval_table, aes(x = lifestage, y = Ilength))+
  geom_boxplot()+
  theme_cowplot()+
  ylab("S-interval length")+
  #geom_jitter(mapping = aes(x = lifestage, y = maxTDopt, col = family))+
  ylim(0,40)

```
Length of the temperature interval that allows for >= 50% survival by life stage


```{r}
# Summary statistics
summary <-Interval_table %>% 
  group_by(lifestage) %>% 
  summarise(median = median(Ilength, na.rm = T), q0.25 =quantile(Ilength,probs =0.25, na.rm = T), q0.75 =quantile(Ilength,probs =0.75, na.rm = T), N = n())#, c(0.25, 0.5, 0.75), type = 1,na.rm = T))

summary
```
 
 
 Median length of the favorable interval (allows for development to occur at 50% efficiency) was `r summary[summary$lifestage == "egg", "median"][[1]] `C for eggs (N = `r summary[summary$lifestage == "egg", "N"][[1]] `), `r summary[summary$lifestage == "larva", "median"][[1]] `C for larvae (N = `r summary[summary$lifestage == "egg", "median"][[1]] `), `r summary[summary$lifestage == "pupa", "N"][[1]] `C for pupae (N = `r summary[summary$lifestage == "egg", "N"][[1]] `), and `r summary[summary$lifestage == "eggtoemergence", "median"][[1]] ` for the whole lifecycle (egg to emergence, N = `r summary[summary$lifestage == "egg", "N"][[1]] `)
 
```{r}

ggplot(filter(Interval_table, quality != "inferred"), aes(x = abs(lat), y = Ilength), col = "quality")+
  geom_point()+
  theme_cowplot()+
  ylab("S-interval length")+
  #geom_jitter(mapping = aes(x = lifestage, y = maxTDopt, col = family))+
  #ylim(10,40)+
  facet_wrap(.~lifestage)
```
Length of the temperature interval that allows for >= 50% survival by absolute latitude

```{r}
ggplot(Interval_table, aes(x = lifestage, y = Imin))+
  geom_boxplot()+
  theme_cowplot()+
  ylab("S-interval minimum (C)")+
  #geom_jitter(mapping = aes(x = lifestage, y = maxTDopt, col = family))+
  ylim(0,40)
```
lowest temperature that allows for 50% survival by lifestage

```{r}
ggplot(filter(Interval_table, quality != "inferred"), aes(x = abs(lat), y = Imin), col = "quality")+
  geom_point()+
  theme_cowplot()+
  ylab("S-interval minimum")+
  #geom_jitter(mapping = aes(x = lifestage, y = maxTDopt, col = family))+
  #ylim(10,40)+
  facet_wrap(.~lifestage)
```

Lowest temperature that allows for >=50% survival by absolute latitude



```{r}
ggplot(Interval_table, aes(x = lifestage, y = Imax))+
 geom_boxplot()+
  theme_cowplot()+
  ylab("S-interval max")+
  #geom_jitter(mapping = aes(x = lifestage, y = maxTDopt, col = family))+
  ylim(10,40)
```

Highest temperature that allows for 50% survival by lifestage

```{r}
ggplot(filter(Interval_table, quality != "inferred"), aes(x = abs(lat), y = Imin), col = "quality")+
  geom_point()+
  theme_cowplot()+
  ylab("S-interval maximum")+
  #geom_jitter(mapping = aes(x = lifestage, y = maxTDopt, col = family))+
  #ylim(10,40)+
  facet_wrap(.~lifestage)
```
Highest temperature that allows for 50% survival by absolute latitude



8. Analyses
Sopts_table
Interval_table
```{r}
library(lme4)
library(car)

hist(Sopts_table$maxSopt)
shapiro.test((Sopts_table$maxSopt))
SoptsAnalysis <- filter(Sopts_table, lifestage != "eggtoemergence", quality!= "inferred")

SoptsAnalysis$absLat <- abs(SoptsAnalysis$lat)

fit1 <- lmer(maxSopt ~ lifestage + absLat + (1|sp) + (1|family), data = SoptsAnalysis) #
summary(fit1)
Anova(fit1) # 
plot(fit1) #
```

maxSopt varied from  `r min(SoptsAnalysis$maxSopt, na.rm = T)`  to max `r max(SoptsAnalysis$maxSopt, na.rm = T)`, mean = `r mean(SoptsAnalysis$maxSopt, na.rm = T)`, SD = `r sd(SoptsAnalysis$maxSopt, na.rm = T)`, N = `r length(unique(SoptsAnalysis$set))` sets from `r length(unique(SoptsAnalysis$sp))` species).Lifestage and absolute latitude explained variation in minSoptLifestage and absolute latitude explained variation in maxSopt



```{r}
library(effects)
library(ggeffects)
SoptsAnalysis$group <- SoptsAnalysis$lifestage
pred <- ggpredict(fit1, terms = c("absLat","lifestage"))

#plot(pred2)
ggplot(pred, aes(x = x, y = predicted, colour = group, fill = group)) +
  geom_line()+
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, col = group), alpha = 0.1)+
  geom_point(data = SoptsAnalysis, mapping = aes(x = abs(lat), y = maxSopt, col =group ),alpha = 0.3 )+
  scale_color_viridis_d()+
  scale_fill_viridis_d()+
  ylab("maxSopt")+
  xlab("Absolute latitude")+
  ylim(10, 35)+
  theme_cowplot()#+
  #facet_grid(.~group)
```



```{r}
#minSopt

hist(SoptsAnalysis$minSopt)
shapiro.test(SoptsAnalysis$minSopt)

fit2 <- lmer(minSopt ~ lifestage + absLat + (1|sp) + (1|family), data = SoptsAnalysis) #
summary(fit2)
Anova(fit2) # 
plot(fit2) #
```


minSopt  `r min(SoptsAnalysis$minSopt, na.rm = T)`  to max `r max(SoptsAnalysis$minSopt, na.rm = T)`, mean = `r mean(SoptsAnalysis$minSopt, na.rm = T)`, SD = `r sd(SoptsAnalysis$minSopt, na.rm = T)`, N = `r length(unique(SoptsAnalysis$set))` sets from `r length(unique(SoptsAnalysis$sp))` species).Lifestage and absolute latitude explained variation in minSopt

```{r}

pred <- ggpredict(fit2, terms = c("absLat","lifestage"))

#plot(pred2)
ggplot(pred, aes(x = x, y = predicted, colour = group, fill = group)) +
  geom_line()+
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, col = group), alpha = 0.1)+
  geom_point(data = SoptsAnalysis, mapping = aes(x = abs(lat), y = minSopt, col =group ), alpha = 0.3)+
  scale_color_viridis_d()+
  scale_fill_viridis_d()+
  ylab("minSopt")+
  xlab("Absolute latitude")+
  ylim(10, 35)+
  theme_cowplot()#+
  #facet_grid(.~group)
```



```{r}
#Imin
IntervalAnalysis_cold <- filter(Interval_table_cold, lifestage != "eggtoemergence", quality!= "inferred")

IntervalAnalysis_cold$absLat <- abs(IntervalAnalysis_cold$lat)
IntervalAnalysis_cold$group <- IntervalAnalysis_cold$lifestage
hist(IntervalAnalysis_cold$Imin)
shapiro.test(IntervalAnalysis_cold$Imin)
fit3 <- lmer(Imin ~ lifestage + absLat + (1|sp) + (1|family), data = IntervalAnalysis_cold) #
summary(fit3)
Anova(fit3) # 
plot(fit3) #
```
The lower limit of S-interval (interval of temperatures that allows for at least 50% survival) ranged from   `r min(IntervalAnalysis_cold$Imin, na.rm = T)`  to max `r max(IntervalAnalysis_cold$Imin, na.rm = T)`, mean = `r mean(IntervalAnalysis_cold$Imin, na.rm = T)`, SD = `r sd(IntervalAnalysis_cold$Imin, na.rm = T)`, N = `r length(unique(IntervalAnalysis_cold$set))` sets from `r length(unique(IntervalAnalysis_cold$sp))` species).Lifestage and absolute latitude explained variation in minSoptAbsolute latitude, but not lifestage affected Imin


```{r}
pred <- ggpredict(fit3, terms = c("absLat","lifestage"))

#plot(pred2)
ggplot(pred, aes(x = x, y = predicted, colour = group, fill = group)) +
  geom_line()+
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, col = group), alpha = 0.1)+
  geom_point(data = IntervalAnalysis, mapping = aes(x = abs(lat), y = Imin, col =group ))+
  scale_color_viridis_d()+
  scale_fill_viridis_d()+
  ylab("S-interval lower limit")+
  xlab("Absolute latitude")+
  #ylim(10, 35)+
  theme_cowplot()#+
  #facet_grid(.~group)
```



```{r}
# Imax
IntervalAnalysis_hot <- filter(Interval_table_hot, lifestage != "eggtoemergence", quality!= "inferred")

IntervalAnalysis_hot$absLat <- abs(IntervalAnalysis_hot$lat)
IntervalAnalysis_hot$group <- IntervalAnalysis_hot$lifestage
hist(Interval_table_hot$Imax)
shapiro.test((Interval_table_hot$Imax))
fit4 <- lmer(Imax ~ lifestage + absLat + (1|sp), data = IntervalAnalysis_hot) # 
summary(fit4)
Anova(fit4) # 
plot(fit4) #
```

The upper limit of S-interval (interval of temperatures that allows for at least 50% survival) ranged from   `r min(IntervalAnalysis_hot$Imax, na.rm = T)`  to max `r max(IntervalAnalysis_hot$Imax, na.rm = T)`, mean = `r mean(IntervalAnalysis_hot$Imax, na.rm = T)`, SD = `r sd(IntervalAnalysis_hot$Imax, na.rm = T)`, N = `r length(unique(IntervalAnalysis_hot$set))` sets from `r length(unique(IntervalAnalysis_hot$sp))` species). There is an issue with the model, but it looks like both lifestage and latitude affected variation in Imax
```{r}
pred <- ggpredict(fit4, terms = c("absLat","lifestage"))

#plot(pred2)
ggplot(pred, aes(x = x, y = predicted, colour = group, fill = group)) +
  geom_line()+
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, col = group), alpha = 0.1)+
  geom_point(data = IntervalAnalysis_hot, mapping = aes(x = abs(lat), y = Imax, col =group ))+
  scale_color_viridis_d()+
  scale_fill_viridis_d()+
  ylab("S-Interval upper limit")+
  xlab("Absolute latitude")+
  #ylim(10, 35)+
  theme_cowplot()#+
  #facet_grid(.~group)
```


```{r}
#Ilength
IntervalAnalysis_both <- filter(Interval_table_both, lifestage != "eggtoemergence", quality!= "inferred")

IntervalAnalysis_both$absLat <- abs(IntervalAnalysis_both$lat)
IntervalAnalysis_both$group <- IntervalAnalysis_both$lifestage
hist(IntervalAnalysis_both$Ilength)
shapiro.test((IntervalAnalysis_both$Ilength))
fit5 <- lmer(Ilength ~ lifestage + absLat + (1|sp), data = IntervalAnalysis_both)# 
summary(fit5)
Anova(fit5) # 
plot(fit5) #
```

```{r}
pred <- ggpredict(fit5, terms = c("absLat","lifestage"))

#plot(pred2)
ggplot(pred, aes(x = x, y = predicted, colour = group, fill = group)) +
  geom_line()+
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, col = group), alpha = 0.1)+
  geom_point(data = IntervalAnalysis_both, mapping = aes(x = abs(lat), y = Ilength, col =group ), alpha = 0.3)+
  scale_color_viridis_d()+
  scale_fill_viridis_d()+
  ylab("S-Interval length")+
  xlab("Absolute latitude")+
  #ylim(10, 35)+
  theme_cowplot()#+
  #facet_grid(.~group)
```

Neither absolute latitude or life stage explained variation in the length of the interval of temperatures that allow for >=50% survival.

Length of the interval of temperatures that allowed for >=50% survival. interval length varied from min `r min(IntervalAnalysis_both$Ilength, na.rm = T)`  to max `r max(IntervalAnalysis_both$Ilength, na.rm = T)`, mean = `r mean(IntervalAnalysis_both$Ilength, na.rm = T)`, SD = `r sd(IntervalAnalysis$Ilength, na.rm = T)`
