---
title: "T0 Analyses"
output: html_notebook
---

1. Import data and housekeeping

2.  Model selection
3.  Phylogenetic correction


```{r}
# load packages
# 
#library(INLA) # install.packages('INLA', repos='https://inla.r-inla-download.org/R/stable')
library(tidyverse)
library(readxl)
library(cowplot)
library(psych) # to get geometric mean
library(lme4)
library(car)
library("dotwhisker")  # to make coefficient plots
library(sjPlot) #prediction plots
library(sjmisc)
library(broom)
library(effects)
library(ggeffects)
library(Hmisc)
library(phyr)
library(ape)
library(nlme)
library(stringr)
```


# 1. Import data and housekeeping
```{r}
# Mariana's desktop

responses <- read_csv("/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/ThermalResponsesTable.csv")
splist <- read_csv("/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/SpeciesList.csv")
tree <- read.tree("GTRG_Unpartitioned_constrain.raxml.bestTree.tre")
```


```{r}
# convert character to factor
responses <- responses %>%
  mutate_if(is.character, factor)
# discard non-pertinent data
# convert character to factor
responses <- responses %>%
  mutate_if(is.character, factor)
list <- splist$sp
# discard non-pertinent data
responses_table <- responses %>% 
  filter(lifestage == "egg"| lifestage == "larva"| lifestage == "pupa",   #quality!= "inferred", 
         quality != "combination", sp%in%list, !is.na(T0)) %>% 
  mutate(absLat = abs(lat))
responses_table$lifestage <- factor(responses_table$lifestage)
responses_table$sp <- factor(responses_table$sp)
T.set <- filter(responses_table, !is.na(T0))
T.set <- select(T.set, set, T0, lifestage, absLat, sp, family)
length(unique(T.set$sp))
T.set$genus <- word(T.set$sp, 1)
T.set$species <- word(T.set$sp, 2)

# save a list of sets included in analyses
T.sets <- T.set %>% 
  select(set)
# write_csv(T.sets,"/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/sets_Tnot.csv")
```

# 2 T0 Temperature threshold to accumulate degree-days
T0 ranged from `r min(T.set$T0, na.rm = T)` to `r max(T.set$T0, na.rm = T)`, mean =  `r mean(T.set$T0, na.rm = T)`, SD = `r sd(T.set$T0, na.rm = T)` N =`r length(!is.na(T.set$T0))`
Latitude ranged from `r min(T.set$absLat, na.rm = T)` to `r max(T.set$absLat, na.rm = T)`

## Model selection

```{r}

T0_i <- lmer(T0 ~ -1 + lifestage + absLat + lifestage:absLat + (1|sp), data = T.set) #
T0_a <- lmer(T0 ~ -1 + lifestage + absLat + (1|sp), data = T.set) #

anova(T0_a, T0_i, test.statistic = "Chisq")# keep additive model

```

```{r}
T0_a_ML <- lmer(T0 ~ -1 + lifestage + absLat +  (1|sp), REML = F, data = T.set)
summary(T0_a_ML)
```
 
 
```{r}
# Plot data and predictions

responses_table$group <- responses_table$lifestage
predto <- ggpredict(T0_i, terms = c("absLat","lifestage"))

T.set$group <- T.set$lifestage

T0_plot <- ggplot(predto, aes(x = x, y = predicted, colour = group)) +
  geom_line()+
   geom_ribbon(aes(ymin = conf.low, ymax = conf.high, col = group), alpha = 0.1)+
  scale_color_viridis_d(option = "plasma", begin = 0, end = 0.8)+ # "magma", "inferno", "plasma"
  #scale_fill_viridis_d(option = "magma", limits = c(0.2,0.8))+
  xlim(0,50)+
  ylab(expression('T'[0]))+
  annotate("text", x = 0, y = 16, label = "F", size = 5)+
  xlab("Absolute latitude")+
  geom_point(data = responses_table, mapping = aes(x = absLat, y = T0, col =group ), alpha = 0.6, shape = 6)+
  theme_cowplot()+
  theme(legend.position = "none")

T0_plot
```
 

# 3 Phylogenetic correction
### 3.1 Adjust data table 
Change species' names of the following species to reflect substitutions due to lack of data and to correct mispellings

Substitutions:
 In tree                  In dataset    
Ephestia columbiella for     Ephestia calidella #### Not in analyses because source population was "inferred"
Episimus tyrius for          Episimus utilis 
Euzopherodes allocrossa for Euzopherodes vapidella ### No development rate data available
Marmara arbutiella for      Marmara gulosa
Protodeltote albidula for   Naranga aenescens

Mispellings: 
Hyphantria cunea instead of   Hypantria cunea
Ameyelois transitella instead of Amyelois transitella,
```{r}
T.set$original_species <- T.set$sp
T.set$sp <- ifelse(T.set$original_species == "Episimus utilis", "Episimus tyrius", ifelse(T.set$original_species == "Marmara gulosa", "Marmara arbutiella", ifelse(T.set$original_species == "Naranga aenescens", "Protodeltote albidula", ifelse(T.set$original_species == "Hypantria cunea", "Hyphantria cunea",  ifelse(T.set$original_species == "Ameyelois transitella", "Amyelois transitella", ifelse(T.set$original_species == "Utethesia ornatrix", "Utetheisa ornatrix", ifelse(T.set$original_species == "Ephestia calidella", "Ephestia columbiella",ifelse(T.set$original_species == "Euzopherodes vapidella", "Euzopherodes allocrossa",as.character(T.set$original_species)))))))))

Dspecies <- tibble(sp = unique(T.set$sp))
V_species <- unique(T.set$sp)
tree_species <-  tibble(sp = as.character(paste(word(tree$tip.label, sep = "_", 4), word(tree$tip.label, sep = "_", 5))))

# setdiff(Dspecies, tree_species)
# all species in T.set are in the tree. 

# setdiff(tree_species, Dspecies) # the tree has 45 species more than T.set

```


```{r}
rename <- as.data.frame(tibble(oldname = tree$tip.label, 
                 newname = tree_species$sp))
tree_relabel <- sub.taxa.label(tree, rename)
T.tree <- keep.tip(tree_relabel, V_species)
plotTree(T.tree)
```



### 2.2.2 Fit model

```{r}
# Whit phylogenetic correction
T0_phylo <- pglmm(T0 ~ -1 + lifestage + absLat + (1|sp__),  bayes= F, data = T.set, cov_ranef = list(sp = T.tree), REML = F)


T0_nophylo <- pglmm(T0 ~ -1 + lifestage + absLat + (1|sp),  bayes= F, data = T.set, cov_ranef = list(sp = T.tree), REML = F)
summary(T0_phylo)

```
```{r}
summary(T0_nophylo)
```
Is phylogenetic history important?
```{r}
R2(T0_phylo,T0_nophylo, phy = S.tree) 
```

Write tables for figures
```{r}
#write_csv(predto,"/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/Output/predto.csv")
#write_csv(T.set,"/Users/marianaabarcazama/Desktop/Projects/ThermalPerformance/Output/Tset.csv")
```



# 4. Variation in maximum development rate (exploratory not in presentation)
```{r}
fit2_fullrandom <- lmer(Max_dr ~ lifestage + absLat + lifestage:absLat + (1|sp), data = as.data.frame(responses_table)) #

fit_red2random <- lmer(Max_dr ~ lifestage + absLat + (1|sp), data = as.data.frame(responses_table), REML = F)

anova(fit2_fullrandom, fit_red2random, test.statistic = "Chisq") # keep interaction model
```

```{r}

# fit better model with ML
fit2_i_randomML <- lmer(Max_dr ~  lifestage + absLat + lifestage:absLat + (1|sp), REML = F, data = as.data.frame(responses_table))
summary(fit2_i_randomML)
```

```{r}
# Plot data and predictions

responses_table$group <- responses_table$lifestage
predrate <- ggpredict(fit2_i_randomML, terms = c("absLat","lifestage"))


rate_plot <- ggplot(predrate, aes(x = x, y = predicted, colour = group, shape = group)) +
  geom_line()+
   geom_ribbon(aes(ymin = conf.low, ymax = conf.high, col = group), alpha = 0.1)+
  scale_color_viridis_d()+
  scale_fill_viridis_d()+
  #xlim(0,50)+
  scale_shape_manual(values = c(1,16,8))+
  #ylab(expression('T'[0]))+
  ylab("Development rate")+
  #annotate("text", x = 0, y = 16, label = "F", size = 5)+
  xlab("Absolute latitude")+
  geom_point(data = responses_table, mapping = aes(x = absLat, y = Max_dr, col =group ), alpha = 0.6)+
  theme_cowplot()+
  theme(legend.position = "none")

rate_plot
```

Add Dlow (run DR2_analyses.Rmd to have all data)
```{r}
T.set$group <- T.set$lifestage
Ttest <- ggplot(predto, aes(x = x, y = predicted, colour = group, fill = group)) +
  geom_line()+
  geom_line(data = predlow, mapping = aes(x = x, y = predicted, colour = group, fill = group))+
  geom_line(data = predto, mapping = aes(x = x, y = predicted, colour = group, fill = group))+
   #geom_ribbon(aes(ymin = conf.low, ymax = conf.high, col = group), alpha = 0.1)+
  scale_color_viridis_d()+
  scale_fill_viridis_d()+
  #scale_shape_manual(values = c(1,16,8))+
  ylab("T0 & Dlow")+
  #ylab(paste(expression(T[0]), "/", expression(D[low])))+
  xlab("Absolute latitude")+
  ylim(0, 40)+
  annotate("text", x = 0, y = 40, label = "A", size = 5)+
  geom_point(data = D.set, mapping = aes(x = absLat, y = Dlower, col = group), alpha = 0.6, shape =16)+
  
  geom_point(data = T.set, mapping = aes(x = absLat, y = T0, col = group), alpha = 0.6, shape = 8)+
  theme_cowplot()+
  theme(legend.title = element_blank(), legend.position = "none")
Ttest
```


